### 2014-12-06 paper.tex
added paper

%        File: paper.tex
%     Created: Sat Dec 06 04:00 PM 2014 E
%
\documentclass[letterpaper,twocolumn,11pt]{report}
% Change the header if you're going to change this.

% Possible packages - uncomment to use
\usepackage{amsmath}       % Needed for math stuff
%\usepackage{lastpage}      % Finds last page
\usepackage{amssymb}       % Needed for some math symbols
\usepackage{graphicx}      % Needed for graphics
\usepackage[usenames,dvipsnames]{xcolor} % Needed for graphics and color
%\usepackage{setspace}      % Needed to set single/double spacing
\usepackage{float}         % Better placement of figures & tables
\usepackage{hyperref}           % Can have actual links
\usepackage{mathpazo}

%Suggested by TeXworks
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{verbatim}      % lets you do verbatim text

% Sets the page margins to 1 inch each side
\usepackage[margin=1in]{geometry}
\geometry{letterpaper}
%\addtolength{\oddsidemargin}{-0.875in} 
%\addtolength{\evensidemargin}{-0.875in} 
%\addtolength{\textwidth}{1.75in} 
%\addtolength{\topmargin}{-.875in} 
%\addtolength{\textheight}{1.75in}

\frenchspacing

% Uncomment this section if you wish to have a header.
\usepackage{fancyhdr} 
\pagestyle{fancy} 
\renewcommand{\headrulewidth}{0.5pt} % customize the layout... 
\lhead{Chen, W., Sim, B., Shi, A.} \chead{Applied Math 205 Final Project} 
\rhead{Fall 2014} 
\lfoot{} \cfoot{\thepage} \rfoot{}

\begin{document}
\title{Parareal Algorithm}
\author{Wesley Chen, Brandon Sim, Andy Shi \\
Applied Math 205 Final Project, Fall 2014}
\date{December 15, 2014}
\maketitle

% No indentation for new paragraphs
\setlength\parindent{0pt}

% Space between each new paragraph
\setlength\parskip{2ex}

% Put your stuff here

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec a diam lectus.
Sed sit amet ipsum mauris. Maecenas congue ligula ac quam viverra nec
consectetur ante hendrerit. Donec et mollis dolor. Praesent et diam eget libero
egestas mattis sit amet vitae augue. Nam tincidunt congue enim, ut porta lorem
lacinia consectetur. Donec ut libero sed arcu vehicula ultricies a non tortor.
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean ut gravida
lorem. Ut turpis felis, pulvinar a semper sed, adipiscing id dolor. Pellentesque
auctor nisi id magna consequat sagittis. Curabitur dapibus enim sit amet elit
pharetra tincidunt feugiat nisl imperdiet. Ut convallis libero in urna ultrices
accumsan. Donec sed odio eros. Donec viverra mi quis quam pulvinar at malesuada
arcu rhoncus. Cum sociis natoque penatibus et magnis dis parturient montes,
nascetur ridiculus mus. In rutrum accumsan ultricies. Mauris vitae nisi at sem
facilisis semper ac in est.

Vivamus fermentum semper porta. Nunc diam velit, adipiscing ut tristique vitae,
sagittis vel odio. Maecenas convallis ullamcorper ultricies. Curabitur ornare,
ligula semper consectetur sagittis, nisi diam iaculis velit, id fringilla sem
nunc vel mi. Nam dictum, odio nec pretium volutpat, arcu ante placerat erat, non
tristique elit urna et turpis. Quisque mi metus, ornare sit amet fermentum et,
tincidunt et orci. Fusce eget orci a orci congue vestibulum. Ut dolor diam,
elementum et vestibulum eu, porttitor vel elit. Curabitur venenatis pulvinar
tellus gravida ornare. Sed et erat faucibus nunc euismod ultricies ut id justo.
Nullam cursus suscipit nisi, et ultrices justo sodales nec. Fusce venenatis
facilisis lectus ac semper. Aliquam at massa ipsum. Quisque bibendum purus
convallis nulla ultrices ultricies. Nullam aliquam, mi eu aliquam tincidunt,
purus velit laoreet tortor, viverra pretium nisi quam vitae mi. Fusce vel
volutpat elit. Nam sagittis nisi dui.

Suspendisse lectus leo, consectetur in tempor sit amet, placerat quis neque.
Etiam luctus porttitor lorem, sed suscipit est rutrum non. Curabitur lobortis
nisl a enim congue semper. Aenean commodo ultrices imperdiet. Vestibulum ut
justo vel sapien venenatis tincidunt. Phasellus eget dolor sit amet ipsum
dapibus condimentum vitae quis lectus. Aliquam ut massa in turpis dapibus
convallis. Praesent elit lacus, vestibulum at malesuada et, ornare et est. Ut
augue nunc, sodales ut euismod non, adipiscing vitae orci. Mauris ut placerat
justo. Mauris in ultricies enim. Quisque nec est eleifend nulla ultrices egestas
quis ut quam. Donec sollicitudin lectus a mauris pulvinar id aliquam urna
cursus. Cras quis ligula sem, vel elementum mi. Phasellus non ullamcorper urna.

Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos
himenaeos. In euismod ultrices facilisis. Vestibulum porta sapien adipiscing
augue congue id pretium lectus molestie. Proin quis dictum nisl. Morbi id quam
sapien, sed vestibulum sem. Duis elementum rutrum mauris sed convallis. Proin
vestibulum magna mi. Aenean tristique hendrerit magna, ac facilisis nulla
hendrerit ut. Sed non tortor sodales quam auctor elementum. Donec hendrerit nunc
eget elit pharetra pulvinar. Suspendisse id tempus tortor. Aenean luctus, elit
commodo laoreet commodo, justo nisi consequat massa, sed vulputate quam urna
quis eros. Donec vel.

\end{document}

### 2014-12-06 introToMPI.py

@@ -7,7 +7,7 @@ import time
 
 def get_big_arrays():
   '''Generate two big random arrays.'''
-  N = 10000000       # A big number, the size of the arrays.
+  N = 100       # A big number, the size of the arrays.
   seed = 42;
   np.random.seed(seed)  # Set the random seed
   return np.random.random(N), np.random.random(N)

### 2014-12-06 paper.tex
latex template start

@@
 %
-\documentclass[letterpaper,twocolumn,11pt]{report}
+\documentclass[letterpaper,twocolumn,11pt]{article}
 % Change the header if you're going to change this.
@@
 \usepackage{mathpazo}
+\usepackage[linesnumbered,vlined,ruled]{algorithm2e} % algorithm
 
@@
 \author{Wesley Chen, Brandon Sim, Andy Shi \\
-Applied Math 205 Final Project, Fall 2014}
+Harvard University, Applied Math 205 Final Project}
 \date{December 15, 2014}
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec a diam lectus.
-Sed sit amet ipsum mauris. Maecenas congue ligula ac quam viverra nec
-consectetur ante hendrerit. Donec et mollis dolor. Praesent et diam eget libero
-egestas mattis sit amet vitae augue. Nam tincidunt congue enim, ut porta lorem
-lacinia consectetur. Donec ut libero sed arcu vehicula ultricies a non tortor.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aenean ut gravida
-lorem. Ut turpis felis, pulvinar a semper sed, adipiscing id dolor. Pellentesque
-auctor nisi id magna consequat sagittis. Curabitur dapibus enim sit amet elit
-pharetra tincidunt feugiat nisl imperdiet. Ut convallis libero in urna ultrices
-accumsan. Donec sed odio eros. Donec viverra mi quis quam pulvinar at malesuada
-arcu rhoncus. Cum sociis natoque penatibus et magnis dis parturient montes,
-nascetur ridiculus mus. In rutrum accumsan ultricies. Mauris vitae nisi at sem
-facilisis semper ac in est.
+\section{Abstract}
 
-Vivamus fermentum semper porta. Nunc diam velit, adipiscing ut tristique vitae,
-sagittis vel odio. Maecenas convallis ullamcorper ultricies. Curabitur ornare,
-ligula semper consectetur sagittis, nisi diam iaculis velit, id fringilla sem
-nunc vel mi. Nam dictum, odio nec pretium volutpat, arcu ante placerat erat, non
-tristique elit urna et turpis. Quisque mi metus, ornare sit amet fermentum et,
-tincidunt et orci. Fusce eget orci a orci congue vestibulum. Ut dolor diam,
-elementum et vestibulum eu, porttitor vel elit. Curabitur venenatis pulvinar
-tellus gravida ornare. Sed et erat faucibus nunc euismod ultricies ut id justo.
-Nullam cursus suscipit nisi, et ultrices justo sodales nec. Fusce venenatis
-facilisis lectus ac semper. Aliquam at massa ipsum. Quisque bibendum purus
-convallis nulla ultrices ultricies. Nullam aliquam, mi eu aliquam tincidunt,
-purus velit laoreet tortor, viverra pretium nisi quam vitae mi. Fusce vel
-volutpat elit. Nam sagittis nisi dui.
+We LIKE TO PLAY LEAGUE.
 
-Suspendisse lectus leo, consectetur in tempor sit amet, placerat quis neque.
-Etiam luctus porttitor lorem, sed suscipit est rutrum non. Curabitur lobortis
-nisl a enim congue semper. Aenean commodo ultrices imperdiet. Vestibulum ut
-justo vel sapien venenatis tincidunt. Phasellus eget dolor sit amet ipsum
-dapibus condimentum vitae quis lectus. Aliquam ut massa in turpis dapibus
-convallis. Praesent elit lacus, vestibulum at malesuada et, ornare et est. Ut
-augue nunc, sodales ut euismod non, adipiscing vitae orci. Mauris ut placerat
-justo. Mauris in ultricies enim. Quisque nec est eleifend nulla ultrices egestas
-quis ut quam. Donec sollicitudin lectus a mauris pulvinar id aliquam urna
-cursus. Cras quis ligula sem, vel elementum mi. Phasellus non ullamcorper urna.
+\section{Background}
 
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos
-himenaeos. In euismod ultrices facilisis. Vestibulum porta sapien adipiscing
-augue congue id pretium lectus molestie. Proin quis dictum nisl. Morbi id quam
-sapien, sed vestibulum sem. Duis elementum rutrum mauris sed convallis. Proin
-vestibulum magna mi. Aenean tristique hendrerit magna, ac facilisis nulla
-hendrerit ut. Sed non tortor sodales quam auctor elementum. Donec hendrerit nunc
-eget elit pharetra pulvinar. Suspendisse id tempus tortor. Aenean luctus, elit
-commodo laoreet commodo, justo nisi consequat massa, sed vulputate quam urna
-quis eros. Donec vel.
+League is cool.
 
+\section{Methodology}
+
+\begin{algorithm}[t]
+ \KwIn{Replication parameter $C$.}
+ \KwIn{$P$ processors arranged into $C$ rows and $T=P/C$ teams.}
+ \KwIn{Set $\mathcal{T}$ of $N$ particles and associated charges and results.}
+ In parallel on all processors, $r = $ rank\;
+ Scatter particles from master to team leaders, $\mathcal{T}_t$\;
+ Broadcast $\mathcal{T}_t$ from team leader to team members\;
+ Copy $\mathcal{T}_t$, the set of targets, to $\mathcal{S}_t$, the set of sources\;
+ Given $c$th member of team $t$, shift $\mathcal{S}_t$ by $c$ along row\;
+ \For{\normalfont{iteration} $i < 1$}{
+   Compute symmetric interaction of $\mathcal{T}_t$ and $\mathcal{S}_t$\;
+ }
+ Reduce the results within a team\;
+ Gather the results between team leaders\;
+ \caption{Symmetric Diagonal $N$-body}
+ \label{alg:symmdiag}
+\end{algorithm}
+
+ROAR
 \end{document}

### 2014-12-06 paper.tex
added algorithm, removed bunch of stuff

@@
 \date{December 15, 2014}
-\maketitle
 
@@
 % Put your stuff here
+\twocolumn[
+    \begin{@twocolumnfalse}
+    \maketitle
+    \begin{abstract}
 
-\section{Abstract}
+    Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec a diam
+    lectus.  Sed sit amet ipsum mauris. Maecenas congue ligula ac quam viverra
+    nec consectetur ante hendrerit. Donec et mollis dolor. Praesent et diam eget
+    libero egestas mattis sit amet vitae augue. Nam tincidunt congue enim, ut
+    porta lorem lacinia consectetur. Donec ut libero sed arcu vehicula ultricies
+    a non tortor.  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+    Aenean ut gravida lorem. Ut turpis felis, pulvinar a semper sed, adipiscing
+    id dolor. Pellentesque auctor nisi id magna consequat sagittis. Curabitur
+    dapibus enim sit amet elit pharetra tincidunt feugiat nisl imperdiet. Ut
+    convallis libero in urna ultrices accumsan.  Donec sed odio eros. Donec
+    viverra mi quis quam pulvinar at malesuada arcu rhoncus. Cum sociis natoque
+    penatibus et magnis dis parturient montes, nascetur ridiculus mus. In rutrum
+    accumsan ultricies. Mauris vitae nisi at sem facilisis semper ac in est.
 
-We LIKE TO PLAY LEAGUE.
+    \end{abstract}
+  \end{@twocolumnfalse}
+]
 
@@
 \begin{algorithm}[t]
- \KwIn{Replication parameter $C$.}
- \KwIn{$P$ processors arranged into $C$ rows and $T=P/C$ teams.}
- \KwIn{Set $\mathcal{T}$ of $N$ particles and associated charges and results.}
- In parallel on all processors, $r = $ rank\;
- Scatter particles from master to team leaders, $\mathcal{T}_t$\;
- Broadcast $\mathcal{T}_t$ from team leader to team members\;
- Copy $\mathcal{T}_t$, the set of targets, to $\mathcal{S}_t$, the set of sources\;
- Given $c$th member of team $t$, shift $\mathcal{S}_t$ by $c$ along row\;
- \For{\normalfont{iteration} $i < 1$}{
-   Compute symmetric interaction of $\mathcal{T}_t$ and $\mathcal{S}_t$\;
+    \KwIn{Temporal discretization $t^n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
+    \KwIn{Coarse scheme $g_{\Delta t}$}
+    \KwIn{Finer scheme $g_{\textnormal{fine}}$}
+    Compute $u_1^{n+1} = g_{\Delta t}(t^n, u^n)$\;
+    Compute the corrections $\delta g^n(u_1^n) = g_{\textnormal{fine}}(t^n,
+    u^n) - g_{\Delta t}(t^n, u^n)$ in parallel\;
+    Add the prediction and correction terms as $u_2^{n+1} = g_{\Delta
+    t}(t^n, u_2^n) + \delta g^n(u_1^n)$\;
+    Repeat steps 2 and 3, incrementing the iteration label and using $u_{k+1}^0
+    = u^0$ as the initial condition\
     }
- Reduce the results within a team\;
- Gather the results between team leaders\;
- \caption{Symmetric Diagonal $N$-body}
- \label{alg:symmdiag}
+ \caption{Parareal}
+ \label{alg:parareal}
 \end{algorithm}

### 2014-12-06 paper.tex
outline of paper

--- 2014-12-06 paper.tex
@@
 \begin{document}
-\title{Parareal Algorithm}
+\title{Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications}
 \author{Wesley Chen, Brandon Sim, Andy Shi \\
@@
 
-We LIKE TO PLAY LEAGUE.
+We implement the parareal algorithm to solve various ODEs and compare in terms of accuracy, speedup and efficiency.
+We explore the effect of using differnt solution operators (a course and a fine as required by the method)  The majority of our code is written in python using mpi4py to parallelize.
+To compare to more performance-based languages, we compare certain codes to an implementation in C++.
 
@@
 
-League is cool.
+\subsection{Numerical Methods and Parallelization}
+Solving systems of differential equations can be a computationally expensive task.
+The error of most algorithms scales on the stepsize of the discretization of time.
+However, stepsize in time is also proportional to the computation required.
+The parareal algorithm, developed by YADA YADA.
+An alternative would be to parallelize by space -
+
+\subsection{Stability of the Parareal}
+
+\subsection{Error of the Parareal}
 
 \section{Methodology}
+
+The general parareal method, as described by YADA YADA can be SEEN BELOW YAR
 
@@
 
-ROAR
+\subsection{Explored Tests}
+We wished to explore certain tests
+
+\section{Results}
+We compared various runs of
+
+\subsection{Comparison to Serial with $G_{fine}$ as \\a Forward Euler with Smaller $\Delta t$}
+
+\subsection{Comparison to Serial with $G_{fine}$ as \\Higher Order Methods}
+
+\subsection{Comparison to a Forward Euler \\Parallelized by Space}
+
+\subsection{Comparison to C++ Implementations}
+
+\subsection{Summary of Performance}
+
+\section{Discussion}
+
+\section{Conclusion and Future Work}
+
+\section{References}
+
 \end{document}

### 2014-12-06 paper.tex
merge conflict fix

@@
 % Put your stuff here
-
-\section{Abstract}
-
+\twocolumn[
+    \begin{@twocolumnfalse}
+    \maketitle
+    \begin{abstract}
 We implement the parareal algorithm to solve various ODEs and compare in terms of accuracy, speedup and efficiency.
@@
 To compare to more performance-based languages, we compare certain codes to an implementation in C++.
-
+    \end{abstract}
+  \end{@twocolumnfalse}
+]
 \section{Background}

### 2014-12-06 paper.tex
minor tweaks on paper, added forward euler

@@
     \begin{abstract}
-We implement the parareal algorithm to solve various ODEs and compare in terms of accuracy, speedup and efficiency.
-We explore the effect of using differnt solution operators (a course and a fine as required by the method)  The majority of our code is written in python using mpi4py to parallelize.
-To compare to more performance-based languages, we compare certain codes to an implementation in C++.
+    We implement the parareal algorithm to solve various ODEs and compare in
+    terms of accuracy, speedup and efficiency. We explore the effect of using
+    differnt solution operators (a coarse and a fine as required by the method)
+    The majority of our code is written in python using mpi4py to parallelize.
+    To compare to more performance-based languages, we compare certain codes to
+    an implementation in C++.
     \end{abstract}
@@
 \subsection{Numerical Methods and Parallelization}
-Solving systems of differential equations can be a computationally expensive task.
-The error of most algorithms scales on the stepsize of the discretization of time.
-However, stepsize in time is also proportional to the computation required.
-The parareal algorithm, developed by YADA YADA.
-An alternative would be to parallelize by space -
+Solving systems of differential equations can be a computationally expensive
+task. The error of most algorithms scales on the stepsize of the discretization
+of time. However, stepsize in time is also proportional to the computation
+required. The parareal algorithm, developed by YADA YADA. An alternative would
+be to parallelize by space -
 
@@
 
-\subsection{Comparison to Serial with $G_{fine}$ as \\a Forward Euler with Smaller $\Delta t$}
+\subsection{Comparison to Serial with $g_{fine}$ as \\a Forward Euler with
+Smaller $\Delta t$}
 
-\subsection{Comparison to Serial with $G_{fine}$ as \\Higher Order Methods}
+\subsection{Comparison to Serial with $g_{fine}$ as \\Higher Order Methods}
 

### 2014-12-06 serial_euler.py

@@ -1,0 +1,50 @@
+import numpy as np
+
+def forward_euler(deriv, init, time_min, time_max, time_step):
+    '''
+    Solves a differential equation using Euler's method from time_min to
+    time_max, using a step of time_step
+
+    deriv: callable; the derivative
+    init: numpy 1D array; initial condition
+    time_min: float; Minimum time
+    time_max: float; Maximum time
+    time_step: float; step size
+    '''
+    dim = init.shape[0]
+    # checks for 1D array
+    assert(init.shape == (dim,))
+
+    # initialize times
+    nsteps = (float(time_max - time_min) / time_step) + 1
+    times = np.linspace(time_min, time_max, nsteps)[1:]
+
+    ys = np.zeros((nsteps, dim))
+    ys[0] = init
+
+    for (ind, t) in enumerate(times):
+        time_ind = ind + 1
+        curr = ys[time_ind-1]
+        ys[time_ind] = curr + time_step * deriv(t, curr)
+
+    return ys
+
+def test():
+    lam = 1.
+
+    def deriv(t, u):
+        return lam * u
+
+    init = np.array([1.])
+
+    res = forward_euler(deriv, init, 0, 1, 0.01)[:,0]
+    
+    exact = np.exp(np.linspace(0,1,len(res)))
+    print exact - res
+
+
+def main():
+    test()
+
+if __name__ == "__main__":
+    main()

### 2014-12-07 introToMPI.py
code structure for parareal temporary push

@@ -5,19 +5,7 @@ from mpi4py import MPI
 import numpy as np
 import time
 
-def get_big_arrays():
-  '''Generate two big random arrays.'''
-  N = 10000      # A big number, the size of the arrays.
-  seed = 42;
-  np.random.seed(seed)  # Set the random seed
-  return np.random.random(N), np.random.random(N)
-
-def serial_dot(a, b):
-  '''The dot-product of the arrays -- slow implementation using for-loop.'''
-  result = 0
-  for k in xrange(0, len(a)):
-    result += a[k] * b[k]
-  return result
+#from serial_euler import forward_euler
 
 def parallel_dot(a, b, comm, p_root=0):
   '''The parallel dot-product of the arrays a and b.

### 2014-12-07 pararealAlgo.py

@@ -1,48 +1,65 @@
-# Example to test python and MPI install from Wesley
+# Generic Parareal Implementation
 # Updated 12/6/2014
 
 from mpi4py import MPI
 import numpy as np
 import time
 
-def get_big_arrays():
-  '''Generate two big random arrays.'''
-  N = 100000       # A big number, the size of the arrays.
-  seed = 42;
-  np.random.seed(seed)  # Set the random seed
-  return np.random.random(N), np.random.random(N)
+from serial_euler import forward_euler
 
-def serial_dot(a, b):
-  '''The dot-product of the arrays -- slow implementation using for-loop.'''
-  result = 0
-  for k in xrange(0, len(a)):
-    result += a[k] * b[k]
-  return result
+# Parareal begin
+def startParareal(deriv, init, k, tmin, tmax, numSteps, comm):
+  '''Starts running k iterations of the parareal algorithm.
+  In each iteration, the program runs the coarse method in serial, then fine corrections in parallel'''
 
-def parallel_dot(a, b, comm, p_root=0):
-  '''The parallel dot-product of the arrays a and b.
+  # get a coarse answer
+  u = serial_coarse(deriv, init, tmin, tmax, numSteps)[:,0]
+
+  # fine time steps
+  qualityFactor = 100
+  fine = numSteps * qualityFactor
+
+  # rethink logic here!
+  if k > 1:
+    for iteration in xrange(k-1):
+
+      # get a coarse answer
+      u = serial_coarse(deriv, init, tmin, tmax, numSteps)[:,0]
+
+      # pass in a coarse answer to refine and correct
+      corrections = parallel_corrections(deriv, u, time, comm, p_root=0)
+
+      # update u with the corrections
+      u += corrections
+
+  return u
+
+def serial_coarse(deriv, init, tmin, tmax, numSteps):
+  '''Runs a coarse numerical method in serial, as part of the parareal algorithm'''
+  # call forward euler
+  return forward_euler(deriv, init, tmin, tmax, numSteps)
+
+def parallel_corrections(u, tmin, tmax, numSteps, comm, p_root=0):
+  '''Parallel Portion of the Parareal Algorithm
   Assumes the arrays exist on process p_root and returns the result to
-  process p_root.
+  process p_root for a given iteration.
   By default, p_root = process 0.'''
   rank = comm.Get_rank()
   size = comm.Get_size()
 
[…]

### 2014-12-07 serial_euler.py

@@ -1,6 +1,6 @@
 import numpy as np
 
-def forward_euler(deriv, init, time_min, time_max, time_step):
+def forward_euler(deriv, init, time_min, time_max, numSteps):
     '''
     Solves a differential equation using Euler's method from time_min to
     time_max, using a step of time_step
@@ -9,15 +9,14 @@ def forward_euler(deriv, init, time_min, time_max, time_step):
     init: numpy 1D array; initial condition
     time_min: float; Minimum time
     time_max: float; Maximum time
-    time_step: float; step size
+    numSteps: int; number of stes
     '''
     dim = init.shape[0]
     # checks for 1D array
     assert(init.shape == (dim,))
 
     # initialize times
-    nsteps = (float(time_max - time_min) / time_step) + 1
-    times = np.linspace(time_min, time_max, nsteps)[1:]
+    times = np.linspace(time_min, time_max, numSteps)[1:]
 
     ys = np.zeros((nsteps, dim))
     ys[0] = init
@@ -29,7 +28,8 @@ def forward_euler(deriv, init, time_min, time_max, time_step):
 
     return ys
 
-def test():
+if __name__ == "__main__":
+
     lam = 1.
 
     def deriv(t, u):
@@ -40,11 +40,5 @@ def test():
     res = forward_euler(deriv, init, 0, 1, 0.01)[:,0]
     
     exact = np.exp(np.linspace(0,1,len(res)))
+    print res
     print exact - res
-
-
-def main():
-    test()
-
-if __name__ == "__main__":
-    main()

### 2014-12-07 pararealAlgo.py
basic logic of parareal complete - uncompiled

@@ -1,68 +1,111 @@
 # Generic Parareal Implementation
 # Updated 12/6/2014
 
+# TODO 
+# Confirm that Euler_Step function does not repeat calculations - else should memoize and look up
+
 from mpi4py import MPI
 import numpy as np
 import time
+import copy
 
 from serial_euler import forward_euler
+from serial_euler import forward_euler_step
 
 # Parareal begin
-def startParareal(deriv, init, k, tmin, tmax, numSteps, comm):
+def startParareal(deriv, init, k, tmin, tmax, numSteps, qualityFactor = 100, comm):
   '''Starts running k iterations of the parareal algorithm.
-  In each iteration, the program runs the coarse method in serial, then fine corrections in parallel'''
+  In each iteration, the program runs the coarse method in serial, then fine corrections in parallel
+  The result is only gathered to the root process at every iteration'''
 
-  # get a coarse answer
-  u = serial_coarse(deriv, init, tmin, tmax, numSteps)[:,0]
-
-  # fine time steps
-  qualityFactor = 100
-  fine = numSteps * qualityFactor
-
-  # rethink logic here!
-  if k > 1:
-    for iteration in xrange(k-1):
-
-      # get a coarse answer
-      u = serial_coarse(deriv, init, tmin, tmax, numSteps)[:,0]
-
-      # pass in a coarse answer to refine and correct
-      corrections = parallel_corrections(deriv, u, time, comm, p_root=0)
-
-      # update u with the corrections
-      u += corrections
-
-  return u
-
-def serial_coarse(deriv, init, tmin, tmax, numSteps):
-  '''Runs a coarse numerical method in serial, as part of the parareal algorithm'''
-  # call forward euler
-  return forward_euler(deriv, init, tmin, tmax, numSteps)
-
-def parallel_corrections(u, tmin, tmax, numSteps, comm, p_root=0):
-  '''Parallel Portion of the Parareal Algorithm
-  Assumes the arrays exist on process p_root and returns the result to
-  process p_root for a given iteration.
-  By default, p_root = process 0.'''
   rank = comm.Get_rank()
   size = comm.Get_size()
 
-  # Save the number of tasks to a variable
-  # numtasks = len(u)
+  dt = (tmax-tmin) / (numSteps - 1)
 
-  # Start and end indices of the local dot product computed by function to allow for undivisble sizes
+  # upast and unext vectors
+  upast = np.zeros(numSteps)
+  unext = np.zeros(numSteps)
+
[…]

### 2014-12-07 pararealAlgo.py
quick fix

@@ -58,7 +58,7 @@ def serial_coarse(deriv, upast, dt, tmin, tmax, numSteps):
   # call forward euler per step and store in vector
   for ind in xrange(numSteps):
     # get a coarse answer
-    upast[ind] = serial_coarse(deriv, times[ind], u[ind], dt)
+    upast[ind] = forward_euler_step(deriv, times[ind], u[ind], dt)
   return upast
 
 def parallel_corrections(upast, tmin, tmax, numSteps, qualityFactor, comm, p_root=0):

### 2014-12-07 pararealAlgo.py
another typo to fix

@@ -58,7 +58,7 @@ def serial_coarse(deriv, upast, dt, tmin, tmax, numSteps):
   # call forward euler per step and store in vector
   for ind in xrange(numSteps):
     # get a coarse answer
-    upast[ind] = forward_euler_step(deriv, times[ind], u[ind], dt)
+    upast[ind] = forward_euler_step(deriv, times[ind], upast[ind], dt)
   return upast
 
 def parallel_corrections(upast, tmin, tmax, numSteps, qualityFactor, comm, p_root=0):

### 2014-12-07 pararealAlgo.py
and one more...

@@ -58,6 +58,7 @@ def serial_coarse(deriv, upast, dt, tmin, tmax, numSteps):
   # call forward euler per step and store in vector
   for ind in xrange(numSteps):
     # get a coarse answer
+    # TODO is this upast[ind] off by one?
     upast[ind] = forward_euler_step(deriv, times[ind], upast[ind], dt)
   return upast
 

### 2014-12-07 serial_euler.py

@@ -36,7 +36,7 @@ if __name__ == "__main__":
         return lam * u
 
     init = np.array([1.])
-
+    
     res = forward_euler(deriv, init, 0, 1, 0.01)[:,0]
     
     exact = np.exp(np.linspace(0,1,len(res)))

### 2014-12-07 pararealAlgo.py
made some changes. hoping

@@ -1,9 +1,6 @@
 # Generic Parareal Implementation
 # Updated 12/6/2014
 
-# TODO 
-# Confirm that Euler_Step function does not repeat calculations - else should memoize and look up
-
 from mpi4py import MPI
 import numpy as np
 import time
@@ -27,6 +24,9 @@ def startParareal(deriv, init, k, tmin, tmax, numSteps, qualityFactor = 100, com
   upast = np.zeros(numSteps)
   unext = np.zeros(numSteps)
 
+  # all coarse
+  upast = forward_euler(deriv, init, tmin, tmax, numSteps)
+
   # if further iterations are required
   if k > 1:
     for iteration in xrange(k-1):
@@ -36,31 +36,36 @@ def startParareal(deriv, init, k, tmin, tmax, numSteps, qualityFactor = 100, com
 
       if rank == 0:
         # get a coarse answer for the next iteration, vectorized
-        unext = serial_coarse(deriv, upast, dt, tmin, tmax, numSteps)[:,0]
+        unext = serial_coarse_with_correction(deriv, init, corrections, dt, tmin, tmax, numSteps)
 
         # update unext for next iteration
         upast = copy.copy(unext)
 
-        # update unext with corrections computed with upast
-        # TODO check the edge case - this is correct?
-        for ind in xrange(numSteps-1):
-          unext[ind+1] += unext[ind] + corrections[ind]
-
   return unext
 
-def serial_coarse(deriv, upast, dt, tmin, tmax, numSteps):
+def serial_coarse_with_correction(deriv, init, corrections, dt, tmin, tmax, numSteps):
   '''Runs a coarse numerical method in serial, as part of the parareal algorithm.
   Will call the stepwise g_coarse function per step'''
+
+  unext = np.zeros(numSteps)
 
   # initialize times
   times = np.linspace(time_min, time_max, numSteps)[1:]
 
+  # initial condition
+  unext[0] = init
+
   # call forward euler per step and store in vector
-  for ind in xrange(numSteps):
+  for ind in xrange(numSteps-1):
     # get a coarse answer
     # TODO is this upast[ind] off by one?
-    upast[ind] = forward_euler_step(deriv, times[ind], upast[ind], dt)
-  return upast
+    unext[ind+1] = forward_euler_step(deriv, times[ind], unext[ind], dt)
+
+    # update unext with corrections computed with upast
+    for ind in xrange(numSteps-1):
+      unext[ind+1] += unext[ind] + corrections[ind]
+
+  return unext
 
[…]

### 2014-12-07 pararealAlgo.py
it runs now

@@ -10,7 +10,8 @@ from serial_euler import forward_euler
 from serial_euler import forward_euler_step
 
 # Parareal begin
-def startParareal(deriv, init, k, tmin, tmax, numSteps, qualityFactor = 100, comm):
+def startParareal(deriv, init, k, tmin, tmax, numSteps, comm, 
+        qualityFactor = 100):
   '''Starts running k iterations of the parareal algorithm.
   In each iteration, the program runs the coarse method in serial, then fine corrections in parallel
   The result is only gathered to the root process at every iteration'''
@@ -20,12 +21,9 @@ def startParareal(deriv, init, k, tmin, tmax, numSteps, qualityFactor = 100, com
 
   dt = (tmax-tmin) / (numSteps - 1)
 
-  # upast and unext vectors
-  upast = np.zeros(numSteps)
-  unext = np.zeros(numSteps)
-
   # all coarse
   upast = forward_euler(deriv, init, tmin, tmax, numSteps)
+  unext = None
 
   # if further iterations are required
   if k > 1:
@@ -47,27 +45,30 @@ def serial_coarse_with_correction(deriv, init, corrections, dt, tmin, tmax, numS
   '''Runs a coarse numerical method in serial, as part of the parareal algorithm.
   Will call the stepwise g_coarse function per step'''
 
-  unext = np.zeros(numSteps)
+  fixed_corrections = np.empty((0, init.shape[0]))
+  # corrections is a list
+  for arr in corrections:
+    fixed_corrections = np.concatenate((fixed_corrections, arr))
+  unext = np.zeros((numSteps+1, init.shape[0]))
 
   # initialize times
-  times = np.linspace(time_min, time_max, numSteps)[1:]
+  times = np.linspace(tmin, tmax, numSteps+1)
 
   # initial condition
   unext[0] = init
 
   # call forward euler per step and store in vector
-  for ind in xrange(numSteps-1):
+  for ind in xrange(numSteps):
     # get a coarse answer
     # TODO is this upast[ind] off by one?
-    unext[ind+1] = forward_euler_step(deriv, times[ind], unext[ind], dt)
+    unext[ind] = forward_euler_step(deriv, times[ind], unext[ind], dt)
 
     # update unext with corrections computed with upast
-    for ind in xrange(numSteps-1):
-      unext[ind+1] += unext[ind] + corrections[ind]
+    unext[ind+1] = unext[ind] + fixed_corrections[ind]
 
   return unext
 
[…]

### 2014-12-07 serial_euler.py

@@ -1,5 +1,8 @@
 import numpy as np
 
+def forward_euler_step(deriv, t, u, dt):
+    return u + dt * deriv(t, u)
+
 def forward_euler(deriv, init, time_min, time_max, numSteps):
     '''
     Solves a differential equation using Euler's method from time_min to
@@ -16,15 +19,16 @@ def forward_euler(deriv, init, time_min, time_max, numSteps):
     assert(init.shape == (dim,))
 
     # initialize times
-    times = np.linspace(time_min, time_max, numSteps)[1:]
+    times = np.linspace(time_min, time_max, numSteps+1)[1:]
+    time_step = float(time_max - time_min) / (numSteps - 1)
 
-    ys = np.zeros((nsteps, dim))
+    ys = np.zeros((numSteps+1, dim))
     ys[0] = init
 
     for (ind, t) in enumerate(times):
         time_ind = ind + 1
         curr = ys[time_ind-1]
-        ys[time_ind] = curr + time_step * deriv(t, curr)
+        ys[time_ind] = forward_euler_step(deriv, t, curr, time_step)
 
     return ys
 
@@ -35,10 +39,12 @@ if __name__ == "__main__":
     def deriv(t, u):
         return lam * u
 
+    nsteps = 200
     init = np.array([1.])
     
-    res = forward_euler(deriv, init, 0, 1, 0.01)[:,0]
-    
-    exact = np.exp(np.linspace(0,1,len(res)))
+    res = forward_euler(deriv, init, 0, 1, nsteps)[:,0]
     print res
+    print len(res)
+    
+    exact = np.exp(np.linspace(0,1,nsteps+1))
     print exact - res

### 2014-12-08 paper.tex
a note about speedups in tex
added poster, modified algo

@@
     = u^0$ as the initial condition\
-    }
  \caption{Parareal}
@@
 \subsection{Comparison to C++ Implementations}
+	The numpy environment combined with the required structure of MPI required certain type/data structure conversions which will take time.not not cunot cu
 

### 2014-12-08 poster/main.tex
added poster, modified algo

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.1 (14/06/14)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{48in} % A0 width: 46.8in
\setlength{\paperheight}{36in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables
\usepackage[linesnumbered,vlined,ruled]{algorithm2e} % algorithm

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{The Parareal Algorithm for Parallel Solution of ODEs} % Poster title

\author{Wesley Chen, Brandon Sim, Andy Shi} % Author(s)

\institute{Harvard University, Applied Math 205 Final Project} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Abstract}

We implement the parareal algorithm to solve various ODEs and compare in terms
of accuracy, speedup and efficiency. We explore the effect of using differnt
solution operators (a coarse and a fine as required by the method) The majority
of our code is written in python using mpi4py to parallelize.  To compare to
more performance-based languages, we compare certain codes to an implementation
in C++.

\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}

Solving systems of differential equations can be a computationally expensive
task. The error of most algorithms scales on the stepsize of the discretization
of time. However, stepsize in time is also proportional to the computation
required. The parareal algorithm, developed by YADA YADA. An alternative would
be to parallelize by space.

%------------------------------------------------

\begin{figure}
\includegraphics[width=0.8\linewidth]{harvard_univ_logo.png}
\caption{Figure caption}
\end{figure}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column

\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------

\begin{block}{Algorithm}
The general parareal method, as described by YADA YADA can be SEEN BELOW YAR

Input: Temporal discretization $t^n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$
Input: Coarse scheme $g_{\Delta t}$
Input: Finer scheme $g_{\textnormal{fine}}$
Compute $u_1^{n+1} = g_{\Delta t}(t^n, u^n)$
Compute the corrections $\delta g^n(u_1^n) = g_{\textnormal{fine}}(t^n,
u^n) - g_{\Delta t}(t^n, u^n)$ in parallel
Add the prediction and correction terms as $u_2^{n+1} = g_{\Delta
t}(t^n, u_2^n) + \delta g^n(u_1^n)$
Repeat steps 2 and 3, incrementing the iteration label and using $u_{k+1}^0
= u^0$ as the initial condition
\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------

\begin{block}{Methods}

Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed laoreet accumsan mattis. Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem. Praesent dictum tempor pulvinar. Suspendisse potenti. Sed tincidunt varius ipsum, et porta nulla suscipit et. Etiam congue bibendum felis, ac dictum augue cursus a. \textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est. In ut orci purus, interdum ornare nibh. Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla. Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam. Maecenas sed diam augue.

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width

%----------------------------------------------------------------------------------------
%	IMPORTANT RESULT
%----------------------------------------------------------------------------------------

% \begin{alertblock}{Important Result}

% Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit. Sed commodo molestie porta. Sed ultrices scelerisque sapien ac commodo. Donec ut volutpat elit.

% \end{alertblock}

%----------------------------------------------------------------------------------------

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again

\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATHEMATICAL SECTION
%----------------------------------------------------------------------------------------

%\begin{block}{Mathematical Section}

%Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.

%\begin{equation}
%E = mc^{2}
%\label{eqn:Einstein}
%\end{equation}
%
%Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
%
%\begin{equation}
%\cos^3 \theta =\frac{1}{4}\cos\theta+\frac{3}{4}\cos 3\theta
%\label{eq:refname}
%\end{equation}
%
%Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
%
%\begin{equation}
%\kappa =\frac{\xi}{E_{\mathrm{max}}} %\mathbb{ZNR}
%\end{equation}
%
%\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\begin{column}{\onecolwid} % The second column within column 2 (column 2.2)

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

\begin{block}{Results}

\begin{figure}
\includegraphics[width=0.8\linewidth]{harvard_univ_logo.png}
\caption{Figure caption}
\end{figure}

Nunc tempus venenatis facilisis. Curabitur suscipit consequat eros non porttitor. Sed a massa dolor, id ornare enim:

\begin{table}
\vspace{2ex}
\begin{tabular}{l l l}
\toprule
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\midrule
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\bottomrule
\end{tabular}
\caption{Table caption}
\end{table}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.2

\end{columns} % End of the split of column 2

\end{column} % End of the second column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The third column

%----------------------------------------------------------------------------------------
%	CONCLUSION
%----------------------------------------------------------------------------------------

\begin{block}{Conclusion}

Nunc tempus venenatis facilisis. \textbf{Curabitur suscipit} consequat eros non porttitor. Sed a massa dolor, id ornare enim. Fusce quis massa dictum tortor \textbf{tincidunt mattis}. Donec quam est, lobortis quis pretium at, laoreet scelerisque lacus. Nam quis odio enim, in molestie libero. Vivamus cursus mi at \textit{nulla elementum sollicitudin}.

\end{block}

%----------------------------------------------------------------------------------------
%	ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Additional Information}

Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum. 
\begin{itemize}
\item Curabitur pellentesque dignissim
\item Eu facilisis est tempus quis
\item Duis porta consequat lorem
\end{itemize}

\end{block}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\begin{block}{References}

\nocite{*} % Insert publications even if they are not cited in the poster
\small{\bibliographystyle{unsrt}
\bibliography{sample}\vspace{0.75in}}

\end{block}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color

\begin{block}{Acknowledgements}

\small{\rmfamily{Nam mollis tristique neque eu luctus. Suspendisse rutrum congue nisi sed convallis. Aenean id neque dolor. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\

\end{block}

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors

\begin{alertblock}{Contact Information}

\begin{itemize}
\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
\item Email: \href{mailto:john@smith.com}{john@smith.com}
\item Phone: +1 (000) 111 1111
\end{itemize}

\end{alertblock}

\begin{center}
\begin{tabular}{ccc}
%\includegraphics[width=0.4\linewidth]{harvard_univ_logo.png} & \hfill &
\includegraphics[width=0.2\linewidth]{seas-rgb.png}
\end{tabular}
\end{center}

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2014-12-08 pararealAlgo.py

@@ -38,6 +38,8 @@ def startParareal(deriv, init, k, tmin, tmax, numSteps, comm,
 
         # update unext for next iteration
         upast = copy.copy(unext)
+  else:
+      return upast
 
   return unext
 
@@ -58,10 +60,10 @@ def serial_coarse_with_correction(deriv, init, corrections, dt, tmin, tmax, numS
   unext[0] = init
 
   # call forward euler per step and store in vector
-  for ind in xrange(numSteps):
+  for ind in xrange(1,numSteps):
     # get a coarse answer
     # TODO is this upast[ind] off by one?
-    unext[ind] = forward_euler_step(deriv, times[ind], unext[ind], dt)
+    unext[ind] = forward_euler_step(deriv, times[ind-1], unext[ind-1], dt)
 
     # update unext with corrections computed with upast
     unext[ind+1] = unext[ind] + fixed_corrections[ind]
@@ -98,7 +100,6 @@ def parallel_corrections(deriv, upast, tmin, tmax, numSteps, qualityFactor, comm
   local_differences = np.empty((0, upast.shape[1]))
   # in a loop, compute fine and then corrections one piece at a time that processor is responsible for, serially
   while (start < end):
-    print("Rank: %d, start: %d" %(rank, start))
     fineResult = forward_euler(deriv, upast[start], times[start], times[start+1], fineNumSteps)
 
     # get difference for the last time step of the fine result and compare to coarse result answer for same time
@@ -139,6 +140,9 @@ if __name__ == '__main__':
   tmin = 0.
   tmax = 1.
   numSteps = 100
+  k = 1
+  qualityFactor = 100
+  fineNumSteps = qualityFactor * numSteps
 
   init = np.array([1.])
   ###
@@ -149,7 +153,8 @@ if __name__ == '__main__':
 
   # Use parareal algorithm
   # we only want column 1 so numpy doesn't complain
-  p_result = startParareal(deriv, init, 2, tmin, tmax, numSteps, comm)
+  p_result = startParareal(deriv, init, k, tmin, tmax, numSteps, comm,
+          qualityFactor = 100)
 
   # Communciation barrier and ending time with MPI timing function
   comm.barrier()
@@ -163,15 +168,16 @@ if __name__ == '__main__':
     s_start = time.time()
 
     # we only want column 1 so numpy doesn't complain
[…]

### 2014-12-09 pararealAlgo.py
reformatted parareal file, added arguments and an additional example

@@ -4,7 +4,8 @@
 from mpi4py import MPI
 import numpy as np
 import time
-import copy
+import argparse
+import itertools
 
 from serial_euler import forward_euler
 from serial_euler import forward_euler_step
@@ -12,185 +13,235 @@ from serial_euler import forward_euler_step
 # Parareal begin
 def startParareal(deriv, init, k, tmin, tmax, numSteps, comm, 
         qualityFactor = 100):
-  '''Starts running k iterations of the parareal algorithm.
-  In each iteration, the program runs the coarse method in serial, then fine corrections in parallel
-  The result is only gathered to the root process at every iteration'''
+    '''Starts running k iterations of the parareal algorithm.  In each
+    iteration, the program runs the coarse method in serial, then fine
+    corrections in parallel The result is only gathered to the root process at
+    every iteration'''
 
-  rank = comm.Get_rank()
-  size = comm.Get_size()
+    rank = comm.Get_rank()
+    size = comm.Get_size()
 
-  dt = (tmax-tmin) / (numSteps - 1)
+    dt = (tmax-tmin) / (numSteps - 1)
 
-  # all coarse
-  upast = forward_euler(deriv, init, tmin, tmax, numSteps)
-  unext = None
+    # all coarse
+    upast = forward_euler(deriv, init, tmin, tmax, numSteps)
+    unext = None
 
-  # if further iterations are required
-  if k > 1:
-    for iteration in xrange(k-1):
+    # if further iterations are required
+    if k > 1:
+        for iteration in xrange(k-1):
 
-      # pass in a coarse answer to refine and correct, bectorized
-      corrections = parallel_corrections(deriv, upast, tmin, tmax, numSteps, qualityFactor, comm, p_root=0)
+            # pass in a coarse answer to refine and correct, vectorized
+            corrections = parallel_corrections(deriv, upast, tmin, tmax,
+                    numSteps, qualityFactor, comm, p_root=0)
 
-      if rank == 0:
-        # get a coarse answer for the next iteration, vectorized
-        unext = serial_coarse_with_correction(deriv, init, corrections, dt, tmin, tmax, numSteps)
+            if rank == 0:
+                # get a coarse answer for the next iteration, vectorized
+                unext = serial_coarse_with_correction(deriv, init, corrections,
+                        dt, tmin, tmax, numSteps)
 
-        # update unext for next iteration
[…]

### 2014-12-09 pararealAlgo.py
reformatted parareal file, added arguments and an additional example

@@ -224,6 +224,11 @@ if __name__ == '__main__':
 
             print("\n\nExample %d:" %iters)
             iters += 1
+            # print parameters:
+            print("Parameters:")
+            print("numSteps:        %d", %numSteps)
+            print("Quality Factor:  %d", %qualityFactor)
+            print("Corrections:     %d", %k)
             # print results if debug mode on
             if args.debug:
                 print("Serial Result   = " + str(s_result))

### 2014-12-09 pararealAlgo.py
more printing

@@ -222,18 +222,20 @@ if __name__ == '__main__':
             e_result_p = exact(coarse_times)
             e_result_s = exact(fine_times)
 
-            print("\n\nExample %d:" %iters)
+            print("\n\nExample %d\n" %iters)
             iters += 1
             # print parameters:
             print("Parameters:")
-            print("numSteps:        %d", %numSteps)
-            print("Quality Factor:  %d", %qualityFactor)
-            print("Corrections:     %d", %k)
+            print("numSteps:        %d" %numSteps)
+            print("Quality Factor:  %d" %qualityFactor)
+            print("Corrections:     %d" %k)
+            print("")
             # print results if debug mode on
             if args.debug:
-                print("Serial Result   = " + str(s_result))
+                print("Serial Result   = \n" + str(s_result))
                 print("Parallel Result = \n" + str(p_result))
-                print("Exact           = " + str(e_result))
+                print("Exact           = \n" + str(e_result))
+                print("")
 
             # compute errors
             p_error = np.abs(p_result - e_result_p)# / abs(e_result)

### 2014-12-09 pararealAlgo.py
added printing to help make csv files

@@ -224,6 +224,11 @@ if __name__ == '__main__':
 
             print("\n\nExample %d\n" %iters)
             iters += 1
+
+            # compute errors
+            p_error = np.abs(p_result - e_result_p)# / abs(e_result)
+            s_error = np.abs(s_result - e_result_s)# / abs(e_result)
+
             # print parameters:
             print("Parameters:")
             print("numSteps:        %d" %numSteps)
@@ -237,10 +242,6 @@ if __name__ == '__main__':
                 print("Exact           = \n" + str(e_result))
                 print("")
 
-            # compute errors
-            p_error = np.abs(p_result - e_result_p)# / abs(e_result)
-            s_error = np.abs(s_result - e_result_s)# / abs(e_result)
-
             # print errors
             #print("Parallel Error = \n" + str(p_error))
             #print("Serial Error = \n" + str(s_error))
@@ -252,3 +253,10 @@ if __name__ == '__main__':
             # print timers
             print("Serial Time:   %f secs" % (s_stop - s_start))
             print("Parallel Time: %f secs" % (p_stop - p_start))
+
+            print("numSteps,Quality Factor,Corrections,Serial Max Abs Error,Parallel Max Abs Error,Serial Mean Abs Error,Parallel Mean Abs Error, Serial Time, Parallel Time")
+
+            print("%d, %d, %d, %e, %e, %e, %e, %f, %f" %(numSteps,
+                qualityFactor, k, np.max(s_error), np.max(p_error),
+                np.mean(s_error), np.mean(p_error), (s_stop - s_start), (p_stop
+                    - p_start)))

### 2014-12-09 data/runscript_16procs.sh
added data

@@ -1,0 +1,31 @@
+#!/bin/bash
+#SBATCH --ntasks 16                #Number of processes
+#SBATCH --nodes 1                  #Number of nodes
+#SBATCH -t 180                 #Runtime in minutes
+#SBATCH -p general            #Partition to submit to
+
+#SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
+#SBATCH -o algo_16procs.out    #File to which standard out will be written
+#SBATCH -e algo_16procs.err    #File to which standard err will be written
+
+# load modules
+module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
+module load centos6/numpy-1.7.1_python-2.7.3
+
+cd ~/parareal/data
+
+# make sure you change the -n argument for mpirun and the output files
+
+# change the quality factor
+k=2
+for q in 10 50 100
+do
+    mpirun -n 16 python ../pararealAlgo.py --nsteps 100 --quality-factor $q --correction-level $k
+done
+
+# change the number of corrections
+q=100
+for k in 1 2 5 10 50 100
+do
+    mpirun -n 64 python ../pararealAlgo.py --nsteps 100 --quality-factor $q --correction-level $k
+done

### 2014-12-09 data/runscript_2procs.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+#SBATCH --ntasks 2                #Number of processes
+#SBATCH --nodes 1                  #Number of nodes
+#SBATCH -t 60                 #Runtime in minutes
+#SBATCH -p general            #Partition to submit to
+
+#SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
+#SBATCH -o algo_2procs.out    #File to which standard out will be written
+#SBATCH -e algo_2procs.err    #File to which standard err will be written
+
+# load modules
+module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
+module load centos6/numpy-1.7.1_python-2.7.3
+
+cd ~/parareal/data
+
+# change the number of corrections
+for k in 2
+do
+    mpirun -n 2 python ../pararealAlgo.py --nsteps 100 --quality-factor 100 --correction-level $k
+done

### 2014-12-09 data/runscript_32procs.sh

@@ -1,0 +1,31 @@
+#!/bin/bash
+#SBATCH --ntasks 32                #Number of processes
+#SBATCH --nodes 1                  #Number of nodes
+#SBATCH -t 300                 #Runtime in minutes
+#SBATCH -p general            #Partition to submit to
+
+#SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
+#SBATCH -o algo_32procs.out    #File to which standard out will be written
+#SBATCH -e algo_32procs.err    #File to which standard err will be written
+
+# load modules
+module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
+module load centos6/numpy-1.7.1_python-2.7.3
+
+cd ~/parareal/data
+
+# make sure you change the -n argument for mpirun
+
+# change the quality factor
+k=2
+for q in 10 50
+do
+    mpirun -n 32 python ../pararealAlgo.py --nsteps 100 --quality-factor $q --correction-level $k
+done
+
+# change the number of corrections
+q=100
+for k in 1 2 5 10 50 100
+do
+    mpirun -n 32 python ../pararealAlgo.py --nsteps 100 --quality-factor $q --correction-level $k
+done

### 2014-12-09 data/runscript_4procs.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+#SBATCH --ntasks 4                #Number of processes
+#SBATCH --nodes 1                  #Number of nodes
+#SBATCH -t 180                 #Runtime in minutes
+#SBATCH -p general            #Partition to submit to
+
+#SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
+#SBATCH -o algo_4procs.out    #File to which standard out will be written
+#SBATCH -e algo_4procs.err    #File to which standard err will be written
+
+# load modules
+module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
+module load centos6/numpy-1.7.1_python-2.7.3
+
+cd ~/parareal/data
+
+# make sure you change the -n argument for mpirun
+
+# change the quality factor
+k=2
+for q in 10 50
+do
+    mpirun -n 4 python ../pararealAlgo.py --nsteps 100 --quality-factor $q --correction-level $k
+done

### 2014-12-09 data/runscript_8procs.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+#SBATCH --ntasks 8                #Number of processes
+#SBATCH --nodes 1                  #Number of nodes
+#SBATCH -t 180                 #Runtime in minutes
+#SBATCH -p general            #Partition to submit to
+
+#SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
+#SBATCH -o algo_8procs.out    #File to which standard out will be written
+#SBATCH -e algo_8procs.err    #File to which standard err will be written
+
+# load modules
+module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
+module load centos6/numpy-1.7.1_python-2.7.3
+
+cd ~/parareal/data
+
+# make sure you change the -n argument for mpirun
+
+# change the quality factor
+k=2
+for q in 100 
+do
+    mpirun -n 8 python ../pararealAlgo.py --nsteps 100 --quality-factor $q --correction-level $k
+done

### 2014-12-09 paper.tex

@@
     \begin{abstract}
-    We implement the parareal algorithm to solve various ODEs and compare in
-    terms of accuracy, speedup and efficiency. We explore the effect of using
-    differnt solution operators (a coarse and a fine as required by the method)
-    The majority of our code is written in python using mpi4py to parallelize.
-    To compare to more performance-based languages, we compare certain codes to
-    an implementation in C++.
+    We implement the parareal algorithm to solve various ODEs and compare in terms of accuracy, speedup and efficiency.
+We explore the effect of using differnt solution operators (a coarse and a fine as required by the method).
+Our code is written in python using mpi4py to parallelize.
+We observe slight speedups in our small scale tests up to 64 processors on Harvard's Odyssey cluster.
+We analyze possible reasons as to why our implementation may not be ideal and some tradeoffs that are taken in the parareal implementation.
+Further optimizations are proposed and rationalized as next steps including ports of to C++ and BLAS libraries which is a more performance-oriented, lower-level language than Python.
+Data is shown for solving for variuos complexities of ODEs from the basic exponential to modelling sound wave propagation as we did in Homework 4.
     \end{abstract}
@@
 
-\subsection{Numerical Methods and Parallelization}
-Solving systems of differential equations can be a computationally expensive
-task. The error of most algorithms scales on the stepsize of the discretization
-of time. However, stepsize in time is also proportional to the computation
-required. The parareal algorithm, developed by YADA YADA. An alternative would
-be to parallelize by space -
+\subsection{Numerical Methods and \\ Parallelization}
+Solving systems of differential equations can be a computationally expensive task.
+The error of most algorithms scales on the stepsize of the discretization of time.
+However, stepsize in time is also proportional to the computation required.
+It would be nice to allow for strong scaling where a problem can be solved in a reasonable time for very small time steps.
+
+Another obstacle to trying to paralleliize numerical methods is that the methods are inherently serial in time - in that evaluation of the next time $n+1$ depends on the previous values, say those at time $n-1$ and $n$.
+This setup does not allow for parallelism.
+
+ An alternative paradigm is a parallelism by space, where the result vector is divided up to different processors.  In this way, the
+
+In general, the
+
+Due to the more theoretical nature of the parareal method, we look to describe its stability and error numerically.
 
@@
 
+FROM THE SLIDES
+
+TODO BKSIM? ASHI carry?
+
 \subsection{Error of the Parareal}
+
+FROM THE SLIDES? + Adjust?
+TODO BKSIM? ASHI carry?
+
+Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.
+However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.
 
@@
 
-The general parareal method, as described by YADA YADA can be SEEN BELOW YAR
+The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is a general algorithm that allows parallelization in time.
+It does so by using a cheaper (lower order, lesser resolution) approximation first, and then makes corrections in parallel.
+The entire process is then iterative and tuned for a given amount of iterations.
+Therefore, the algorithm is k repetitions of a serial coarse method updated by a finer method run in parallel for subsections.
 
@@
 
+\subsection{Visualization of Diagram}
+bksim TODO
+
+\subsection{Analysis of Speedup and Efficiency}
+
+ashi TODO from our computation on the paper
+
 \subsection{Explored Tests}
-We wished to explore certain tests
+We wished to explore certain tests that would span the space of possible parameters to the parareal as well as possible applications in different ODE systems.  Our baseline was looking to compare
 
 \section{Results}
-We compared various runs of
+
+TODO CONSIDER IF WE WANT TO MOVE SOME OF THE ANALYSIS TO THE DISCUSSION SECTION AND FRAME THE DISCUSSION WITH THE SAME SUBSECTIONS
 
@@
 
+We looked
+
+We see that the speedups are
+
+Note that increasing the iterations
+
+ashi TODO add figures and check my logic/bs to match it to the data
+
 \subsection{Comparison to Serial with $g_{fine}$ as \\Higher Order Methods}
 
+TODO team?
+
 \subsection{Comparison to a Forward Euler \\Parallelized by Space}
+
+For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute.
+
+bksim TODO
 
@@
 
+TODO
+
+\subsection{Strong and Weak Scaling Analysis}
+
+ashi TODO by taking some numbers and following %https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance
+
 \section{Discussion}
+
+\subsection{Demonstrating Scaling}
+
+As we can see, we do see the effects of strong scaling
+
+TODO analyze
+
+\subsection{Possible Optimizations}
+
+TODO Wesley
 
@@
 
+We see that (HOPE IT WORKS WELL OR ELSE...)
+Most of the further work would be in optimization.
+We are trying to measure performance in python, which is not the ideal benchmarking language for efficiency and speedups.
+A port over to C++ using the MPI libraries in C++ rather than the mpi4py libraries in python would be ideal.
+Other optimizations mentioned above coud also be implemented.
+To further explore the parareal algorithm would involve many test cases and to see how the parareal algorithm compares to other approaches - since the actual time and iterations necessary depend on the system of ODEs to solve.
+
+On the other hand, paralleism by space already requires a very problem-specific design.
+The strong scaling efficiency of parallelism by space is high and is conceptually embarrisingly parallel and easier to create.
+However, the limitations are in the requirement for the code to be designed for the problem - how to divide by space so as to allow for the maximum number of even divisions.
+The lack of generalism is not as beautiful as the parareal algorithm but may lend itself to faster speedups.
+
+All in all, we have implemented and begun to explore some techniques of looking for strong and weak scaling efficiencies for solving ODEs.
+The parareal algorithm is beautiful in its theoretical advantages - in terms of stability, error and efficiency.
+However, the method, as a very generalized method, requires deeper analysis for the specific problem.
+The main variable will be in the difference in the coarse and fine methods (be it lower and higher order or a higher and lower resolution for the same method).
+The tradeoffs that must be computed and optimized for would be problem specific.
+
 \section{References}
 
+Bibtex Citations for the Slides, and for %https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance
+
 \end{document}

### 2014-12-09 poster/main.tex
updated algo to hopefully use a bit less time; updated poster

@@
 
-Input: Temporal discretization $t^n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$
-Input: Coarse scheme $g_{\Delta t}$
-Input: Finer scheme $g_{\textnormal{fine}}$
-Compute $u_1^{n+1} = g_{\Delta t}(t^n, u^n)$
-Compute the corrections $\delta g^n(u_1^n) = g_{\textnormal{fine}}(t^n,
-u^n) - g_{\Delta t}(t^n, u^n)$ in parallel
-Add the prediction and correction terms as $u_2^{n+1} = g_{\Delta
-t}(t^n, u_2^n) + \delta g^n(u_1^n)$
-Repeat steps 2 and 3, incrementing the iteration label and using $u_{k+1}^0
-= u^0$ as the initial condition
+\begin{itemize}
+    \item \textbf{Input}: Temporal discretization $t_n = t_0 + n \Delta t, \, n =
+        1,2,\ldots,N$
+    \item \textbf{Input}: Coarse scheme $g_{\Delta t}$
+    \item \textbf{Input}: Finer scheme $g_{\textnormal{fine}}$
+\end{itemize}
+
+\begin{enumerate}
+    \item Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$
+    \item Compute the corrections $\delta g_n(u_1_n) =
+        g_{\textnormal{fine}}(t_n, u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel
+    \item Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta
+        t}(t_n, u^2_n) + \delta g_n(u^1_n)$
+    \item Repeat steps 2 and 3, incrementing the iteration label and using
+        $u^{k+1}_0 = u^1_0$ as the initial condition
+\end{enumerate}
 \end{block}

### 2014-12-09 paper.tex
trying to get more data; minor changes in paper

--- 2014-12-08 paper.tex
@@
 Harvard University, Applied Math 205 Final Project}
-\date{December 15, 2014}
+\date{December 10, 2014}
 
@@
 
-\subsection{Comparison to Serial with $g_{fine}$ as \\a Forward Euler with
+\subsection{Comparison to Serial with $g_{fine}$ as a Forward Euler with
 Smaller $\Delta t$}
 
-\subsection{Comparison to Serial with $g_{fine}$ as \\Higher Order Methods}
+\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}
 
-\subsection{Comparison to a Forward Euler \\Parallelized by Space}
+\subsection{Comparison to a Forward Euler Parallelized by Space}
 
 \subsection{Comparison to C++ Implementations}
-	The numpy environment combined with the required structure of MPI required certain type/data structure conversions which will take time.not not cunot cu
+    The numpy environment combined with the required structure of MPI required
+    certain type/data structure conversions which will take time.not not cunot
+    cu
 

### 2014-12-09 paper.tex

--- 2014-12-09 paper.tex
@@
     \begin{abstract}
-    We implement the parareal algorithm to solve various ODEs and compare in terms of accuracy, speedup and efficiency.
-We explore the effect of using differnt solution operators (a coarse and a fine as required by the method).
-Our code is written in python using mpi4py to parallelize.
-We observe slight speedups in our small scale tests up to 64 processors on Harvard's Odyssey cluster.
-We analyze possible reasons as to why our implementation may not be ideal and some tradeoffs that are taken in the parareal implementation.
-Further optimizations are proposed and rationalized as next steps including ports of to C++ and BLAS libraries which is a more performance-oriented, lower-level language than Python.
-Data is shown for solving for variuos complexities of ODEs from the basic exponential to modelling sound wave propagation as we did in Homework 4.
+    We implement the parareal algorithm to solve various ODEs and compare in
+    terms of accuracy, speedup and efficiency. We explore the effect of using
+    differnt solution operators (a coarse and a fine as required by the method).
+    Our code is written in python using mpi4py to parallelize. We observe
+    slight speedups in our small scale tests up to 64 processors on Harvard's
+    Odyssey cluster. We analyze possible reasons as to why our implementation
+    may not be ideal and some tradeoffs that are taken in the parareal
+    implementation. Further optimizations are proposed and rationalized as next
+    steps including ports of to C++ and BLAS libraries which is a more
+    performance-oriented, lower-level language than Python. Data is shown for
+    solving for variuos complexities of ODEs from the basic exponential to
+    modelling sound wave propagation as we did in Homework 4.
     \end{abstract}
@@
 \subsection{Numerical Methods and \\ Parallelization}
-Solving systems of differential equations can be a computationally expensive task.
-The error of most algorithms scales on the stepsize of the discretization of time.
-However, stepsize in time is also proportional to the computation required.
-It would be nice to allow for strong scaling where a problem can be solved in a reasonable time for very small time steps.
+Solving systems of differential equations can be a computationally expensive
+task. The error of most algorithms scales on the stepsize of the discretization
+of time. However, stepsize in time is also proportional to the computation
+required. It would be nice to allow for strong scaling where a problem can be
+solved in a reasonable time for very small time steps.
 
-Another obstacle to trying to paralleliize numerical methods is that the methods are inherently serial in time - in that evaluation of the next time $n+1$ depends on the previous values, say those at time $n-1$ and $n$.
-This setup does not allow for parallelism.
+Another obstacle to trying to paralleliize numerical methods is that the methods
+are inherently serial in time---in that evaluation of the next time $n+1$
+depends on the previous values, say those at time $n-1$ and $n$. This setup
+does not allow for parallelism.
 
- An alternative paradigm is a parallelism by space, where the result vector is divided up to different processors.  In this way, the
+ An alternative paradigm is a parallelism by space, where the result vector is
+ divided up to different processors.  In this way, the
 
@@
 
-Due to the more theoretical nature of the parareal method, we look to describe its stability and error numerically.
+Due to the more theoretical nature of the parareal method, we look to describe
+its stability and error numerically.
 
@@
 
-Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.
-However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.
+Show that this parareal method will approach, with large enough $k$ (enough
+iterations) to approach the error of the fine method. However, with too large
+of a k and a lower quality factor (ratio of fine to coarse), the time for the
+parareal could potentially take longer than the direct serial computation.
 
@@
 
-The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is a general algorithm that allows parallelization in time.
-It does so by using a cheaper (lower order, lesser resolution) approximation first, and then makes corrections in parallel.
-The entire process is then iterative and tuned for a given amount of iterations.
-Therefore, the algorithm is k repetitions of a serial coarse method updated by a finer method run in parallel for subsections.
+The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
+a general algorithm that allows parallelization in time.  It does so by using a
+cheaper (lower order, lesser resolution) approximation first, and then makes
+corrections in parallel. The entire process is then iterative and tuned for a
+given amount of iterations. Therefore, the algorithm is $k$ repetitions of a
+serial coarse method updated by a finer method run in parallel for subsections.
 
@@
     \KwIn{Finer	 scheme $g_{\textnormal{fine}}$}
-    Compute $u_1^{n+1} = g_{\Delta t}(t^n, u^n)$\;
-    Compute the corrections $\delta g^n(u_1^n) = g_{\textnormal{fine}}(t^n,
-    u^n) - g_{\Delta t}(t^n, u^n)$ in parallel\;
-    Add the prediction and correction terms as $u_2^{n+1} = g_{\Delta
-    t}(t^n, u_2^n) + \delta g^n(u_1^n)$\;
-    Repeat steps 2 and 3, incrementing the iteration label and using $u_{k+1}^0
-    = u^0$ as the initial condition\
+    Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
+    Compute the corrections $\delta g_n(u^1_n) = g_{\textnormal{fine}}(t_n,
+    u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel\;
+    Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta t}(t_n,
+    u^2_n) + \delta g_n(u^1_n)$\;
+    Repeat steps 2 and 3, incrementing the iteration label and using $u^{k+1}_0
+    = u^1_0$ as the initial condition\;
  \caption{Parareal}
@@
 
-\subsection{Visualization of Diagram}
+\subsection{Visualization Diagram}
 bksim TODO
@@
 \subsection{Explored Tests}
-We wished to explore certain tests that would span the space of possible parameters to the parareal as well as possible applications in different ODE systems.  Our baseline was looking to compare
+
+We wished to explore certain tests that would span the space of possible
+parameters to the parareal as well as possible applications in different ODE
+systems. Our baseline was looking to compare
 
@@
 
-\subsection{Comparison to Serial with $g_{fine}$ as \\a Forward Euler with
+\subsection{Comparison to Serial with $g_{fine}$ as a Forward Euler with
 Smaller $\Delta t$}
+
+\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}
 
@@
 
-\subsection{Comparison to Serial with $g_{fine}$ as \\Higher Order Methods}
+\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}
 
@@
 
-\subsection{Comparison to a Forward Euler \\Parallelized by Space}
+\subsection{Comparison to a Forward Euler Parallelized by Space}
 
@@
 \subsection{Comparison to C++ Implementations}
-	The numpy environment combined with the required structure of MPI required certain type/data structure conversions which will take time.not not cunot cu
+
+The numpy environment combined with the required structure of MPI required
+certain type/data structure conversions which will take time.not not cunot
+cu
 
@@
 
-ashi TODO by taking some numbers and following %https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance
+ashi TODO by taking some numbers and following
+\url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
 
@@
 
-We see that (HOPE IT WORKS WELL OR ELSE...)
-Most of the further work would be in optimization.
-We are trying to measure performance in python, which is not the ideal benchmarking language for efficiency and speedups.
-A port over to C++ using the MPI libraries in C++ rather than the mpi4py libraries in python would be ideal.
-Other optimizations mentioned above coud also be implemented.
-To further explore the parareal algorithm would involve many test cases and to see how the parareal algorithm compares to other approaches - since the actual time and iterations necessary depend on the system of ODEs to solve.
+We see that (HOPE IT WORKS WELL OR ELSE...) Most of the further work would be in
+optimization. We are trying to measure performance in Python, which is not the
+ideal benchmarking language for efficiency and speedups. A port over to C++
+using the MPI libraries in C++ rather than the mpi4py libraries in python would
+be ideal. Other optimizations mentioned above could also be implemented. To
+further explore the parareal algorithm would involve many test cases and to see
+how the parareal algorithm compares to other approaches---since the actual time
+and iterations necessary depend on the system of ODEs to solve.
 
-On the other hand, paralleism by space already requires a very problem-specific design.
-The strong scaling efficiency of parallelism by space is high and is conceptually embarrisingly parallel and easier to create.
-However, the limitations are in the requirement for the code to be designed for the problem - how to divide by space so as to allow for the maximum number of even divisions.
-The lack of generalism is not as beautiful as the parareal algorithm but may lend itself to faster speedups.
+On the other hand, paralleism by space already requires a very problem-specific
+design. The strong scaling efficiency of parallelism by space is high and is
+conceptually embarrisingly parallel and easier to create. However, the
+limitations are in the requirement for the code to be designed for the
+problem---how to divide by space so as to allow for the maximum number of even
+divisions. The lack of generalism is not as beautiful as the parareal algorithm
+but may lend itself to faster speedups.
 
-All in all, we have implemented and begun to explore some techniques of looking for strong and weak scaling efficiencies for solving ODEs.
-The parareal algorithm is beautiful in its theoretical advantages - in terms of stability, error and efficiency.
-However, the method, as a very generalized method, requires deeper analysis for the specific problem.
-The main variable will be in the difference in the coarse and fine methods (be it lower and higher order or a higher and lower resolution for the same method).
-The tradeoffs that must be computed and optimized for would be problem specific.
+All in all, we have implemented and begun to explore some techniques of looking
+for strong and weak scaling efficiencies for solving ODEs. The parareal
+algorithm is beautiful in its theoretical advantages---in terms of stability,
+error and efficiency. However, the method, as a very generalized method,
+requires deeper analysis for the specific problem. The main variable will be in
+the difference in the coarse and fine methods (be it lower and higher order or a
+higher and lower resolution for the same method). The tradeoffs that must be
+computed and optimized for would be problem specific.
 
@@
 
-Bibtex Citations for the Slides, and for %https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance
+Bibtex Citations for the Slides, and for \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
 

### 2014-12-09 paper.tex
another paper update, more to come

--- 2014-12-09 paper.tex
@@
 This setup does not allow for parallelism.
+The only way to incorporate parallelism is to create schemes that have parallelizable components - such as an update or a refinement to a serial portion.
+As will be discussed and analyzed in the metodology section, one such scheme would be the parareal algorithm.
+For an overview, the parareal algorithm first does a coarse (low-order and fast) approximation, then seeks to iteratively correct using smaller, more refined numerical methods done in parallel from approximated starting points approximated by the coarse solution.
 
- An alternative paradigm is a parallelism by space, where the result vector is divided up to different processors.  In this way, the
+An alternative paradigm to applying parallelism to numerical methods is the parallelism by space paradigm, where the result vector space is divided up to different regions and sent to different responsible processors.
+In this way, each processor sees a similar but smaller-scale problem - weak scaling.
+This method is expected to outperform the parallel in time paradigm simply because the task is somewhat embarassingly parallel, being able to be naturally divided into smaller but identical in method problems.
+There is no serial component in this algorithm and should be exhibit both strong and weak scaling.
 
-In general, the
-
-Due to the more theoretical nature of the parareal method, we look to describe its stability and error numerically.
-
-\subsection{Stability of the Parareal}
-
-FROM THE SLIDES
-
-TODO BKSIM? ASHI carry?
-
-\subsection{Error of the Parareal}
-
-FROM THE SLIDES? + Adjust?
-TODO BKSIM? ASHI carry?
-
-Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.
-However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.
+In general, the division in space paradigm is more intuitive to parallelize and will often lead to greater speedups because there is no serial component.
+Yet it is worth noting that knowing how to divide the result vector requires knowledge of problem-specific details.
+Thus, this approach lacks the generalisability and theoretical derivaions that schemes that parallelize in time can offer.
 
@@
 
-\subsection{Visualization of Diagram}
+\subsection{Visualization of Scheme}
 bksim TODO
+
+\subsection{Stability of the Parareal}
+
+FROM THE SLIDES
+
+TODO BKSIM? ASHI carry?
+
+\subsection{Error of the Parareal}
+
+FROM THE SLIDES? + Adjust?
+TODO BKSIM? ASHI carry?
+
+Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.
+However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.
 
@@
 
-ashi TODO from our computation on the paper
+ashi TODO ashia from wesley's handwritten analysis
 
@@
 
-\section{Results}
-
-TODO CONSIDER IF WE WANT TO MOVE SOME OF THE ANALYSIS TO THE DISCUSSION SECTION AND FRAME THE DISCUSSION WITH THE SAME SUBSECTIONS
+\section{Results and Discussion}
 
@@
 
-We looked
+We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using
 
-We see that the speedups are
+Varying Iterations:
 
-Note that increasing the iterations
+Varying Quality Factor:
+We define the quality factor in this case for how many times smaller the Euler method used for $g_{fine}$ is compared to the Euler method used for $g_{\Delta t}$
+
+Changing the ODE
 
@@
 
-TODO team?
+TODO team
 
-\subsection{Comparison to a Forward Euler \\Parallelized by Space}
+IDEALLY WE SHOW THE SAME GRAPHS AS THE ABOVE WITH ALL THE VARIATIONS TESTED
+
+\subsection{Parallelism by Space Paradigm}
 
@@
 bksim TODO
+
+\subsection{Parallelism Paradigm Comparisons}
+
+TODO compare the computation times (and perhaps a slight change in accuracy of the results) between both parareal methods the best in the sections before hand
+
+TODO bsim and ashi show graphs and some times for the forward euler parareal versions - the fastest ones or both if we have obht implemented
 
@@
 
-TODO
+TODO team
 
@@
 ashi TODO by taking some numbers and following %https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance
-
-\section{Discussion}
 

### 2014-12-09 paper.tex

--- 2014-12-09 paper.tex
@@
 
-Another obstacle to trying to paralleliize numerical methods is that the methods
-are inherently serial in time---in that evaluation of the next time $n+1$
-depends on the previous values, say those at time $n-1$ and $n$. This setup
-does not allow for parallelism.
+Another obstacle to trying to paralleliize numerical methods is that the methods are inherently serial in time - in that evaluation of the next time $n+1$ depends on the previous values, say those at time $n-1$ and $n$.
+This setup does not allow for parallelism.
+The only way to incorporate parallelism is to create schemes that have parallelizable components - such as an update or a refinement to a serial portion.
+As will be discussed and analyzed in the metodology section, one such scheme would be the parareal algorithm.
+For an overview, the parareal algorithm first does a coarse (low-order and fast) approximation, then seeks to iteratively correct using smaller, more refined numerical methods done in parallel from approximated starting points approximated by the coarse solution.
 
- An alternative paradigm is a parallelism by space, where the result vector is
- divided up to different processors.  In this way, the
+An alternative paradigm to applying parallelism to numerical methods is the parallelism by space paradigm, where the result vector space is divided up to different regions and sent to different responsible processors.
+In this way, each processor sees a similar but smaller-scale problem - weak scaling.
+This method is expected to outperform the parallel in time paradigm simply because the task is somewhat embarassingly parallel, being able to be naturally divided into smaller but identical in method problems.
+There is no serial component in this algorithm and should be exhibit both strong and weak scaling.
 
-In general, the
-
-Due to the more theoretical nature of the parareal method, we look to describe
-its stability and error numerically.
-
-\subsection{Stability of the Parareal}
-
-FROM THE SLIDES
-
-TODO BKSIM? ASHI carry?
-
-\subsection{Error of the Parareal}
-
-FROM THE SLIDES? + Adjust?
-TODO BKSIM? ASHI carry?
-
-Show that this parareal method will approach, with large enough $k$ (enough
-iterations) to approach the error of the fine method. However, with too large
-of a k and a lower quality factor (ratio of fine to coarse), the time for the
-parareal could potentially take longer than the direct serial computation.
+In general, the division in space paradigm is more intuitive to parallelize and will often lead to greater speedups because there is no serial component.
+Yet it is worth noting that knowing how to divide the result vector requires knowledge of problem-specific details.
+Thus, this approach lacks the generalisability and theoretical derivaions that schemes that parallelize in time can offer.
 
@@
 
-\subsection{Visualization Diagram}
+\subsection{Visualization of Scheme}
+
 bksim TODO
+
+\subsection{Stability of the Parareal}
+
+FROM THE SLIDES
+
+TODO BKSIM? ASHI carry?
+
+\subsection{Error of the Parareal}
+
+FROM THE SLIDES? + Adjust?
+TODO BKSIM? ASHI carry?
+
+Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.
+However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.
 
@@
 
-\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}
+We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using GET PARAMETERES
 
-We looked
+Varying Iterations:
 
-We see that the speedups are
+Varying Quality Factor:
+We define the quality factor in this case for how many times smaller the Euler method used for $g_{fine}$ is compared to the Euler method used for $g_{\Delta t}$
 
-Note that increasing the iterations
+Changing the ODE
 
@@
 
-TODO team?
+TODO team
 
-\subsection{Comparison to a Forward Euler Parallelized by Space}
+IDEALLY WE SHOW THE SAME GRAPHS AS THE ABOVE WITH ALL THE VARIATIONS TESTED
+
+\subsection{Parallelism by Space Paradigm}
 
@@
 \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
-
-\section{Discussion}
 

### 2014-12-09 paper.tex
minor changes on paper
style modifications to paper

@@
-%        File: paper.tex
-%     Created: Sat Dec 06 04:00 PM 2014 E
+%ssssFile: paper.tex
+%ss Created: Sat Dec 06 04:00 PM 2014 E
 %
-\documentclass[letterpaper,twocolumn,11pt]{article}
+\documentclass[letterpaper,12pt]{article}
 % Change the header if you're going to change this.
@@
 \usepackage{amsmath} % Needed for math stuff
-%\usepackage{lastpage}      % Finds last page
+%\usepackage{lastpage}sss% Finds last page
 \usepackage{amssymb} % Needed for some math symbols
-\usepackage{graphicx}      % Needed for graphics
+\usepackage{graphicx}% Needed for graphics
 \usepackage[usenames,dvipsnames]{xcolor} % Needed for graphics and color
-%\usepackage{setspace}      % Needed to set single/double spacing
+%\usepackage{setspace}sss% Needed to set single/double spacing
 \usepackage{float} % Better placement of figures & tables
@@
 % Put your stuff here
-\twocolumn[
-    \begin{@twocolumnfalse}
+%\twocolumn[
+%\begin{@twocolumnfalse}
 \maketitle
@@
 Our code is written in python using mpi4py to parallelize. We observe
-    slight speedups in our small scale tests up to 64 processors on Harvard's
+light speedups in our small scale tests up to 64 processors on Harvard's
 Odyssey cluster. We analyze possible reasons as to why our implementation
@@
 \end{abstract}
-    \end{@twocolumnfalse}
-]
+%\end{@twocolumnfalse}
+%]
+
+\clearpage
+
+\tableofcontents
+
+\clearpage
+
 \section{Background}
 
-\subsection{Numerical Methods and \\ Parallelization}
+\subsection{Numerical Methods and Parallelization}
 Solving systems of differential equations can be a computationally expensive
@@
 
-Another obstacle to trying to paralleliize numerical methods is that the methods are inherently serial in time - in that evaluation of the next time $n+1$ depends on the previous values, say those at time $n-1$ and $n$.
-This setup does not allow for parallelism.
-The only way to incorporate parallelism is to create schemes that have parallelizable components - such as an update or a refinement to a serial portion.
-As will be discussed and analyzed in the metodology section, one such scheme would be the parareal algorithm.
-For an overview, the parareal algorithm first does a coarse (low-order and fast) approximation, then seeks to iteratively correct using smaller, more refined numerical methods done in parallel from approximated starting points approximated by the coarse solution.
+Another obstacle to trying to paralleliize numerical methods is that the methods
+are inherently serial in time---in that evaluation of the next time $n+1$
+depends on the previous values, say those at time $n-1$ and $n$. This setup
+does not allow for parallelism. The only way to incorporate parallelism is to
+create schemes that have parallelizable components---such as an update or a
+refinement to a serial portion. As will be discussed and analyzed in the
+metodology section, one such scheme would be the parareal algorithm.sFor an
+overview, the parareal algorithm first does a coarse (low-order and fast)
+approximation, then seeks to iteratively correct using smaller, more refined
+numerical methods done in parallel from approximated starting points
+approximated by the coarse solution.
 
-An alternative paradigm to applying parallelism to numerical methods is the parallelism by space paradigm, where the result vector space is divided up to different regions and sent to different responsible processors.
-In this way, each processor sees a similar but smaller-scale problem - weak scaling.
-This method is expected to outperform the parallel in time paradigm simply because the task is somewhat embarassingly parallel, being able to be naturally divided into smaller but identical in method problems.
-There is no serial component in this algorithm and should be exhibit both strong and weak scaling.
+An alternative paradigm to applying parallelism to numerical methods is the
+parallelism by space paradigm, where the result vector space is divided up to
+different regions and sent to different responsible processors. In this way,
+each processor sees a similar but smaller-scale problem---weak scaling. This
+method is expected to outperform the parallel in time paradigm simply because
+the task is somewhat embarassingly parallel, being able to be naturally divided
+into smaller but identical in method problems. There is no serial component in
+this algorithm and should be exhibit both strong and weak scaling.
 
-In general, the division in space paradigm is more intuitive to parallelize and will often lead to greater speedups because there is no serial component.
-Yet it is worth noting that knowing how to divide the result vector requires knowledge of problem-specific details.
-Thus, this approach lacks the generalisability and theoretical derivaions that schemes that parallelize in time can offer.
+In general, the division in space paradigm is more intuitive to parallelize and
+will often lead to greater speedups because there is no serial component. Yet
+it is worth noting that knowing how to divide the result vector requires
+knowledge of problem-specific details. Thus, this approach lacks the
+general stability and theoretical derivations that schemes that parallelize in
+time can offer.
 
@@
 The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
-a general algorithm that allows parallelization in time.  It does so by using a
+a general algorithm that allows parallelization in time.sIt does so by using a
 cheaper (lower order, lesser resolution) approximation first, and then makes
@@
 \begin{algorithm}[t]
-    \KwIn{Temporal discretization $t^n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
+    \KwIn{Temporal discretization $t_n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
     \KwIn{Coarse scheme $g_{\Delta t}$}
-    \KwIn{Finer	 scheme $g_{\textnormal{fine}}$}
+    \KwIn{Finersscheme $g_{\textnormal{fine}}$}
     Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
@@
 
-Show that this parareal method will approach, with large enough k (enough iterations) to approach the error of the fine method.
-However, with too large of a k and a lower quality factor (ratio of fine to coarse), the time for the parareal could potentially take longer than the direct serial computation.
+Show that this parareal method will approach, with large enough $k$ (enough
+iterations) to approach the error of the fine method. However, with too large of
+a $k$ and a lower quality factor (ratio of fine to coarse), the time for the
+parareal could potentially take longer than the direct serial computation.
 
@@
 
-We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using GET PARAMETERES
+We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using GET PARAMETERS
 
@@
 
-For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute.
+For the speaker problem, we designed a problem-specific method to parallelize in
+space and divide Pierce Hall into subsections such that each processor would
+receive a division of the building to compute.
 
@@
 
-TODO compare the computation times (and perhaps a slight change in accuracy of the results) between both parareal methods the best in the sections before hand
+TODO compare the computation times (and perhaps a slight change in accuracy of
+the results) between both parareal methods the best in the sections before hand
 
-TODO bsim and ashi show graphs and some times for the forward euler parareal versions - the fastest ones or both if we have obht implemented
+TODO bsim and ashi show graphs and some times for the forward Euler parareal
+versions---the fastest ones or both if we have both implemented
 
@@
 The numpy environment combined with the required structure of MPI required
-certain type/data structure conversions which will take time.not not cunot
-cu
+certain type/data structure conversions which will take time.not not cunot cu
 
@@
 
-On the other hand, paralleism by space already requires a very problem-specific
+On the other hand, parallelism by space already requires a very problem-specific
 design. The strong scaling efficiency of parallelism by space is high and is
-conceptually embarrisingly parallel and easier to create. However, the
+conceptually embarrassingly parallel and easier to create. However, the
 limitations are in the requirement for the code to be designed for the
 problem---how to divide by space so as to allow for the maximum number of even
-divisions. The lack of generalism is not as beautiful as the parareal algorithm
+divisions. The lack of generality is not as beautiful as the parareal algorithm
 but may lend itself to faster speedups.

### 2014-12-09 paper.tex
changed poster and writeup

@@
 terms of accuracy, speedup and efficiency. We explore the effect of using
-differnt solution operators (a coarse and a fine as required by the method).
+different solution operators (a coarse and a fine as required by the method).
 Our code is written in python using mpi4py to parallelize. We observe
@@
 performance-oriented, lower-level language than Python. Data is shown for
-solving for variuos complexities of ODEs from the basic exponential to
-modelling sound wave propagation as we did in Homework 4.
+solving for various complexities of ODEs from the basic exponential to
+modeling sound wave propagation as we did in Homework 4.
 \end{abstract}

### 2014-12-09 poster/main.tex
changed poster and writeup

@@
 \usepackage{graphicx}  % Required for including images
+\usepackage{hyperref}
+\usepackage{natbib}
 
 \usepackage{booktabs} % Top and bottom rules for tables
+\usepackage{microtype}
 \usepackage[linesnumbered,vlined,ruled]{algorithm2e} % algorithm
@@
 
-We implement the parareal algorithm to solve various ODEs and compare in terms
-of accuracy, speedup and efficiency. We explore the effect of using differnt
-solution operators (a coarse and a fine as required by the method) The majority
-of our code is written in python using mpi4py to parallelize.  To compare to
-more performance-based languages, we compare certain codes to an implementation
-in C++.
+We implement the parareal algorithm to solve various ODEs and compare in
+terms of accuracy, speedup and efficiency. We explore the effect of using
+different solution operators (a coarse and a fine as required by the method).
+Our code is written in python using mpi4py to parallelize. We observe
+light speedups in our small scale tests up to 64 processors on Harvard's
+Odyssey cluster. We analyze possible reasons as to why our implementation
+may not be ideal and some tradeoffs that are taken in the parareal
+implementation. Further optimizations are proposed and rationalized as next
+steps including ports of to C++ and BLAS libraries which is a more
+performance-oriented, lower-level language than Python. Data is shown for
+solving for various complexities of ODEs from the basic exponential to
+modeling sound wave propagation as we did in homework 4.
 
@@
 
-\begin{block}{Introduction}
+\begin{block}{Algorithm}
 
-Solving systems of differential equations can be a computationally expensive
-task. The error of most algorithms scales on the stepsize of the discretization
-of time. However, stepsize in time is also proportional to the computation
-required. The parareal algorithm, developed by YADA YADA. An alternative would
-be to parallelize by space.
+The parareal algorithm, developed by Lions, Maday and Turinici in 2001, is
+a general algorithm that allows parallelization in time.sIt does so by using a
+cheaper (lower order, lesser resolution) approximation first, and then makes
+corrections in parallel. The entire process is then iterative and tuned for a
+given amount of iterations. Therefore, the algorithm is $k$ repetitions of a
+serial coarse method updated by a finer method run in parallel for subsections.
 
-%------------------------------------------------
-
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{harvard_univ_logo.png}
-\caption{Figure caption}
-\end{figure}
-
-\end{block}
-
-%----------------------------------------------------------------------------------------
-
-\end{column} % End of the first column
-
-\begin{column}{\sepwid}\end{column} % Empty spacer column
-
-\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)
-
-\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column
-
-\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)
-
-%----------------------------------------------------------------------------------------
-%	MATERIALS
-%----------------------------------------------------------------------------------------
-
-\begin{block}{Algorithm}
-The general parareal method, as described by YADA YADA can be SEEN BELOW YAR
+Subscript: time index. Superscript: iteration index.
 
@@
     \item Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$
-    \item Compute the corrections $\delta g_n(u_1_n) =
+    \item Compute the corrections $\delta g_n(u^1_n) =
         g_{\textnormal{fine}}(t_n, u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel
@@
 
-\end{column} % End of column 2.1
+\end{column} % End of the first column
 
-\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)
+\begin{column}{\sepwid}\end{column} % Empty spacer column
+
+\begin{column}{\onecolwid} % Begin a column which is two columns wide (column 2)
+
+%\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column
+
+%\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)
 
 %----------------------------------------------------------------------------------------
-%	METHODS
+%	MATERIALS
 %----------------------------------------------------------------------------------------
-
-\begin{block}{Methods}
-
-Lorem ipsum dolor \textbf{sit amet}, consectetur adipiscing elit.
-Sed laoreet accumsan mattis.
-Integer sapien tellus, auctor ac blandit eget, sollicitudin vitae lorem.
-Praesent dictum tempor pulvinar.
-Suspendisse potenti.
-Sed tincidunt varius ipsum, et porta nulla suscipit et.
-Etiam congue bibendum felis, ac dictum augue cursus a.
-\textbf{Donec} magna eros, iaculis sit amet placerat quis, laoreet id est.
-In ut orci purus, interdum ornare nibh.
-Pellentesque pulvinar, nibh ac malesuada accumsan, urna nunc convallis tortor, ac vehicula nulla tellus eget nulla.
-Nullam lectus tortor, \textit{consequat tempor hendrerit} quis, vestibulum in diam.
-Maecenas sed diam augue.
-
-\end{block}
 
@@
 
-\end{column} % End of column 2.2
+%\end{column} % End of column 2.1
 
-\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width
+%\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)
+
+%----------------------------------------------------------------------------------------
+
+%\end{column} % End of column 2.2
+
+%\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width
 
@@
 
-\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again
+%\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column again
 
-\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
+%\begin{column}{\onecolwid} % The first column within column 2 (column 2.1)
 
@@
 
-%\begin{block}{Mathematical Section}
+\begin{block}{Theoretical Results}
 
-%Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
+\textbf{Convergence}
 
-%\begin{equation}
-%E = mc^{2}
-%\label{eqn:Einstein}
-%\end{equation}
-%
-%Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
-%
-%\begin{equation}
-%\cos^3 \theta =\frac{1}{4}\cos\theta+\frac{3}{4}\cos 3\theta
-%\label{eq:refname}
-%\end{equation}
-%
-%Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin. Nam quis odio enim, in molestie libero. Vivamus cursus mi at nulla elementum sollicitudin.
-%
-%\begin{equation}
-%\kappa =\frac{\xi}{E_{\mathrm{max}}} %\mathbb{ZNR}
-%\end{equation}
-%
-%\end{block}
+Assume the coarse operator $g_{\Delta t}$ is order $m$ and is Lipshitz, and the
+fine solution operator $g_{\textnormal{fine}}$ is a sufficiently accurate
+approximation to the analytic operator so we may replace $g_{\textnormal{fine}}
+\to g$.
+
+\emph{Theorem}: The order of accuracy of the parareal method with coarse
+solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. Can be proved by
+induction.
+
+\textbf{Stability}
+
+With the parareal method, it is possible to combine ODE solvers. Stability
+region depends on both $g_{\Delta t}$ and $g_{\textnormal{fine}}$, and the
+equation being solved.
+
+\textbf{Runtime and Speedup}
+
+Let's say the coarse operator runs in time $t$, and the fine operator runs in
+time $Q \cdot t$. Assume we have $N$ processors and we perform $k$ correction
+steps. Then, the runtime of parareal is, assuming negligible setup and
+aggregation time,
+
+\begin{equation}
+t + k(t + \frac{Qt}{N}).
+\end{equation}
+
+In order for there to be a speedup relative to the fine operator, we require
+that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
+- 1 - k}$.
+
+\end{block}
+
+\begin{block}{Methods}
+
+We tested some stuff on Odyssey.
+
+\end{block}
 
@@
 
+\begin{column}{\sepwid}\end{column}
+
 \begin{column}{\onecolwid} % The second column within column 2 (column 2.2)
+
+%----------------------------------------------------------------------------------------
+%	METHODS
+%----------------------------------------------------------------------------------------
 
@@
 
-\begin{block}{Results}
+\begin{block}{Preliminary Results}
 
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{harvard_univ_logo.png}
-\caption{Figure caption}
+\includegraphics[width=0.75\textwidth]{../data/runtime_vs_cpus.pdf}
+\caption{Running time vs. number of CPUs.}
+\label{fig:run_v_cpu}
 \end{figure}
 
-Nunc tempus venenatis facilisis. Curabitur suscipit consequat eros non porttitor. Sed a massa dolor, id ornare enim:
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{../data/error_vs_corrections.pdf}
+\caption{Error vs. number of correction steps.}
+\label{fig:err_v_k}
+\end{figure}
 
-\begin{table}
-\vspace{2ex}
-\begin{tabular}{l l l}
-\toprule
-\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
-\midrule
-Treatment 1 & 0.0003262 & 0.562 \\
-Treatment 2 & 0.0015681 & 0.910 \\
-Treatment 3 & 0.0009271 & 0.296 \\
-\bottomrule
-\end{tabular}
-\caption{Table caption}
-\end{table}
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{../data/error_vs_qualityfactor.pdf}
+\caption{Error vs. quality factor of the fine operator}
+\label{fig:err_v_q}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{../data/runtime_vs_corrections.pdf}
+\caption{Running time as a function of the number of correction steps.}
+\label{fig:run_v_k}
+\end{figure}
 
@@
 
-\end{columns} % End of the split of column 2
+%\end{columns} % End of the split of column 2
 
-\end{column} % End of the second column
+%\end{column} % End of the second column
 
@@
 
-Nunc tempus venenatis facilisis.
-\textbf{Curabitur suscipit} consequat eros non porttitor.
-Sed a massa dolor, id ornare enim.
-Fusce quis massa dictum tortor \textbf{tincidunt mattis}.
-Donec quam est, lobortis quis pretium at, laoreet scelerisque lacus.
-Nam quis odio enim, in molestie libero.
-Vivamus cursus mi at \textit{nulla elementum sollicitudin}.
-
 \end{block}
@@
 \begin{block}{Additional Information}
-
-Maecenas ultricies feugiat velit non mattis. Fusce tempus arcu id ligula varius dictum. 
-\begin{itemize}
-\item Curabitur pellentesque dignissim
-\item Eu facilisis est tempus quis
-\item Duis porta consequat lorem
-\end{itemize}
 
@@
 \begin{center}
-\begin{tabular}{ccc}
-%\includegraphics[width=0.4\linewidth]{harvard_univ_logo.png} & \hfill &
 \includegraphics[width=0.2\linewidth]{seas-rgb.png}
-\end{tabular}
 \end{center}

### 2014-12-09 poster/main.tex
more poster files

@@
 
-We implement the parareal algorithm to solve various ODEs and compare in
+We implement the Parareal algorithm to solve various ODEs and compare in
 terms of accuracy, speedup and efficiency. We explore the effect of using
@@
 Odyssey cluster. We analyze possible reasons as to why our implementation
-may not be ideal and some tradeoffs that are taken in the parareal
+may not be ideal and some tradeoffs that are taken in the Parareal
 implementation. Further optimizations are proposed and rationalized as next
@@
 
-The parareal algorithm, developed by Lions, Maday and Turinici in 2001, is
+The Parareal algorithm, developed by Lions, Maday and Turinici in 2001, is
 a general algorithm that allows parallelization in time.sIt does so by using a
@@
 
-\emph{Theorem}: The order of accuracy of the parareal method with coarse
+\emph{Theorem}: The order of accuracy of the Parareal method with coarse
 solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. Can be proved by
@@
 
-With the parareal method, it is possible to combine ODE solvers. Stability
+With the Parareal method, it is possible to combine ODE solvers. Stability
 region depends on both $g_{\Delta t}$ and $g_{\textnormal{fine}}$, and the
@@
 time $Q \cdot t$. Assume we have $N$ processors and we perform $k$ correction
-steps. Then, the runtime of parareal is, assuming negligible setup and
+steps. Then, the runtime of Parareal is, assuming negligible setup and
 aggregation time,
@@
 
-We tested some stuff on Odyssey.
+We implemented the Parareal algorithm in Python, using mpi4py to parallelize it.
+Our coarse operator was forward Euler step, with 100 steps, while the fine
+operator was forward Euler, with $Q * 100$ steps, where $Q$ is the quality
+factor. We tested the Parareal algorithm on two sets of differential equations:
+
+\[
+y'(t) = f(t, y) = \lambda y, \, y(0) = 1
+\]
+\[
+y''(t) + 2y'(t) + 5y(t) = 0, \, y(0) = 1 \, y'(0) = 0.
+\]
+
+We ran the algorithm on different numbers of processors and varied $k$ and $Q$.
 

### 2014-12-09 paper.tex
more updates

--- 2014-12-09 paper.tex
@@
 Yet it is worth noting that knowing how to divide the result vector requires knowledge of problem-specific details.
-Thus, this approach lacks the generalisability and theoretical derivaions that schemes that parallelize in time can offer.
+Thus, this approach lacks the generalizability and theoretical derivaions that schemes that parallelize in time can offer.
 
@@
 
+The beauty is the parareal method can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{fine}$ solution operators.
+The $g_{fine}$ solution is to a more computationally expensive method but assumed to provide either greater resolution or lower error.
+Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.
+The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code.
+
 \subsection{Visualization of Scheme}
 
-bksim TODO
+bksim TODO put a graph because maybe it'll help? the slides have a version
 
@@
 parameters to the parareal as well as possible applications in different ODE
-systems. Our baseline was looking to compare
+systems. Because there were so many different versions and parameteres for the parareal to test,
 
@@
 
-We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using GET PARAMETERES
+We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using WHAT WERE THE PARAMETERS - Coarse and Fine Num Steps
 
@@
 For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute.
+We wish to strong scaling, in which the computation time of our parallel method will decrease as we increase the number of processors.
+We compare this version to the serial version we did for Homework 4.
 
-bksim TODO
+bksim TODO GET DATA, AS REQUIRED FROM ABOVE, SHOW GRAPHS
 
 \subsection{Parallelism Paradigm Comparisons}
+
+Now with 3 different parallel implementations, two in time wiht the parareal, one in space,  we compare across take a breather, don't overwork,
 
@@
 
-TODO Wesley
+TODO WESLEY
 

### 2014-12-09 paper.tex

--- 2014-12-09 paper.tex
@@
 
-We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using GET PARAMETERS
+We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using WHAT WERE THE PARAMETERS - Coarse and Fine Num Steps
 
@@
 
-For the speaker problem, we designed a problem-specific method to parallelize in
-space and divide Pierce Hall into subsections such that each processor would
-receive a division of the building to compute.
+For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute.
+We wish to strong scaling, in which the computation time of our parallel method will decrease as we increase the number of processors.
+We compare this version to the serial version we did for Homework 4.
 
-bksim TODO
+bksim TODO GET DATA, AS REQUIRED FROM ABOVE, SHOW GRAPHS
 
@@
 
-TODO compare the computation times (and perhaps a slight change in accuracy of
-the results) between both parareal methods the best in the sections before hand
+Now with 3 different parallel implementations, two in time wiht the parareal, one in space,  we compare across TODO WESLSEY
+
+TODO compare the computation times (and perhaps a slight change in accuracy of the results) between both parareal methods the best in the sections before hand
 

### 2014-12-09 poster/main.tex
POSTER COMPLETE
Conflicts:
	paper.out
	paper.tex

@@
 
-\title{The Parareal Algorithm for Parallel Solution of ODEs} % Poster title
+\title{Parallelizing Numerical Methods for Solving ODEs: The Parareal Algorithm} % Poster title
 
@@
 
-We implement the Parareal algorithm to solve various ODEs and compare in
-terms of accuracy, speedup and efficiency. We explore the effect of using
-different solution operators (a coarse and a fine as required by the method).
-Our code is written in python using mpi4py to parallelize. We observe
+We look to achieve strong scaling in parallelizing numerical ODE methods.
+We test both parallelism by result space as well as parallelism by time, with the Parareal algorithm, to solve various ODEs.
+We measure performance and compare in
+terms of accuracy, speedup and efficiency. We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
 light speedups in our small scale tests up to 64 processors on Harvard's
@@
 implementation. Further optimizations are proposed and rationalized as next
-steps including ports of to C++ and BLAS libraries which is a more
-performance-oriented, lower-level language than Python. Data is shown for
-solving for various complexities of ODEs from the basic exponential to
-modeling sound wave propagation as we did in homework 4.
+steps including ports of to C++ using BLAS. Our code is written in python using mpi4py.
 
@@
 a general algorithm that allows parallelization in time.sIt does so by using a
-cheaper (lower order, lesser resolution) approximation first, and then makes
+cheaper (lower order or lesser resolution) approximation first, and then makes
 corrections in parallel. The entire process is then iterative and tuned for a
@@
 
-Subscript: time index. Superscript: iteration index.
-
+Parareal Algorithm:\\
+Subscript: Time [0,n], Superscript: Iteration [1,k]
 \begin{itemize}
@@
         t}(t_n, u^2_n) + \delta g_n(u^1_n)$
-    \item Repeat steps 2 and 3, incrementing the iteration label and using
+    \item Recompute corrections and adjustment, for each k \\ using
         $u^{k+1}_0 = u^1_0$ as the initial condition
@@
 %\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)
-
-%----------------------------------------------------------------------------------------
-%	MATERIALS
-%----------------------------------------------------------------------------------------
-
-%----------------------------------------------------------------------------------------
 
@@
 %----------------------------------------------------------------------------------------
-
-\begin{block}{Theoretical Results}
+\begin{block}{Theoretical Analysis}
 
@@
 
-\textbf{Runtime and Speedup}
+\textbf{Parallel Performance Tradeoff Analysis}
 
@@
 
-In order for there to be a speedup relative to the fine operator, we require
+The first t seconds comes from the first coarse aprpoximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
 that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
 - 1 - k}$.
+This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{fine}$.
+However, this optimal k is a problem-specific parameter and must be optimized for.
 
 \end{block}
+
+%----------------------------------------------------------------------------------------
+%	METHODS
+%----------------------------------------------------------------------------------------
 
@@
 %----------------------------------------------------------------------------------------
-%	METHODS
-%----------------------------------------------------------------------------------------
-
-%----------------------------------------------------------------------------------------
 %	RESULTS
@@
 
-\begin{block}{Preliminary Results}
+\begin{block}{Preliminary \\ Implementation Results}
 
@@
 %----------------------------------------------------------------------------------------
-%	CONCLUSION
+%	Discussion
 %----------------------------------------------------------------------------------------
 
-\begin{block}{Conclusion}
+\begin{block}{Discussion and Next Steps}
 
+Our preliminary results do not look as optimal as we had hoped.
+We see a significant overhead in the iterative Parareal algorithm when compared to the serial computations of $g_{fine}$ directly.
+However, we do see that the Parareal algorithm itself displays both strong and weak scaling, as expected.
+Yet the biggest tradeoff we look to better demonstrate will be the heavy dependence on the two methods chosen for $g_{\Delta t}$ and $g_{fine}$.
+\\
+
+Most of the immediate work will involve increased data collection with large N.
+The Parareal algorithm provides maximum speedup when N approaches the number of discretizations in time for $g_{\Delta t}$ which would allow all corrections to be made at one step.
+We are also in the process of implementing and switching out various $g_{fine}$ schemes.
+We are looking to employ higher order methods at the same temporal resolution as opposed to a higher-resolution equal-order method.
+This should help show our ever convergence of the Parareal method and may also emphasize overall speedup as well.
 \end{block}
@@
 %----------------------------------------------------------------------------------------
-%	ADDITIONAL INFORMATION
+%	Conclusion and Next Steps
 %----------------------------------------------------------------------------------------
 
-\begin{block}{Additional Information}
+\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
+\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors
 
-\end{block}
+\begin{alertblock}{Conclusion}
+
+We have implemented and begun to explore some techniques of looking
+for strong and weak scaling efficiencies for solving ODEs. The Parareal
+algorithm is beautiful in its theoretical advantages---in terms of stability,
+error and efficiency. However, the method, as a very generalized method,
+requires deeper analysis for the specific problem.
+
+\end{alertblock}
 
@@
 
+\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
 \begin{block}{References}
 
-\nocite{*} % Insert publications even if they are not cited in the poster
-\small{\bibliographystyle{unsrt}
-\bibliography{sample}\vspace{0.75in}}
+\begin{itemize}
+\item Lions, J., Maday, Y and Turinici, G. "A 'Parareal' in Time Discretizion of PDE's". 2001. \textit{Comptes Rendus de l'Academie des Sciences Series I Mathematics} 332(7): 661-668\\
+
+\item Field, Scott. "Parareal Methods Field Talk". Dec, 2009. Retried from \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
+
+\end{itemize}
+
+%\nocite{*} % Insert publications even if they are not cited in the poster
+%\small{\bibliographystyle{unsrt}
+%\bibliography{sample}\vspace{0.75in}}
 
 \end{block}
-
-%----------------------------------------------------------------------------------------
-%	ACKNOWLEDGEMENTS
-%----------------------------------------------------------------------------------------
-
-\setbeamercolor{block title}{fg=red,bg=white} % Change the block title color
-
-\begin{block}{Acknowledgements}
-
-\small{\rmfamily{Nam mollis tristique neque eu luctus.
-Suspendisse rutrum congue nisi sed convallis.
-Aenean id neque dolor.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.}} \\
-
-\end{block}
-
-%----------------------------------------------------------------------------------------
-%	CONTACT INFORMATION
-%----------------------------------------------------------------------------------------
-
-\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
-\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors
-
-\begin{alertblock}{Contact Information}
-
-\begin{itemize}
-\item Web: \href{http://www.university.edu/smithlab}{http://www.university.edu/smithlab}
-\item Email: \href{mailto:john@smith.com}{john@smith.com}
-\item Phone: +1 (000) 111 1111
-\end{itemize}
-
-\end{alertblock}
-
-\begin{center}
-\includegraphics[width=0.2\linewidth]{seas-rgb.png}
-\end{center}
-
-%----------------------------------------------------------------------------------------
 

### 2014-12-09 poster/main.tex
poster typos fixed

@@
 The Parareal algorithm, developed by Lions, Maday and Turinici in 2001, is
-a general algorithm that allows parallelization in time.sIt does so by using a
+a general algorithm that allows parallelization in time. It does so by using a
 cheaper (lower order or lesser resolution) approximation first, and then makes
@@
 
-The first t seconds comes from the first coarse aprpoximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
+The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
 that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
@@
 
-\item Field, Scott. "Parareal Methods Field Talk". Dec, 2009. Retried from \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
+\item Field, Scott. "Parareal Methods Field Talk". Dec, 2009. Retrieved from \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
 

### 2014-12-10 paper.tex
expanding more on paper

@@
 \begin{abstract}
-We implement the parareal algorithm to solve various ODEs and compare in
-terms of accuracy, speedup and efficiency. We explore the effect of using
-different solution operators (a coarse and a fine as required by the method).
-Our code is written in python using mpi4py to parallelize. We observe
+
+We look to achieve strong scaling in parallelizing numerical ODE methods.
+We test both parallelism by result space as well as parallelism by time, with the Parareal algorithm, to solve various ODEs.
+We measure performance and compare in
+terms of accuracy, speedup and efficiency. We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
 light speedups in our small scale tests up to 64 processors on Harvard's
 Odyssey cluster. We analyze possible reasons as to why our implementation
-may not be ideal and some tradeoffs that are taken in the parareal
+may not be ideal and some tradeoffs that are taken in the Parareal
 implementation. Further optimizations are proposed and rationalized as next
-steps including ports of to C++ and BLAS libraries which is a more
-performance-oriented, lower-level language than Python. Data is shown for
-solving for various complexities of ODEs from the basic exponential to
-modeling sound wave propagation as we did in Homework 4.
+steps including ports of to C++ using BLAS.
+Our code is written in python using mpi4py.
+We compare tests from various data sources from basic exponential ODEs to larger grids of ODEs to solve for in applications like diffusion and sound wave propagation.
+
 \end{abstract}
@@
 
-Another obstacle to trying to paralleliize numerical methods is that the methods
-are inherently serial in time---in that evaluation of the next time $n+1$
+Another obstacle to trying to parallelize numerical methods is that the methods
+are inherently serial in time---in that evaluation of the next time point $n+1$
 depends on the previous values, say those at time $n-1$ and $n$. This setup
@@
 create schemes that have parallelizable components---such as an update or a
-refinement to a serial portion. As will be discussed and analyzed in the
-metodology section, one such scheme would be the parareal algorithm.sFor an
+refinement built on top of a more basic method first computed in serial. As will be discussed and analyzed in the
+metodology section, one such scheme would be the parareal algorithm. As an
 overview, the parareal algorithm first does a coarse (low-order and fast)
@@
 method is expected to outperform the parallel in time paradigm simply because
-the task is somewhat embarassingly parallel, being able to be naturally divided
+the task is embarassingly parallel, being able to be naturally divided
 into smaller but identical in method problems. There is no serial component in
@@
 In general, the division in space paradigm is more intuitive to parallelize and
-will often lead to greater speedups because there is no serial component. Yet
-it is worth noting that knowing how to divide the result vector requires
-knowledge of problem-specific details. Thus, this approach lacks the
-general stability and theoretical derivations that schemes that parallelize in
+will often lead to greater speedups because there is no serial component.
+The main tradeoff however, is that knowing how to divide the resultant system requires knowledge of problem-specific details.
+Thus, this approach lacks the
+general stability and theoretical proofs (of error and convergence) that schemes which parallelize in
 time can offer.
+
+\subsection{Measuring Parallel Performance}
 
@@
 The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
-a general algorithm that allows parallelization in time.sIt does so by using a
-cheaper (lower order, lesser resolution) approximation first, and then makes
-corrections in parallel. The entire process is then iterative and tuned for a
-given amount of iterations. Therefore, the algorithm is $k$ repetitions of a
-serial coarse method updated by a finer method run in parallel for subsections.
+a generalalized algorithm which allows for parallelization in time. It does so by using a
+cheaper ($g_{\Delta t}$)---lower order or lesser resolution, approximation first which it then corrects by evaluating segments with a finer method ($g_{fine}$).
+The correction step from the approximation can be done in parallel.
+Multiple iterations can be performed to achieve an error provably asymptotically equal to a full serial computation of $g_{fine}$.
+The exact number of iterations required to achieve certain levels of convergence depends on the problem and must be tuned top optimize for speedup for a given system.
+In summary, the algorithm is $k$ repetitions of a
+correction process which requires running updated by a finer method run in parallel for subsections.
 
@@
 
+With the Parareal method, it is possible to combine ODE solvers. The stability region depends on both $g_{\Delta t}$ and $g_{\textnormal{fine}}$, and the
+equation being solved.
+
 FROM THE SLIDES
@@
 TODO BKSIM? ASHI carry?
+
+\subsection{Convergence of the Parareal}
+
+TODO BKSIM
+
+Assume the coarse operator $g_{\Delta t}$ is order $m$ and is Lipshitz, and the
+fine solution operator $g_{\textnormal{fine}}$ is a sufficiently accurate
+approximation to the analytic operator so we may replace $g_{\textnormal{fine}}
+\to g$.
+
+\emph{Theorem}: The order of accuracy of the Parareal method with coarse
+solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. Can be proved by
+induction.
+
+SHALL WE DO IT THEN?
 
@@
 
-\subsection{Analysis of Speedup and Efficiency}
+\subsection{Parallel Tradeoff Analysis}
 
-ashi TODO ashia from wesley's handwritten analysis
+ashi TODO
+
+Let's say the coarse operator runs in time $t$, and the fine operator runs in
+time $Q \cdot t$. Assume we have $N$ processors and we perform $k$ correction
+steps. Then, the runtime of Parareal is, assuming negligible setup and
+aggregation time,
+
+\begin{equation}
+t + k(t + \frac{Qt}{N}).
+\end{equation}
+
+The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
+that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
+- 1 - k}$.
+This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{fine}$.
+However, this optimal k is a problem-specific parameter and must be optimized for.
 
@@
 
-We looked at both $e^x$ as well as TODO WHAT OTHER SCHEME using WHAT WERE THE PARAMETERS - Coarse and Fine Num Steps
+We implemented the Parareal algorithm in Python, using mpi4py to parallelize it.
+Our coarse operator was forward Euler step, with 100 steps, while the fine
+operator was forward Euler, with $Q * 100$ steps, where $Q$ is the quality
+factor. We tested the Parareal algorithm on two sets of differential equations:
+
+\[
+y'(t) = f(t, y) = \lambda y, \, y(0) = 1
+\]
+\[
+y''(t) + 2y'(t) + 5y(t) = 0, \, y(0) = 1 \, y'(0) = 0.
+\]
+
+We ran the algorithm on different numbers of processors and varied $k$ and $Q$.
 
@@
 
-ashi TODO add figures and check my logic/bs to match it to the data
+TODO ashi sort out the figure and match and replace if needbe
+
+\begin{figure}
+\includegraphics[width=0.75\textwidth]{data/runtime_vs_cpus.pdf}
+\caption{Running time vs. number of CPUs.}
+\label{fig:run_v_cpu}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{data/error_vs_corrections.pdf}
+\caption{Error vs. number of correction steps.}
+\label{fig:err_v_k}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{data/error_vs_qualityfactor.pdf}
+\caption{Error vs. quality factor of the fine operator}
+\label{fig:err_v_q}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{data/runtime_vs_corrections.pdf}
+\caption{Running time as a function of the number of correction steps.}
+\label{fig:run_v_k}
+\end{figure}
 

### 2014-12-10 pararealAlgo.py

@@ -92,7 +92,7 @@ def parallel_corrections(deriv, upast, tmin, tmax, numSteps, qualityFactor,
 
     # values for the fine computations
     dt = float(tmax - tmin) / (numSteps - 1)
-    fineNumSteps = numSteps*qualityFactor
+    fineNumSteps = numSteps * qualityFactor
     fineDt = dt / qualityFactor
 
     # broadcast upast

### 2014-12-10 serial_integrate.py
added more integration methods

@@ -1,0 +1,92 @@
+import numpy as np
+
+def integrate(deriv, init, time_min, time_max, numSteps, step):
+    '''
+    Solves a differential equation using a callable method from time_min to
+    time_max, using a step of time_step
+
+    deriv: callable; the derivative
+    init: numpy 1D array; initial condition
+    time_min: float; Minimum time
+    time_max: float; Maximum time
+    numSteps: int; number of stes
+    step: callable; takes deriv, current time, current y, and time step and
+    returns the next current y
+    '''
+    dim = init.shape[0]
+    # checks for 1D array
+    assert(init.shape == (dim,))
+
+    # initialize times
+    times = np.linspace(time_min, time_max, numSteps+1)
+    time_step = float(time_max - time_min) / (numSteps - 1)
+
+    ys = np.zeros((numSteps+1, dim))
+    ys[0] = init
+
+    for ind in xrange(numSteps):
+        t = times[ind]
+        curr = ys[ind]
+        ys[ind + 1] = step(deriv, t, curr, time_step)
+
+    return ys
+
+
+def forward_euler_step(deriv, t, u, dt):
+    return(u + dt * deriv(t, u))
+
+def forward_euler(deriv, init, time_min, time_max, numSteps):
+    return(integrate(deriv, init, time_min, time_max, numSteps,
+            forward_euler_step))
+
+def rk4_step(deriv, t, u, dt):
+    k1 = deriv(t, u)
+    k2 = deriv(t + (dt/2.), u + 0.5*k1*dt)
+    k3 = deriv(t + (dt/2.), u + 0.5*k2*dt)
+    k4 = deriv(t + dt, u + k3*dt)
+
+    return(u + ((dt/6.) * (k1 + 2*k1 + 2*k3 + k4)))
+
+def midpoint_step(deriv, t, u, dt):
+    k1 = dt * deriv(t, u)
+    return(u + dt*deriv(t + 0.5*dt, u + 0.5*k1))
+
+def midpoint_method(deriv, init, time_min, time_max, numSteps):
+    return(integrate(deriv, init, time_min, time_max, numSteps, midpoint_step))
+
+def rk4(deriv, init, time_min, time_max, numSteps):
+    return(integrate(deriv, init, time_min, time_max, numSteps, rk4_step))
+
+if __name__ == "__main__":
+
+    def deriv(t, u):
+        y = u[0]
[…]

### 2014-12-10 paper.tex

@@
 
-ashi TODO
+We look to compute a theoretical maximum speedup enabled by the nature of the Parareal algorithm.
+As this is a tradeoff between error (with parameter k being the number of iterations of corrections, each which takes another cycle of $g_{\Delta t}$, we seek to express the tradeoff as a function of the quality factor, $Q$, the number of iterations, $k$ and the number of processors $n$.
+Our quality factor is defined as the multiplicative factor of additional time necessary to
 
-Let's say the coarse operator runs in time $t$, and the fine operator runs in
-time $Q \cdot t$. Assume we have $N$ processors and we perform $k$ correction
-steps. Then, the runtime of Parareal is, assuming negligible setup and
+Then, the runtime of Parareal is, assuming negligible setup and
 aggregation time,
@@
 
+\subsection{Expectations}
+TODO WESLEY
+\subsection{General Remarks about Resuts}
+One of the TODO WESEY
+
 \subsection{Comparison to Serial with $g_{fine}$ as a Forward Euler with
@@
 Varying Iterations:
-
+]
 Varying Quality Factor:
@@
 
-Now with 3 different parallel implementations, two in time wiht the parareal, one in space,  we compare across TODO WESLSEY
+Now with 3 different parallel implementations, two in time with the parareal, one in space, we compare across the methods to see if we can make any statements about general performance of our different paradigms.
+
+TODO WESLEY
 

### 2014-12-10 paper.tex
another hour push

@@
 a generalalized algorithm which allows for parallelization in time. It does so by using a
-cheaper ($g_{\Delta t}$)---lower order or lesser resolution, approximation first which it then corrects by evaluating segments with a finer method ($g_{fine}$).
+cheaper ($g_{\Delta t}$)---lower order or lesser resolution, approximation first which it then corrects by evaluating segments with a finer method ($g_{\textnormal{fine}}$).
 The correction step from the approximation can be done in parallel.
-Multiple iterations can be performed to achieve an error provably asymptotically equal to a full serial computation of $g_{fine}$.
+Multiple iterations can be performed to achieve an error provably asymptotically equal to a full serial computation of $g_{\textnormal{fine}}$.
 The exact number of iterations required to achieve certain levels of convergence depends on the problem and must be tuned top optimize for speedup for a given system.
@@
 
-The beauty is the parareal method can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{fine}$ solution operators.
-The $g_{fine}$ solution is to a more computationally expensive method but assumed to provide either greater resolution or lower error.
+The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators.
+The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.
 Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.
 The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code.
+We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.
 
-\subsection{Visualization of Scheme}
+\subsection{Visualization of Parareal Scheme}
 
@@
 
+TODO WESLEY
+
 \begin{equation}
@@
 - 1 - k}$.
-This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{fine}$.
+This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{\textnormal{fine}}$.
 However, this optimal k is a problem-specific parameter and must be optimized for.
+
+TODO PARALLEL IN SPACE ANALYSIS
+
+\subsection{Incorporating Communication}
+
+TODO WESLEY
+
+TODO PARALLEL IN SPACE ANALYSIS
+
+\subsection{Amdahl's Law Analysis}
+
+TODO WESLEY
+
+TODO PARALLEL IN SPACE ANALYSIS
 
@@
 \subsection{General Remarks about Resuts}
-One of the TODO WESEY
+One of the TODO WESLEY
 
-\subsection{Comparison to Serial with $g_{fine}$ as a Forward Euler with
+\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as a Forward Euler with
 Smaller $\Delta t$}
@@
 Varying Quality Factor:
-We define the quality factor in this case for how many times smaller the Euler method used for $g_{fine}$ is compared to the Euler method used for $g_{\Delta t}$
+We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$
 
@@
 
-\subsection{Comparison to Serial with $g_{fine}$ as Higher Order Methods}
+\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as Higher Order Methods}
 
@@
 
-\subsection{Strong and Weak Scaling Analysis}
+\subsection{Demonstrating Scaling}
 
 ashi TODO by taking some numbers and following
-\url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
-
-\subsection{Demonstrating Scaling}
 
@@
 
-\subsection{Possible Optimizations}
-
-TODO WESLEY
-
-\section{Conclusion and Future Work}
+\subsection{Possible Optimizations and Future Work}
 
@@
 
+TODO Wesley
+
+\section{Conclusion and Future Work}
+
 All in all, we have implemented and begun to explore some techniques of looking

### 2014-12-10 paper.tex
ready to parallelize the process of paper writing

@@
 
-\subsection{Measuring Parallel Performance}
+\subsection{Measuring Parallel Performance and Amdahl's Law}
+
+Performance of parallel algorithms are usually measured in terms of 3 different metrics: speedup, strong scalling efficiency and weak scaling efficiency.
 
@@
 TODO WESLEY
-\subsection{General Remarks about Resuts}
+\subsection{General Remarks about Results}
 One of the TODO WESLEY

### 2014-12-10 paper.tex
minor changes to paper
added more stuff to paper about error

@@
-%ssssFile: paper.tex
-%ss Created: Sat Dec 06 04:00 PM 2014 E
+%File: paper.tex
+%Created: Sat Dec 06 04:00 PM 2014 E
 %
@@
 
-\tableofcontents
+%\tableofcontents
 
-\clearpage
+%\clearpage
 
@@
 The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
-a generalalized algorithm which allows for parallelization in time. It does so by using a
-cheaper ($g_{\Delta t}$)---lower order or lesser resolution, approximation first which it then corrects by evaluating segments with a finer method ($g_{\textnormal{fine}}$).
-The correction step from the approximation can be done in parallel.
-Multiple iterations can be performed to achieve an error provably asymptotically equal to a full serial computation of $g_{\textnormal{fine}}$.
-The exact number of iterations required to achieve certain levels of convergence depends on the problem and must be tuned top optimize for speedup for a given system.
-In summary, the algorithm is $k$ repetitions of a
-correction process which requires running updated by a finer method run in parallel for subsections.
+a generalized algorithm which allows for parallelization in time. It does so
+by using a cheaper ($g_{\Delta t}$)---a lower order or lesser resolution
+approximation first, which it then corrects by evaluating segments with a finer
+method $g_{\textnormal{fine}}$. The correction step from the approximation can
+be done in parallel. Multiple iterations can be performed to achieve an error
+provably asymptotically equal to a full serial computation of
+$g_{\textnormal{fine}}$. The exact number of iterations required to achieve
+certain levels of convergence depends on the problem and must be tuned top
+optimize for speedup for a given system. In summary, the algorithm is $k$
+repetitions of a correction process which requires running updated by a finer
+method run in parallel for subsections.
 
@@
     \KwIn{Coarse scheme $g_{\Delta t}$}
-    \KwIn{Finersscheme $g_{\textnormal{fine}}$}
+    \KwIn{Finer scheme $g_{\textnormal{fine}}$}
     Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
-    Compute the corrections $\delta g_n(u^1_n) = g_{\textnormal{fine}}(t_n,
+    Compute the corrections $\delta g(u^1_n) = g_{\textnormal{fine}}(t_n,
     u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel\;
     Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta t}(t_n,
-    u^2_n) + \delta g_n(u^1_n)$\;
+    u^2_n) + \delta g(u^1_n)$\;
     Repeat steps 2 and 3, incrementing the iteration label and using $u^{k+1}_0
@@
 
+\subsection{Convergence of the Parareal}
+
+Assume the coarse operator $g_{\Delta t}$ is Lipschitz and order $m$: for some
+constant $L$, 
+\[ |g_{\Delta t}(t_n, u) - g_{\Delta t}(t_n, v)| \leq (1 + L \Delta t) |u - v|
+\, \forall t \in (0, t_n), \]
+\[ |u(t_n) - u^1_N | \leq C(\Delta t)^m |u_0|. \]
+
+Additionally, assume the function $u$ is bounded on $(0, t_n)$. Furthermore,
+assume the fine solution operator $g_{\textnormal{fine}}$ is a sufficiently
+accurate approximation to the analytic operator so we may replace
+$g_{\textnormal{fine}} \to g$.
+
+\textbf{Theorem}: The order of accuracy of the Parareal method with coarse
+solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
+this using induction [Bal].
+
+\emph{Proof}: By induction.
+
+Case $k = 1$: This is just the coarse operator, which is order $m$.
+
+Assume that this is true for $k$, that $|u(t_N) - u^k_N| \leq C(\Delta t)^{mk}
+|u_0|$.
+
+For the case $k + 1$, we have
+\[
+\begin{aligned}
+    |u(t_N) - u^{k+1}_N| &= 
+    |g(u(t_{N-1})) - g_{\Delta t}(u^{k+1}_{N-1}) - \delta g(u^k_{N-1})| \\
+    &= |g_{\Delta t}(u(t_{N-1})) - g_{\Delta t}(u^{k+1}_{N-1}) - \delta
+    g(u^k_{N-1}) + \delta g(u(t_{N-1}))| \\
+    &\leq |g_{\Delta t}(u(t_{N-1})) - g_{\Delta t}(u^{k+1}_{N-1})| - |\delta
+    g(u^k_{N-1}) + \delta g(u(t_{N-1}))| \\
+    &\leq (1 + C \Delta t) |u(t_{N-1} - u_{N-1}^{k+1}| + C(\Delta t)^{m+1} |
+    u_{N-1}^k - u(t_{N-1}) | \\
+    &\leq (1 + C \Delta t) |u(t_{N-1}) - u_{N-1}^{k+1}| + C(\Delta t)^{m(k+1)+1}
+    | u_0 | \\
+\end{aligned}
+\]
+
+We can continue to expand $|u(t_{N-1} - u_{N-1}^{k+1}|$ and get time indices of
+$N-1, N-2, \ldots, 2, 1$. This implies that 
+\[ |u(t_N) - u^{k+1}_N| \leq C(\Delta t)^{m(k+1)} |u_0|, \]
+as desired.
+
+\subsection{Error of the Parareal}
+
+With the assumptions of the previous section, this parareal method will
+approach, with large enough $k$ (correction iterations) to approach the error of
+the fine method. However, there is a time vs. accuracy tradeoff. Let $Q$ be the
+\emph{Quality Factor} for $g_{\Delta t}$ and $g_{\textnormal{fine}}$. Suppose
+that, for constant number of processors and $\Delta t$, $g_{\Delta t}$ runs in
+time $T$. Then, $g_{\textnormal{fine}}$ runs in time $QT$. With too large of a
+$k$ and a lower $Q$, the time for the Parareal could potentially take longer
+than the direct serial computation. We note that there could be other
+relationships between the time of the coarse and fine schemes other than
+multiplication by a constant term.
+
+Disregarding the time relationship between $g_{\Delta t}$ and
+$g_{\textnormal{fine}}$, we can look at what happens to the error as $k \to N$
+(recall that $N$ is the total number of time points in the time discretization).
+
+For $k = 1, 2, \ldots$ we have 
+\[u_{n+1}^{k+1} = g_{\Delta t} (t_n, u_n^{k+1}) + \big(
+g_{\textnormal{fine}}(t_n, u_n^k) - g_{\Delta t}(t_n, u_n^k). \]
+
+As $k \to N$, the Parareal algorithm gives $u_{n}^{k+1} = u^k_n$, so the order
+of accuracy approaches that of $g_{\textnormal{fine}}$.
+
+% I don't understand the above part ~ Andy
+
 \subsection{Stability of the Parareal}
@@
 
-TODO BKSIM? ASHI carry?
-
-\subsection{Convergence of the Parareal}
-
-TODO BKSIM
-
-Assume the coarse operator $g_{\Delta t}$ is order $m$ and is Lipshitz, and the
-fine solution operator $g_{\textnormal{fine}}$ is a sufficiently accurate
-approximation to the analytic operator so we may replace $g_{\textnormal{fine}}
-\to g$.
-
-\emph{Theorem}: The order of accuracy of the Parareal method with coarse
-solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. Can be proved by
-induction.
-
-SHALL WE DO IT THEN?
-
-\subsection{Error of the Parareal}
-
-FROM THE SLIDES? + Adjust?
-TODO BKSIM? ASHI carry?
-
-Show that this parareal method will approach, with large enough $k$ (enough
-iterations) to approach the error of the fine method. However, with too large of
-a $k$ and a lower quality factor (ratio of fine to coarse), the time for the
-parareal could potentially take longer than the direct serial computation.
+ASHI carry?
 
@@
 \begin{equation}
-t + k(t + \frac{Qt}{N}).
+t + k \left(t + \frac{Qt}{N} \right).
 \end{equation}
@@
 parameters to the parareal as well as possible applications in different ODE
-systems. Because there were so many different versions and parameteres for the parareal to test,
+systems. Because there were so many different versions and parameters for the Parareal to test,
 
@@
 \begin{figure}
+\begin{center}
 \includegraphics[width=0.75\textwidth]{data/runtime_vs_cpus.pdf}
@@
 \label{fig:run_v_cpu}
+\end{center}
 \end{figure}
@@
 \begin{figure}
+\begin{center}
 \includegraphics[width=0.75\linewidth]{data/error_vs_corrections.pdf}
@@
 \label{fig:err_v_k}
+\end{center}
 \end{figure}
@@
 \begin{figure}
+\begin{center}
 \includegraphics[width=0.75\linewidth]{data/error_vs_qualityfactor.pdf}
@@
 \label{fig:err_v_q}
+\end{center}
 \end{figure}
@@
 \begin{figure}
+\begin{center}
 \includegraphics[width=0.75\linewidth]{data/runtime_vs_corrections.pdf}
@@
 \label{fig:run_v_k}
+\end{center}
 \end{figure}
@@
 
+%\bibliographystyle{unsrt}
+%\bibliography{poster/sample}
+
+Guilliame Bal paper
+
+Stability of the Parareal Algorithm by Staff et al
+
 \end{document}

### 2014-12-10 paper.tex
wesley updates
added some stuff

@@
 \[u_{n+1}^{k+1} = g_{\Delta t} (t_n, u_n^{k+1}) + \big(
-g_{\textnormal{fine}}(t_n, u_n^k) - g_{\Delta t}(t_n, u_n^k). \]
+g_{\textnormal{fine}}(t_n, u_n^k) - g_{\Delta t}(t_n, u_n^k)\big). \]
 
@@
 
-FROM THE SLIDES
+Let us consider the ODE $du/dt = \lambda u$. Let $g_{\textnormal{fine}}(t_n,
+u_n) = \bar{g_{\textnormal{fine}}}u_n$ and $g_{\Delta t}(t_n, u_n) =
+\bar{g_{\Delta t}} u_n$. As shown by Staff et al. [cite], the parareal method
+becomes
 
-ASHI carry?
+\[ u_n^k = \left( \sum_{j=0}^k {n \choose j} (\bar{g_{\textnormal{fine}}} -
+g_{\Delta t})^j \bar{g_{\Delta t}}^{n-j} \right) u_0 = H(\bar{g_{\Delta
+t}}, \bar{g_{\textnormal{fine}}}, n, k, \lambda)u_0 \]
+
+This is stable if $\textnormal{max}_{n, k} |H| \leq 1$.
+
+When $\lambda \in \mathbb{R}$ and $\lambda \leq 0$, Staff et. al also show that
+
+\[
+    \begin{aligned}
+    |H| &\leq \sum_{j=0}^n {n \choose j} |\bar{g_{\textnormal{fine}}} -
+    g_{\Delta t}|^j |\bar{g_{\Delta t}}|^{n-j} \\
+    &= (|\bar{g_{\textnormal{fine}}} - g_{\Delta t}| + |\bar{g_{\Delta
+    t}}|)^n \leq 1
+    \end{aligned}
+\]
+
+To satisfy these conditions, we have that 
+\begin{enumerate}
+    \item $| \bar{g_{\textnormal{fine}}} | \leq 1$: This is the usual stability
+        requirement.
+    \item $|\bar{g_{\textnormal{fine}}} - 2 \bar{g_{\Delta t}}| \leq 1$.
+\end{enumerate}
 
@@
 TODO WESLEY
+
 \subsection{General Remarks about Results}
-One of the TODO WESLEY
+LETS SEE IF I HAVE TO DISCLAIM ANY GLOBAL FAILURES IN OUR RESULTS
 
@@
 \[
-y'(t) = f(t, y) = \lambda y, \, y(0) = 1
+y'(t) = f(t, y) = \lambda y, \, \, y(0) = 1
 \]
 \[
-y''(t) + 2y'(t) + 5y(t) = 0, \, y(0) = 1 \, y'(0) = 0.
+y''(t) + 2y'(t) + 5y(t) = 0, \, \, y(0) = 1, \, y'(0) = 0.
 \]

### 2014-12-10 parallel_space/custom_plot.py
merge

@@ -1,0 +1,105 @@
+import numpy as np
+from math import *
+import matplotlib.pyplot as plt
+import sys
+
+# Returns a scaled value of a function in the range 0 to 1, truncating it if
+# necessary
+def fscale(v,vmin,vsca):
+    vs=(v-vmin)*vsca
+    if vs<0: vs=0
+    if vs>1: vs=1
+    return vs
+
+# Returns a red->white->blue color scheme
+def palette1(v):
+    v*=2
+    if(v>1):
+        v=2-v
+        return (v,v,min(1,0.5+3*v))
+    else:
+        return (min(1,0.5+3*v),v,v)
+
+# Outputs a 2D image from a field using the palette1 color scheme
+# fn: the filename to save to
+# p: the 2D field to plot
+# wa: the matrix describing walls to overlay
+# (vmin,vmax): the field range
+# ups: the upsampling factor to use (so each field point is convert into an ups
+#      by ups square)
+def plot1(fn,p,wa,vmin,vmax,ups):
+
+    # Check matrix dimensions are the same
+    (m,n)=p.shape
+    if (m,n)!=wa.shape:
+        print "Matrix dimension mismatch"
+
+    # Set up output array and scaling constant
+    o=np.zeros((m*ups,n*ups,3))
+    vsca=1.0/(vmax-vmin)
+
+    # Assemble the output array
+    for i in range(m):
+        iu=i*ups
+        for j in range(n):
+            ju=j*ups
+            if wa[i,j]==1:
+                o[iu:iu+ups,ju:ju+ups,0]=0
+                o[iu:iu+ups,ju:ju+ups,1]=0
+                o[iu:iu+ups,ju:ju+ups,2]=0
+            else:
+                (re,gr,bl)=palette1(fscale(p[i,j],vmin,vsca))
+                o[iu:iu+ups,ju:ju+ups,0]=re
+                o[iu:iu+ups,ju:ju+ups,1]=gr
+                o[iu:iu+ups,ju:ju+ups,2]=bl
+    
+    # Save the image
+    plt.figure()
+    plt.imshow(o)
+    plt.show()
+    plt.imsave(fn,o)
+
+# Returns a black->red->yellow color scheme (matching Gnuplot's PM3D scheme)
+def palette2(v):
+    if v>0.5:
+        vs=0
+    else:
+        vs=sin(2*pi*v)
+    return (sqrt(v),v*v*v,vs)
+
[…]

### 2014-12-10 parallel_space/testing.py

@@ -1,0 +1,92 @@
+# Parallelism by Space 
+import numpy as np
+import custom_plot
+import copy
+from mpi4py import MPI
+import time
+import argparse
+import itertools
+
+# (a)
+# pressure in region S
+def pressure_s(t, p=10., w=100.*np.pi):
+    return p*np.sin(w*t)
+
+def pressure(up, right, down, left, center, prev, c=34300., h=36.6):
+    dt = h/2./c
+    return c*c/h/h*dt*dt*(up+right+down+left-4*center)+2.*center-prev
+
+p_B = [0,0] # person at (10, 167)
+p_C = [0,0] # person at (35, 73)
+p_M = [0,0] # person at (67, 115)
+
+h = 36.6
+c = 34300.
+
+pierce = np.loadtxt("pierce.txt", dtype=np.int8)
+
+# simulates from t = 0 to given t
+def simulate(t, fn):
+    region = np.zeros([100, 200])
+    
+    # set initial condition in region S
+    for j in xrange(57, 61):
+        for k in xrange(15, 19):
+            region[j][k] = 1
+    
+    p0 = np.zeros([100, 200])
+    p1 = np.zeros([100, 200])
+    for (x, y),v in np.ndenumerate(p1):
+        if region[x][y] == 1:
+            p1[x][y] = pressure_s(h/2./c)
+    
+    p_pprev = copy.copy(p0)
+    p_prev = copy.copy(p1)
+    p_curr = copy.copy(p_prev)
+    
+    t0 = h/c
+    dt = h/2./c
+
+    for t in np.arange(t0, t+dt/2., dt):
+        for (x, y), v in np.ndenumerate(p_prev):
+            # nothing on map for these values
+            if x == 0 or x == 99 or y == 0 or y == 199:
+                continue
+            
+            # region S is driven to satisfy this condition
+            if region[x][y] == 1:
+                p_curr[x][y] = pressure_s(t)
+                continue
+                
+            # if x, y is not a wall
+            if pierce[x][y] == 0:
+                up = p_prev[x][y-1]
+                right = p_prev[x+1][y]
+                down = p_prev[x][y+1]
+                left = p_prev[x-1][y]
+            
+                # if neighbors are walls, ghost nodes
+                if pierce[x][y-1] == 1:
+                    up = p_prev[x][y]
+                if pierce[x+1][y] == 1:
+                    right = p_prev[x][y]
+                if pierce[x][y+1] == 1:
[…]

### 2014-12-10 paper.tex
minor changes to paper

@@
 
-The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
+The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$.
+In order for there to be a speedup relative to the fine operator, we require
 that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q

### 2014-12-10 paper.tex
temp push to sync

@@
 
+\subsubsection{Parareal}
+
 We look to compute a theoretical maximum speedup enabled by the nature of the Parareal algorithm.
@@
 
-Then, the runtime of Parareal is, assuming negligible setup and
-aggregation time,
-
-TODO WESLEY
+Then, the runtime of Parareal is, assuming negligible overhead and ignoring communication speeds and letting $t$ be the time in seconds to compute the coarse $g_{\Delta t}$:
 
@@
 
-The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$.
-In order for there to be a speedup relative to the fine operator, we require
-that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q
-- 1 - k}$.
-This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, k, is important for error convergence to that of $g_{\textnormal{fine}}$.
-However, this optimal k is a problem-specific parameter and must be optimized for.
+The first $t$ seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$.
+Then for each of the $k$ correction iterations, the $t$ term comes from $n$, element by element evaluations of  $g_{\Delta t}$ followed by a parallel $\frac{1}{N}$ evaluation of $g_{\textnormal{fine}}$.
+In order for there to be a speedup relative to the fine operator, we can rearrange to get:
 
-TODO PARALLEL IN SPACE ANALYSIS
+\begin{equation}
+\begin{aligned}
+t + k\left( t + \frac{Qt}{N} \right) &< Qt \\
+k &< \frac{Q - 1}{1 + Q/N} \\
+N &> \frac{Qk}{Q - 1 - k} \\
+\end{aligned}
+\end{equation}
+
+This can be tuned with by either adjusting the the quality factor or reducing the number of iterations, but the number of iterations, $k$, is important for error convergence to that of $g_{\textnormal{fine}}$.
+However, finding the optimal $k$ is a problem-specific parameter and traded off for the desired accepted error.
+
+\subsubsection{Parallel By Space}
 
@@
 
-TODO WESLEY
+\subsubsection{Parareal}
 
-TODO PARALLEL IN SPACE ANALYSIS
+\subsubsection{Parallel By Space}
 

### 2014-12-10 paper.tex
4:30 push

@@
 
+\subsubsection{Speedup}
+
+Speedup, $S$ is defined simply as the ratio of the serial implementation to the parallelized version with a given number of processors.
+These numbers are usually reported per number of processors that are tested:
+
+\begin{equation}
+S = \frac{T_\textnormal{serial}}{T_\textnormal{parallel}}
+\end{equation}
+
+\subsubsection{Strong Scaling Efficiency}
+
+Strong scaling 
+Programs that are difficult to parallelize will have efficiency expressions like $\frac{1}{\ln p}$ which approaches zero as the number of processors increases by a large amount.
+
+\begin{equation}
+\textnormal{Eff}_{\textnormal{strong}} = \frac{t_1}{N * t_N} * 100\%
+\end{equation}
+
+\subsubsection{Weak Scaling Efficiency}
+
+\begin{equation}
+\textnormal{Eff}_{\textnormal{weak}} = \frac{t_1}{t_N} * 100\%
+\end{equation}
+
 \section{Methodology}
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}
     \KwIn{Temporal discretization $t_n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
@@
 We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.
-
-\subsection{Visualization of Parareal Scheme}
-
-bksim TODO put a graph because maybe it'll help? the slides have a version
 
@@
 
-\begin{equation}
+\[
 \begin{aligned}
@@
 \end{aligned}
-\end{equation}
+\]
 
@@
 
+For the parallelism by space, we are dividing our grid on which we which to solve our wave diffusion equation.
+
 \subsection{Incorporating Communication}
+
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$
 
@@
 
+The Parareal method TODO does this scale well?
+
 \subsubsection{Parallel By Space}
+
+Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of
+
+\[
+est
+\]
 

### 2014-12-10 paper.tex
parareal picture

--- 2014-12-10 paper.tex
@@
 \usepackage[margin=1in]{geometry}
+% force pictures to go where you want
+\usepackage{float}
+
 \geometry{letterpaper}
@@
 \begin{abstract}
-
 We look to achieve strong scaling in parallelizing numerical ODE methods.
@@
 
-bksim TODO put a graph because maybe it'll help? the slides have a version
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=0.75\linewidth]{parareal_visualization.png}
+\caption{Graphical representation of the parareal scheme.}
+\label{fig:pararealvis}
+\end{center}
+\end{figure}
 
@@
 
-The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$.
-In order for there to be a speedup relative to the fine operator, we require
+The first t seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$,  In order for there to be a speedup relative to the fine operator, we require
 that $t + k(t + Qt/N) < Qt$, or $k < \frac{Q - 1}{1 + Q/N}$ or $N > \frac{Qk}{Q

### 2014-12-10 serial_integrate.py

@@ -72,7 +72,7 @@ if __name__ == "__main__":
     nsteps_lst = [200, 400, 800, 1600, 3200, 6400, 12800, 25600, 51200]
     init = np.array([1., 0.])
     tmin = 0.0
-    tmax = 20.
+    tmax = 30.
     #print("type,nsteps,err")
     with open('out', 'wb') as f:
         f.write('type,nsteps,err\n')

### 2014-12-10 paper.tex
small change in paper typo

--- 2014-12-10 paper.tex
@@
 \begin{abstract}
-
 We look to achieve strong scaling in parallelizing numerical ODE methods.
 We test both parallelism by result space as well as parallelism by time, with the Parareal algorithm, to solve various ODEs.
-We measure performance and compare in
-terms of accuracy, speedup and efficiency. We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
+We measure performance and compare in terms of accuracy, speedup and efficiency.
+We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
 light speedups in our small scale tests up to 64 processors on Harvard's
@@
 We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.
+
+\subsection{Visualization of Parareal Scheme}
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=0.75\linewidth]{parareal_visualization.png}
+\caption{Graphical representation of the parareal scheme.}
+\label{fig:pararealvis}
+\end{center}
+\end{figure}
 
@@
 For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute.
-We wish to strong scaling, in which the computation time of our parallel method will decrease as we increase the number of processors.
+We wish to show strong scaling, in which the computation time of our parallel method will decrease as we increase the number of processors.
 We compare this version to the serial version we did for Homework 4.

### 2014-12-10 paper.tex
5:30 push

@@
 Performance of parallel algorithms are usually measured in terms of 3 different metrics: speedup, strong scalling efficiency and weak scaling efficiency.
+Strong and weak scaling are common measures for the different ways a program can benefit from increased parallellism and more processors as well as the limitatios that dictate the speedups.
 
@@
 
-Strong scaling 
+Strong scaling is a way to measure performance of a given application depending on if it is CPU-bound. This means that 
 Programs that are difficult to parallelize will have efficiency expressions like $\frac{1}{\ln p}$ which approaches zero as the number of processors increases by a large amount.
@@
 \end{equation}
+
+\subsection{Amdahl's Law}
+
+The final principle when discussing parallel performance is the potential existence of an Amdahl speedup limit.  o7777777777
 

### 2014-12-10 paper.tex
6:30 push

@@
 Performance of parallel algorithms are usually measured in terms of 3 different metrics: speedup, strong scalling efficiency and weak scaling efficiency.
-Strong and weak scaling are common measures for the different ways a program can benefit from increased parallellism and more processors as well as the limitatios that dictate the speedups.
+Strong and weak scaling are common measures for the different ways a program can benefit from increased parallellism and more processors as well as the limitations that dictate the speedups.
+In each of the below subsections, $T(n)$ represents the computation of using $n$ processors and the same notation for speedup.
 
@@
 \begin{equation}
-S = \frac{T_\textnormal{serial}}{T_\textnormal{parallel}}
+S(n) = \frac{T_\textnormal{serial}}{T_\textnormal{parallel in n cores}}
 \end{equation}
@@
 
-Strong scaling is a way to measure performance of a given application depending on if it is CPU-bound. This means that 
-Programs that are difficult to parallelize will have efficiency expressions like $\frac{1}{\ln p}$ which approaches zero as the number of processors increases by a large amount.
+Strong scaling is a way to measure performance of a given application depending on if it is CPU-bound.
+This means that given a fixed problem size, with more processors, the computation time will decrease for each.
+Programs that are difficult to parallelize will have efficiency expressions like $\frac{1}{\ln p}$ which approaches zero as the number of processors increases by a large amount meaning that adding additional processors does not give much more speedup.
+We can compute the strong scaling efficiency with the follow expression:
 
 \begin{equation}
-\textnormal{Eff}_{\textnormal{strong}} = \frac{t_1}{N * t_N} * 100\%
+\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n * T(n)} * 100\%
 \end{equation}
@@
 
+Weak scaling measures the memory-bound nature of the parallelized application.
+A weak scaling parallel implementation is one that allows a larger problem proportionally increased for each processor to be solved in the same time as the original.
+This is usually the more intuitive usage and measure to quantify parallel application performance.
+Below is the computation for weak scaling efficiency:
+
 \begin{equation}
-\textnormal{Eff}_{\textnormal{weak}} = \frac{t_1}{t_N} * 100\%
+\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} * 100\%
 \end{equation}
@@
 
-The final principle when discussing parallel performance is the potential existence of an Amdahl speedup limit.  o7777777777
+The final principle when discussing parallel performance is the potential existence of an Amdahl speedup limit.
+Amdahl's Law seeks to describe the maximum parallel speedup in an overall piece of code when the problem is not embarassingly parallel --- not all of the system can be divided up in parallel.
+The parallel in time methods must always have some serial component due to the time-dependent nature of the ODE's which will fall under the limitations stated by Amdahl's Law.
+His law expresses both best time given $n$ processors as well as the speedup for $n$ processors as a function of $B$, the fractional amount of the algorithm that can be parallelized:
+
+\begin{equation}
+T(n) = T(1) \left( B + \frac{1}{n} (1-B) \right)
+\end{equation}
+
+\begin{equation}
+S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
+\end{equation}
 
@@
 
+The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators.
+The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.
+Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.
+The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code.
+We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.
+
+\subsection{Parareal Algorithm and Visualization}
+
 \begin{algorithm}
@@
 \end{algorithm}
-
-The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators.
-The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.
-Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.
-The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code.
-We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.
-
-\subsection{Visualization of Parareal Scheme}
 
@@
 As this is a tradeoff between error (with parameter k being the number of iterations of corrections, each which takes another cycle of $g_{\Delta t}$, we seek to express the tradeoff as a function of the quality factor, $Q$, the number of iterations, $k$ and the number of processors $n$.
-Our quality factor is defined as the multiplicative factor of additional time necessary to
+Our quality factor is defined as the multiplicative factor of additional time necessary to TODO
 
@@
 For the parallelism by space, we are dividing our grid on which we which to solve our wave diffusion equation.
+This division is inherently parallelizable and allows us to follow a very similar approach in computing this as in the serial version - just divided in space, where the stencil approach is well-defined.
 
@@
 
+One of the main drawbacks of using Amdahl's Law as benchmarks in analysis is in the generally abstract way the paralleizable fraction is determined, the $B$ term.
+The value of this is estimated based on an understanding of the algorithm from a high level and not derived from any underlying mathematical expressions.
+
 \subsubsection{Parareal}
@@
 
-Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of
+Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound (with $B$ of 1):
 
@@
 \subsection{Explored Tests}
+
+WESLEY ENDED HERE, BEFORE THIS IS WESLEY APPROVED
 

### 2014-12-10 paper.tex
almost done up until results sections

@@
 The parallel in time methods must always have some serial component due to the time-dependent nature of the ODE's which will fall under the limitations stated by Amdahl's Law.
-His law expresses both best time given $n$ processors as well as the speedup for $n$ processors as a function of $B$, the fractional amount of the algorithm that can be parallelized:
+His law expresses both best time given $n$ processors as well as the speedup for $n$ processors as a function of $B$, the fractional amount of the algorithm that cannot be parallelized---due to strict serial constraints:
 
@@
 As this is a tradeoff between error (with parameter k being the number of iterations of corrections, each which takes another cycle of $g_{\Delta t}$, we seek to express the tradeoff as a function of the quality factor, $Q$, the number of iterations, $k$ and the number of processors $n$.
-Our quality factor is defined as the multiplicative factor of additional time necessary to TODO
+Our quality factor is defined as the multiplicative factor of additional time necessary to compute the solution using the $g_{\textnormal{fine}}$ scheme instead of the $g_{\Delta t}$ method.
 
-Then, the runtime of Parareal is, assuming negligible overhead and ignoring communication speeds and letting $t$ be the time in seconds to compute the coarse $g_{\Delta t}$:
+Then, the runtime of Parareal is, assuming negligible overhead and ignoring communication bandwidth and letting $t$ be the time in seconds to compute the coarse $g_{\Delta t}$:
 
@@
 
-Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound (with $B$ of 1):
+Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):
 
 \[
-est
+S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
 \]
+
+With embarassingly parallel problem setups, we can optimally achieve up to $n$ times speedup with $n$ processors though this number if quite ideal and will be very difficult to approach (large scaling will have different memory management and greater communication overheads and costs).
 
@@
 
-WESLEY ENDED HERE, BEFORE THIS IS WESLEY APPROVED
+We wished to explore certain tests that would demonstrate the effects of changing various parameters be it from the solution schemes in the parareal to the exact system we were trying to solve.
 
-We wished to explore certain tests that would span the space of possible
-parameters to the parareal as well as possible applications in different ODE
-systems. Because there were so many different versions and parameters for the Parareal to test,
+Due to the sheer number of things to explore, we usually try to compare each run to a benchmark we decided was most useful for comparison's sake and the list of parameters we wished to test are:
+
+\begin{itemize}
+\item
+\item
+\end{itemize}
 
@@
 
+WESLEY ENDED HERE, BEFORE THIS IS WESLEY APPROVED
+
 \subsection{Expectations}
-TODO WESLEY
+We hoped to show that under the properly tuned parameters (namely the $k$ number of iterations),
 

### 2014-12-10 paper.tex
fixed typos in paper

--- 2014-12-10 paper.tex
@@
 refinement built on top of a more basic method first computed in serial. As will be discussed and analyzed in the
-metodology section, one such scheme would be the parareal algorithm. As an
+methodology section, one such scheme would be the parareal algorithm. As an
 overview, the parareal algorithm first does a coarse (low-order and fast)
@@
 \begin{equation}
-\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n * T(n)} * 100\%
+\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n \times T(n)} \times 100\%
 \end{equation}
@@
 \begin{equation}
-\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} * 100\%
+\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} \times 100\%
 \end{equation}
@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$
 

### 2014-12-10 paper.tex
minor modifications on paper

--- 2014-12-10 paper.tex
@@
 refinement built on top of a more basic method first computed in serial. As will be discussed and analyzed in the
-methodology section, one such scheme would be the parareal algorithm. As an
+metodology section, one such scheme would be the parareal algorithm. As an
 overview, the parareal algorithm first does a coarse (low-order and fast)
@@
 \begin{equation}
-\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n \times T(n)} \times 100\%
+\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n \cdot T(n)} \cdot 100\%
 \end{equation}
@@
 \begin{equation}
-\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} \times 100\%
+\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} \cdot 100\%
 \end{equation}
@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$
 
@@
 Our coarse operator was forward Euler step, with 100 steps, while the fine
-operator was forward Euler, with $Q * 100$ steps, where $Q$ is the quality
+operator was forward Euler, with $Q \cdot 100$ steps, where $Q$ is the quality
 factor. We tested the Parareal algorithm on two sets of differential equations:

### 2014-12-10 paper.tex
adding bibliography

@@
 
-The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
-a generalized algorithm which allows for parallelization in time. It does so
-by using a cheaper ($g_{\Delta t}$)---a lower order or lesser resolution
-approximation first, which it then corrects by evaluating segments with a finer
-method $g_{\textnormal{fine}}$. The correction step from the approximation can
-be done in parallel. Multiple iterations can be performed to achieve an error
-provably asymptotically equal to a full serial computation of
-$g_{\textnormal{fine}}$. The exact number of iterations required to achieve
-certain levels of convergence depends on the problem and must be tuned top
-optimize for speedup for a given system. In summary, the algorithm is $k$
-repetitions of a correction process which requires running updated by a finer
-method run in parallel for subsections.
+The parareal algorithm, developed by Lions, Maday and Turinici in 2001
+\cite{lions2001parareal}, is a generalized algorithm which allows for
+parallelization in time. It does so by using a cheaper ($g_{\Delta t}$)---a
+lower order or lesser resolution approximation first, which it then corrects by
+evaluating segments with a finer method $g_{\textnormal{fine}}$. The correction
+step from the approximation can be done in parallel. Multiple iterations can be
+performed to achieve an error provably asymptotically equal to a full serial
+computation of $g_{\textnormal{fine}}$. The exact number of iterations required
+to achieve certain levels of convergence depends on the problem and must be
+tuned top optimize for speedup for a given system. In summary, the algorithm is
+$k$ repetitions of a correction process which requires running updated by a
+finer method run in parallel for subsections.
 
@@
 
-\section{References}
+\bibliographystyle{unsrt}
+\bibliography{paper}
 
 Bibtex Citations for the Slides, and for \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
-
-%\bibliographystyle{unsrt}
-%\bibliography{poster/sample}
 

### 2014-12-10 paper.tex
added biblio

@@
 \usepackage[linesnumbered,vlined,ruled]{algorithm2e} % algorithm
+\usepackage{microtype}
 
@@
 
-Performance of parallel algorithms are usually measured in terms of 3 different metrics: speedup, strong scalling efficiency and weak scaling efficiency.
-Strong and weak scaling are common measures for the different ways a program can benefit from increased parallellism and more processors as well as the limitations that dictate the speedups.
-In each of the below subsections, $T(n)$ represents the computation of using $n$ processors and the same notation for speedup.
+Performance of parallel algorithms are usually measured in terms of 3 different
+metrics: speedup, strong scalling efficiency and weak scaling efficiency.
+Strong and weak scaling are common measures for the different ways a program can
+benefit from increased parallellism and more processors as well as the
+limitations that dictate the speedups~\cite{parallelscaling}. In each of the below subsections, $T(n)$ represents the computation of using $n$ processors and the same notation for speedup.
 
@@
 
-The parareal algorithm, developed by Lions, Maday and Turinici in 2001
-\cite{lions2001parareal}, is a generalized algorithm which allows for
+The parareal algorithm, developed by Lions, Maday and Turinici in
+2001~\cite{lions2001parareal}, is a generalized algorithm which allows for
 parallelization in time. It does so by using a cheaper ($g_{\Delta t}$)---a
@@
 solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
-this using induction [Bal].
+this using induction \cite{bal2005convergence}.
 
@@
 u_n) = \bar{g_{\textnormal{fine}}}u_n$ and $g_{\Delta t}(t_n, u_n) =
-\bar{g_{\Delta t}} u_n$. As shown by Staff et al. [cite], the parareal method
-becomes
+\bar{g_{\Delta t}} u_n$. As shown by Staff et al. \cite{staff2005stability}, the parareal method becomes
 
@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$
 
@@
 Varying Iterations:
-]
+
 Varying Quality Factor:
@@
 
+\nocite{*}
 \bibliographystyle{unsrt}
@@
 
-Guilliame Bal paper
-
-Stability of the Parareal Algorithm by Staff et al
-
 \end{document}

### 2014-12-10 paper.tex
pushing lease no conflicts

--- 2014-12-10 paper.tex
@@
 \begin{equation}
-\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n \times T(n)} \times 100\%
+\textnormal{Eff}_{\textnormal{strong}} = \frac{T(1)}{n * T(n)} * 100\%
 \end{equation}
@@
 \begin{equation}
-\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} \times 100\%
+\textnormal{Eff}_{\textnormal{weak}} = \frac{T(1)}{T(n)} * 100\%
 \end{equation}
@@
 t + k \left(t + \frac{Qt}{N} \right).
+\label{eq:scheme}
 \end{equation}
@@
 This division is inherently parallelizable and allows us to follow a very similar approach in computing this as in the serial version - just divided in space, where the stencil approach is well-defined.
+A fully parallelizable approach allows for $n$ times speedup and no computational tradeoffs if we ignore communication.
+As we are dividing our processors into a processor grid (dividing a 2D space into a 2D arrangement of processors) we are only limited by when the processor grid size approaches the problem size - upon which further processors would not receive any tasks since that would require a division by less than a unit of computation.
 
@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$ explained in the below table.
+
+\begin{table}
+\centering
+\begin{tabular}{|c|l|}
+{\bf Variable} & {\bf Description}\\
+\hline\hline
+$\alpha$ & Latency of communication. (sec)\\
+$\beta$ & Inverse communication bandwidth. (sec/byte)\\
+$\gamma$ & Inverse computational performance. (sec/flop)\\
+\hline
+\end{tabular}
+\caption{Table of Communication Architecture Parameters}
+\label{tab:notation}
+\end{table}
 
 \subsubsection{Parareal}
+TODO THIS IS CONFUSING Wesley
 
 \subsubsection{Parallel By Space}
+TODO THIS IS CONFUSING TOO Wesley
 
@@
 
-The Parareal method TODO does this scale well?
+We first try to express $B$ as function of the quality factor $Q$ and the number of correction steps of the parareal $k$.
+
+From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + K(t+ Qt)$.
+Of this, the only parallelizable part is the kQt term.
+Thus to find the expression of the parallelizable portion, we simplify:
+
+\[
+
+\]
+
+Now to replace this into Amdahl's Law, we have:
+
+\[
+S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
+\]
 
@@
 
-\subsection{Explored Tests}
+\subsection{Planned Tests to Run}
 
@@
 \begin{itemize}
-\item
-\item
+\item General Scaling with Incrased $N$
+\item Adjusting $K$ in the Parareal
+\item Using Different Integration Schemes in Parareal for $g_{\textnormal{fine}}$
+\item Run the Solver on Various Complexities of ODEs
 \end{itemize}
@@
 Our coarse operator was forward Euler step, with 100 steps, while the fine
-operator was forward Euler, with $Q \cdot 100$ steps, where $Q$ is the quality
+operator was forward Euler, with $Q * 100$ steps, where $Q$ is the quality
 factor. We tested the Parareal algorithm on two sets of differential equations:

### 2014-12-10 paper.tex

--- 2014-12-10 paper.tex
@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$ explained in the below table.
+
+\begin{table}
+\centering
+\begin{tabular}{|c|l|}
+{\bf Variable} & {\bf Description}\\
+\hline\hline
+$\alpha$ & Latency of communication. (sec)\\
+$\beta$ & Inverse communication bandwidth. (sec/byte)\\
+$\gamma$ & Inverse computational performance. (sec/flop)\\
+\hline
+\end{tabular}
+\caption{Table of Communication Architecture Parameters}
+\label{tab:notation}
+\end{table}
 
 \subsubsection{Parareal}
+TODO THIS IS CONFUSING Wesley
 
 \subsubsection{Parallel By Space}
+TODO THIS IS CONFUSING TOO Wesley
 
@@
 
-The Parareal method TODO does this scale well?
+We first try to express $B$ as function of the quality factor $Q$ and the number of correction steps of the parareal $k$.
+
+From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + K(t+ Qt)$.
+Of this, the only parallelizable part is the $kQt$ term.
+Thus to find the expression of the parallelizable portion, we simplify:
+
+\[
+AA
+\]
+
+Now to replace this into Amdahl's Law, we have:
+
+\[
+S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
+\]
 

### 2014-12-10 paper.tex
paper updates

@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$ explained in the below table.
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$ explained in the below table.
 
@@
 
-For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute.
-We wish to show strong scaling, in which the computation time of our parallel method will decrease as we increase the number of processors.
-We compare this version to the serial version we did for Homework 4.
+To demonstrate parallelism by space, we solve the two-dimensional wave equation by partioning a two-dimensional space domain amongst multiple processors. Specifically, we look to numerically solve
 
-bksim TODO GET DATA, AS REQUIRED FROM ABOVE, SHOW GRAPHS
+\begin{equation*}
+\frac{\partial^2 u}{\partial t^2} = \frac{\partial^2 u}{\partial x^2} + \frac{\partial^2 u}{\partial y^2}
+\end{equation*}
+
+on the two-dimensional domain $(x, y) \in [0,1]\times[0,1]$ for evolution in time.
+We partition the domain by space using a second order central difference approximation, assigning a rectangular portion of the two-dimensional domain to each of a grid of $(P_x, P_y)$ processors.
+This requires the passing of information along the edges of the processors to each other, as shown graphically in the figure below:
+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=0.5\linewidth]{decomp.png}
+\caption{Parallelization by space schematic.}
+\label{fig:spaceparallel}
+\end{center}
+\end{figure}
+
+Note that each processor receives information from only those adjacent to it; that is, the top left processor in the graphic does not communicate directly with the bottom right processor in the graphic.
+Rather, information is passed only along shared boundary points to propagate the wave equation.
+We run this code for two different grid sizes, $128\times 128$ and $256\times 256$, as well as in serial.
+
+In serial, we have:
+
+\begin{center}
+\begin{tabular}{|c|c|} \hline
+($N_x, N_y$) & Time (seconds)\\ \hline
+(128, 128) & 0.987\\ \hline
+(256, 256) & 7.889\\ \hline
+\end{tabular}
+\end{center}
+
+In parallel, we have:
+
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|} \hline
+($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
+(128, 128) & 2 & 2 &  \\ \hline
+(128, 128) & 4 & 2 & 2.31 \\ \hline
+(128, 128) & 4 & 4 &  \\ \hline
+(128, 128) & 8 & 8 &  \\ \hline
+(256, 256) & 2 & 2 & 2.57 \\ \hline
+(256, 256) & 4 & 2 & 11.32 \\ \hline
+(256, 256) & 4 & 4 & 13.65 \\ \hline
+(256, 256) & 8 & 8 & 17.44 \\ \hline
+\end{tabular}
+\end{center}
+
+Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, linear scaling is achieved.
+However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed).
+Hence, this is an example of weak scaling.
+However, an advantage to this paradigm is that because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes.
+Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.
 
@@
 
-As we can see, we do see the effects of strong scaling
+As we can see, we do see the effects of WEAK scaling
 
@@
 
-On the other hand, parallelism by space already requires a very problem-specific
-design. The strong scaling efficiency of parallelism by space is high and is
-conceptually embarrassingly parallel and easier to create. However, the
+On the other hand, parallelism by space already requires a more problem-specific
+design. Parallelism by space is intuitive, and conceptually, is embarrassingly parallel and easier to create. However, the
 limitations are in the requirement for the code to be designed for the

### 2014-12-10 parallel_space/parallel_wave512.py

@@ -111,7 +111,7 @@ if __name__ == '__main__':
 # Global constants
   xMin, xMax = 0.0, 1.0     # Domain boundaries
   yMin, yMax = 0.0, 1.0     # Domain boundaries
-  Nx = 256                   # Number of total grid points in x
+  Nx = 512                   # Number of total grid points in x
   Ny = Nx                   # Number of total grid points in y
   dx = (xMax-xMin)/(Nx-1)   # Grid spacing, Delta x
   dy = (yMax-yMin)/(Ny-1)   # Grid spacing, Delta y

### 2014-12-10 parallel_space/runscript32x1.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --ntasks 32                #Number of processes
+#SBATCH --nodes 1                  #Number of nodes
+#SBATCH -t 60                 #Runtime in minutes
+#SBATCH -p general            #Partition to submit to
+
+#SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
+#SBATCH -o algo_32x1procs.out    #File to which standard out will be written
+#SBATCH -e algo_32x1procs.err    #File to which standard err will be written
+
+# load modules
+module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
+module load centos6/numpy-1.7.1_python-2.7.3
+
+mpirun -n 32 python parallel_wave.py 32 1

### 2014-12-10 parallel_space/runscript4x2.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH --ntasks 8                #Number of processes
+#SBATCH --nodes 1                  #Number of nodes
+#SBATCH -t 60                 #Runtime in minutes
+#SBATCH -p general            #Partition to submit to
+
+#SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
+#SBATCH -o algo_4x2procs.out    #File to which standard out will be written
+#SBATCH -e algo_4x2procs.err    #File to which standard err will be written
+
+# load modules
+module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
+module load centos6/numpy-1.7.1_python-2.7.3
+
+mpirun -n 8 python parallel_wave.py 4 2

### 2014-12-10 parallel_space/runscript4x2_512.sh

@@ -1,16 +1,15 @@
 #!/bin/bash
-#SBATCH --ntasks 2                #Number of processes
+#SBATCH --ntasks 8                #Number of processes
 #SBATCH --nodes 1                  #Number of nodes
 #SBATCH -t 60                 #Runtime in minutes
 #SBATCH -p general            #Partition to submit to
 
 #SBATCH --mem-per-cpu=200     #Memory per cpu in MB (see also --mem)
-#SBATCH -o algo_2procs.out    #File to which standard out will be written
-#SBATCH -e algo_2procs.err    #File to which standard err will be written
+#SBATCH -o algo512_4x2procs.out    #File to which standard out will be written
+#SBATCH -e algo512_4x2procs.err    #File to which standard err will be written
 
 # load modules
 module load centos6/mpi4py-1.3.1_python-2.7.3_openmpi-1.6.4_gcc-4.8.0
 module load centos6/numpy-1.7.1_python-2.7.3
-module load centos6/Anaconda-1.8.0
 
-mpirun -n 4 python P2A.py 2 2
+mpirun -n 8 python parallel_wave512.py 4 2

### 2014-12-10 paper.tex
paper changes
trying to get changes

@@
 
-Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):
+Because this task is embarassingly parallel and nearly 100\% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):
 
@@
 
-Varying Iterations:
+%Varying Iterations:
 
 Varying Quality Factor:
-We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$
-
-Changing the ODE
-
-TODO ashi sort out the figure and match and replace if needbe
+We define the quality factor in this case for how many times smaller the Euler
+method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for
+$g_{\Delta t}$.
 
@@
 \end{figure}
+
+Figure~\ref{fig:run_v_cpu} shows the running time (wall clock) for the Parareal
+algorithm
 
@@
 Rather, information is passed only along shared boundary points to propagate the wave equation.
-We run this code for two different grid sizes, $128\times 128$ and $256\times 256$, as well as in serial.
+We run this code for three different grid sizes: $128\times 128$, $256\times 256$, and $512\times 512$, in both serial and parallel.
 
@@
 (256, 256) & 7.889\\ \hline
+(512, 512) & 63.913\\ \hline
 \end{tabular}
@@
 (256, 256) & 8 & 8 & 17.44 \\ \hline
+(512, 512) & 2 & 2 & \\ \hline
+(512, 512) & 4 & 2 &  \\ \hline
+(512, 512) & 4 & 4 &  \\ \hline
+(512, 512) & 8 & 8 &  \\ \hline
 \end{tabular}

### 2014-12-10 paper.tex

@@
 \begin{equation}
-S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
+S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
 \end{equation}
@@
 Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.
-The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code.
+The exact optimum pairing is problem-specific and the code is written so that a different ``n-order method step'' function could be written and then replaced into the code.
 We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.
@@
 solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
-this using induction \cite{bal2005convergence}.
+this using induction~\cite{bal2005convergence}.
 
@@
 of accuracy approaches that of $g_{\textnormal{fine}}$.
-
-% I don't understand the above part ~ Andy
 
@@
 
-From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + K(t+ Qt)$.
+From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + k(t+ Qt)$.
 Of this, the only parallelizable part is the $kQt$ term.
-Thus to find the expression of the parallelizable portion, we simplify:
+Thus to find the expression of the non-parallelizable portion, we simplify:
 
 \[
-AA
+B = \frac{t+kt}{t+kt+kQt} = \frac{1+k}{kQ+k+1}
 \]
@@
 \[
-S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
+\begin{aligned}
+S(n) &= \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{1}{\left( B + \frac{1}{n} (1-B) \right)}\\
+       &= \frac{T(1)}{T(1) \left( \frac{1+k}{kQ+k+1} + \frac{1}{n} (1- \frac{1+k}{kQ+k+1}) \right)} \\
+       &= \frac{n(kQ+k+1)}{n+nk+kQ}
+\end{aligned}
 \]
+
+As written below, we have used $Q = 100$ with various $k$.
+
+When $k = 2$ and $Q = 100$, for maximal speedup (fewest iterations) of $S(n) = \frac{203n}{3n+200} \approx 67$
+
+When $k = 10$ and $Q = 100$, for more iterations and less error, we'd see an Amdahl bounded max speedup of $S(n) = \frac{1011n}{11n+1000} \approx 91$
+
+It is important to note that this speedup is measured with respect to the parareal agorithm done in parallel.
+This does not compare the Parareal to doing $g_{\textnormal{fine}}$ but rather shows how given infinite processors, our speedup will approach certain asymptotic bounds.
+We do see however large constants in the denominator which means that we will get better speedup as we increase n past a certain size.
 
@@
 \[
-S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
+S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
 \]
@@
 (128, 128) & 0.987\\ \hline
-(256, 256) & 7.889\\ \hline
-(512, 512) & 63.913\\ \hline
+(256, 256) & 7.889\\ \hline %real
+(512, 512) & 63.913\\ \hline %real
 \end{tabular}
@@
 ($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
-(128, 128) & 2 & 2 &  \\ \hline
+(128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
 (128, 128) & 4 & 2 & 2.31 \\ \hline
-(128, 128) & 4 & 4 &  \\ \hline
-(128, 128) & 8 & 8 &  \\ \hline
-(256, 256) & 2 & 2 & 2.57 \\ \hline
+(128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
+(128, 128) & 8 & 8 & (still queued on Odyssey) \\ \hline
+(256, 256) & 2 & 2 & 2.57 \\ \hline %real
 (256, 256) & 4 & 2 & 11.32 \\ \hline
@@
 (256, 256) & 8 & 8 & 17.44 \\ \hline
-(512, 512) & 2 & 2 & \\ \hline
-(512, 512) & 4 & 2 &  \\ \hline
-(512, 512) & 4 & 4 &  \\ \hline
-(512, 512) & 8 & 8 &  \\ \hline
+(512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
+(512, 512) & 4 & 2 & (still queued on Odyssey)\\ \hline
+(512, 512) & 4 & 4 & (still queued on Odyssey)\\ \hline %check
+(512, 512) & 8 & 8 & (still queued on Odyssey) \\ \hline %check
 \end{tabular}
@@
 
-Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, linear scaling is achieved.
+Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved.
 However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed).
@@
 
-\subsection{Parallelism Paradigm Comparisons}
-
-Now with 3 different parallel implementations, two in time with the parareal, one in space, we compare across the methods to see if we can make any statements about general performance of our different paradigms.
-
-TODO WESLEY
-
-TODO compare the computation times (and perhaps a slight change in accuracy of the results) between both parareal methods the best in the sections before hand
-
-TODO bsim and ashi show graphs and some times for the forward Euler parareal
-versions---the fastest ones or both if we have both implemented
-
-\subsection{Comparison to C++ Implementations}
-
-The numpy environment combined with the required structure of MPI required
-certain type/data structure conversions which will take time.not not cunot cu
-
 \subsection{Summary of Performance}
@@
 
-\subsection{Demonstrating Scaling}
+\subsection{Demonstrating Scaling and Efficiency}
 
-ashi TODO by taking some numbers and following
+\subsection{Parallel in Space}
+
+bsim TODO eventually
 
@@
 
+TODO NUMPY WRITE FATST HING
+
 On the other hand, parallelism by space already requires a more problem-specific
@@
 
-TODO Wesley
+TODO WRITE ABOUT BETTER DATA COLLECTION
 

### 2014-12-10 paper.tex
paper updates

--- 2014-12-10 paper.tex
@@
 (128, 128) & 0.987\\ \hline
-(256, 256) & 7.889\\ \hline
+(256, 256) & 7.889\\ \hline %real
+(512, 512) & 63.913\\ \hline %real
 \end{tabular}
@@
 (256, 256) & 8 & 8 & 17.44 \\ \hline
+(512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
+(512, 512) & 4 & 2 & (still queued on Odyssey)\\ \hline
+(512, 512) & 4 & 4 & (still queued on Odyssey)\\ \hline %check
+(512, 512) & 8 & 8 & (still queued on Odyssey) \\ \hline %check
 \end{tabular}

### 2014-12-10 paper.tex
wrote some stuff

--- 2014-12-10 paper.tex
@@
 \usepackage{amsmath} % Needed for math stuff
-%\usepackage{lastpage}sss% Finds last page
+%\usepackage{lastpage} % Finds last page
 \usepackage{amssymb}  % Needed for some math symbols
-\usepackage{graphicx}% Needed for graphics
+\usepackage{graphicx} % Needed for graphics
 \usepackage[usenames,dvipsnames]{xcolor} % Needed for graphics and color
-%\usepackage{setspace}sss% Needed to set single/double spacing
+%\usepackage{setspace} % Needed to set single/double spacing
 \usepackage{float} % Better placement of figures & tables
@@
 \usepackage[margin=1in]{geometry}
-% force pictures to go where you want
-\usepackage{float}
-
 \geometry{letterpaper}
-%\addtolength{\oddsidemargin}{-0.875in} 
-%\addtolength{\evensidemargin}{-0.875in} 
-%\addtolength{\textwidth}{1.75in} 
-%\addtolength{\topmargin}{-.875in} 
-%\addtolength{\textheight}{1.75in}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[htb]
 \begin{center}
@@
 
+Example 1:
 \[
@@
 \]
+
+Example 2:
 \[
@@
 We ran the algorithm on different numbers of processors and varied $k$ and $Q$.
+We define the quality factor $Q$ in this case for how many times smaller the
+Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method
+used for $g_{\Delta t}$.
 
-%Varying Iterations:
-
-Varying Quality Factor:
-We define the quality factor in this case for how many times smaller the Euler
-method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for
-$g_{\Delta t}$.
-
-\begin{figure}
+\begin{figure}[htb]
 \begin{center}
@@
 Figure~\ref{fig:run_v_cpu} shows the running time (wall clock) for the Parareal
-algorithm
+algorithm compared to different number of CPUs used. The serial version is the
+Euler method with $Q$ times as many steps. As we can see, the running time
+decreases as more processors are added. However, there is a base time associated
+with setting up the parallel infrastructure, so asymptotically the running time
+is nonzero.
 
-\begin{figure}
+\begin{figure}[htb]
 \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[htb]
 \begin{center}

### 2014-12-10 paper.tex
please no more conflicts please
Conflicts:
	paper.tex

--- 2014-12-10 paper.tex
@@
 Most of the time this will be a higher order method, but can also be a greater sampling of the same order method used in the coarse version.
-The exact optimum pairing is problem-specific and the code is written so that a different ``n-order method step'' function could be written and then replaced into the code.
+The exact optimum pairing is problem-specific and the code is written so that a different "n-order method step" function could be written and then replaced into the code.
 We experiment with testing different pairs of $g_{\Delta t}$ and $g_{\textnormal{fine}}$ to see how influential the choice of solution scheme is.
@@
 solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
-this using induction~\cite{bal2005convergence}.
+this using induction \cite{bal2005convergence}.
 
@@
 of accuracy approaches that of $g_{\textnormal{fine}}$.
+
+% I don't understand the above part ~ Andy
 
@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$ explained in the below table.
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $gamma$ explained in the below table.
 
@@
 
-Because this task is embarassingly parallel and nearly 100\% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):
+Because this task is embarassingly parallel and nearly 100% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):
 
@@
 
-%Varying Iterations:
+Varying Iterations:
 
 Varying Quality Factor:
-We define the quality factor in this case for how many times smaller the Euler
-method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for
-$g_{\Delta t}$.
+We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$
+
+Changing the ODE
+
+TODO ashi sort out the figure and match and replace if needbe
 
@@
 \end{figure}
-
-Figure~\ref{fig:run_v_cpu} shows the running time (wall clock) for the Parareal
-algorithm
 
@@
 
-To demonstrate parallelism by space, we solve the two-dimensional wave equation by partioning a two-dimensional space domain amongst multiple processors. Specifically, we look to numerically solve
+For the speaker problem, we designed a problem-specific method to parallelize in space and divide Pierce Hall into subsections such that each processor would receive a division of the building to compute.
+We wish to show strong scaling, in which the computation time of our parallel method will decrease as we increase the number of processors.
+We compare this version to the serial version we did for Homework 4.
 
-\begin{equation*}
-\frac{\partial^2 u}{\partial t^2} = \frac{\partial^2 u}{\partial x^2} + \frac{\partial^2 u}{\partial y^2}
-\end{equation*}
-
-on the two-dimensional domain $(x, y) \in [0,1]\times[0,1]$ for evolution in time.
-We partition the domain by space using a second order central difference approximation, assigning a rectangular portion of the two-dimensional domain to each of a grid of $(P_x, P_y)$ processors.
-This requires the passing of information along the edges of the processors to each other, as shown graphically in the figure below:
-
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width=0.5\linewidth]{decomp.png}
-\caption{Parallelization by space schematic.}
-\label{fig:spaceparallel}
-\end{center}
-\end{figure}
-
-Note that each processor receives information from only those adjacent to it; that is, the top left processor in the graphic does not communicate directly with the bottom right processor in the graphic.
-Rather, information is passed only along shared boundary points to propagate the wave equation.
-We run this code for three different grid sizes: $128\times 128$, $256\times 256$, and $512\times 512$, in both serial and parallel.
-
-In serial, we have:
-
-\begin{center}
-\begin{tabular}{|c|c|} \hline
-($N_x, N_y$) & Time (seconds)\\ \hline
-(128, 128) & 0.987\\ \hline
-(256, 256) & 7.889\\ \hline %real
-(512, 512) & 63.913\\ \hline %real
-\end{tabular}
-\end{center}
-
-In parallel, we have:
-
-\begin{center}
-\begin{tabular}{|c|c|c|c|c|} \hline
-($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
-(128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
-(128, 128) & 4 & 2 & 2.31 \\ \hline
-(128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
-(128, 128) & 8 & 8 & (still queued on Odyssey) \\ \hline
-(256, 256) & 2 & 2 & 2.57 \\ \hline %real
-(256, 256) & 4 & 2 & 11.32 \\ \hline
-(256, 256) & 4 & 4 & 13.65 \\ \hline
-(256, 256) & 8 & 8 & 17.44 \\ \hline
-(512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
-(512, 512) & 4 & 2 & (still queued on Odyssey)\\ \hline
-(512, 512) & 4 & 4 & (still queued on Odyssey)\\ \hline %check
-(512, 512) & 8 & 8 & (still queued on Odyssey) \\ \hline %check
-\end{tabular}
-\end{center}
-
-Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved.
-However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed).
-Hence, this is an example of weak scaling.
-However, an advantage to this paradigm is that because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes.
-Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.
+bksim TODO GET DATA, AS REQUIRED FROM ABOVE, SHOW GRAPHS
 
@@
 
-As we can see, we do see the effects of WEAK scaling
+As we can see, we do see the effects of strong scaling
 
@@
 
-On the other hand, parallelism by space already requires a more problem-specific
-design. Parallelism by space is intuitive, and conceptually, is embarrassingly parallel and easier to create. However, the
+On the other hand, parallelism by space already requires a very problem-specific
+design. The strong scaling efficiency of parallelism by space is high and is
+conceptually embarrassingly parallel and easier to create. However, the
 limitations are in the requirement for the code to be designed for the

### 2014-12-10 paper.tex

--- 2014-12-10 paper.tex
@@
 
-When $k = 2$ and $Q = 100$, for maximal speedup (fewest iterations) of $S(n) = \frac{203n}{3n+200} \approx 67$
+When $k = 2$ and $Q = 100$, for maximal speedup (fewest iterations) of:
+\[
+\begin{aligned}
+S(n) &= \frac{203n}{3n+200} \\
+\lim_{n\to \infty}S(n) &\approx 67
+\end{aligned}
+\]
 
-When $k = 10$ and $Q = 100$, for more iterations and less error, we'd see an Amdahl bounded max speedup of $S(n) = \frac{1011n}{11n+1000} \approx 91$
+When $k = 10$ and $Q = 100$, for more iterations and less error, we'd see an Amdahl bounded max speedup of 
+\[
+\begin{aligned}
+S(n) &= \frac{1011n}{11n+1000} \\
+\lim_{n\to \infty}S(n) &\approx 91
+\end{aligned}
+\]
 
@@
 \section{Results and Discussion}
-
-WESLEY ENDED HERE, BEFORE THIS IS WESLEY APPROVED
 
@@
 
-\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as Higher Order Methods}
-
-TODO team
-
-IDEALLY WE SHOW THE SAME GRAPHS AS THE ABOVE WITH ALL THE VARIATIONS TESTED
-
 \subsection{Parallelism by Space Paradigm}
@@
 Rather, information is passed only along shared boundary points to propagate the wave equation.
-We run this code for three different grid sizes: $128\times 128$, $256\times 256$, and $512\times 512$, in both serial and parallel.
+We run this code for two different grid sizes: $256\times 256$ and $512\times 512$, in both serial and parallel.
 
@@
 ($N_x, N_y$) & Time (seconds)\\ \hline
-(128, 128) & 0.987\\ \hline
+%(128, 128) & 0.987\\ \hline
 (256, 256) & 7.889\\ \hline %real
@@
 \begin{tabular}{|c|c|c|c|c|} \hline
-($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) \\ \hline
-(128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
-(128, 128) & 4 & 2 & 2.31 \\ \hline
-(128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
-(128, 128) & 8 & 8 & (still queued on Odyssey) \\ \hline
-(256, 256) & 2 & 2 & 2.57 \\ \hline %real
-(256, 256) & 4 & 2 & 11.32 \\ \hline
-(256, 256) & 4 & 4 & 13.65 \\ \hline
-(256, 256) & 8 & 8 & 17.44 \\ \hline
-(512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
-(512, 512) & 4 & 2 & (still queued on Odyssey)\\ \hline
-(512, 512) & 4 & 4 & (still queued on Odyssey)\\ \hline %check
-(512, 512) & 8 & 8 & (still queued on Odyssey) \\ \hline %check
+($N_x, N_y$) & $P_x$ & $P_y$ & Time (sec) & Speedup \\ \hline
+%(128, 128) & 2 & 2 & (still queued on Odyssey) & \\ \hline
+%(128, 128) & 4 & 2 & (still queued on Odyssey) & \\ \hline %check
+%(128, 128) & 4 & 4 & (still queued on Odyssey) & \\ \hline
+%(128, 128) & 8 & 8 & (still queued on Odyssey) & \\ \hline
+(256, 256) & 2 & 2 & 2.57 & 3.07\\ \hline %real
+(256, 256) & 4 & 2 & 11.32 & \\ \hline %check
+(256, 256) & 4 & 4 & 13.65 & \\ \hline %check
+(256, 256) & 8 & 8 & 17.44 & \\ \hline %check
+(512, 512) & 2 & 2 & 39.65 & 1.61\\ \hline %real
+(512, 512) & 4 & 2 & (still queued on Odyssey)& \\ \hline
+(512, 512) & 4 & 4 & (still queued on Odyssey)&\\ \hline %check
+(512, 512) & 8 & 8 & (still queued on Odyssey)&\\ \hline %check
 \end{tabular}

### 2014-12-10 paper.tex

--- 2014-12-10 paper.tex
@@
 \usepackage{amsmath} % Needed for math stuff
-%\usepackage{lastpage} % Finds last page
+%\usepackage{lastpage}sss% Finds last page
 \usepackage{amssymb} % Needed for some math symbols
-\usepackage{graphicx} % Needed for graphics
+\usepackage{graphicx}% Needed for graphics
 \usepackage[usenames,dvipsnames]{xcolor} % Needed for graphics and color
-%\usepackage{setspace} % Needed to set single/double spacing
+%\usepackage{setspace}sss% Needed to set single/double spacing
 \usepackage{float} % Better placement of figures & tables
@@
 \usepackage[margin=1in]{geometry}
+% force pictures to go where you want
+\usepackage{float}
+
 \geometry{letterpaper}
+%\addtolength{\oddsidemargin}{-0.875in} 
+%\addtolength{\evensidemargin}{-0.875in} 
+%\addtolength{\textwidth}{1.75in} 
+%\addtolength{\topmargin}{-.875in} 
+%\addtolength{\textheight}{1.75in}
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[H]
 \begin{center}
@@
 
-Example 1:
 \[
@@
 \]
-
-Example 2:
 \[
@@
 We ran the algorithm on different numbers of processors and varied $k$ and $Q$.
-We define the quality factor $Q$ in this case for how many times smaller the
-Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method
-used for $g_{\Delta t}$.
 
-\begin{figure}[htb]
+Varying Iterations:
+
+Varying Quality Factor:
+We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$
+
+Changing the ODE
+
+TODO ashi sort out the figure and match and replace if needbe
+
+\begin{figure}
 \begin{center}
@@
 
-Figure~\ref{fig:run_v_cpu} shows the running time (wall clock) for the Parareal
-algorithm compared to different number of CPUs used. The serial version is the
-Euler method with $Q$ times as many steps. As we can see, the running time
-decreases as more processors are added. However, there is a base time associated
-with setting up the parallel infrastructure, so asymptotically the running time
-is nonzero.
-
-\begin{figure}[htb]
+\begin{figure}
 \begin{center}
@@
 
-\begin{figure}[htb]
+\begin{figure}
 \begin{center}

### 2014-12-10 paper.tex
merge

@@
 \begin{equation}
-S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
+S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)}
 \end{equation}
@@
 
-From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + k(t+ Qt)$.
+From our analysis in equation~\ref{eq:scheme}, when we set $N$ to 1 for the generic setup, we see that our total computation is $t + K(t+ Qt)$.
 Of this, the only parallelizable part is the $kQt$ term.
-Thus to find the expression of the non-parallelizable portion, we simplify:
+Thus to find the expression of the parallelizable portion, we simplify:
 
 \[
-B = \frac{t+kt}{t+kt+kQt} = \frac{1+k}{kQ+k+1}
+AA
 \]
@@
 \[
-\begin{aligned}
-S(n) &= \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{1}{\left( B + \frac{1}{n} (1-B) \right)}\\
-       &= \frac{T(1)}{T(1) \left( \frac{1+k}{kQ+k+1} + \frac{1}{n} (1- \frac{1+k}{kQ+k+1}) \right)} \\
-       &= \frac{n(kQ+k+1)}{n+nk+kQ}
-\end{aligned}
+S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
 \]
-
-As written below, we have used $Q = 100$ with various $k$.
-
-When $k = 2$ and $Q = 100$, for maximal speedup (fewest iterations) of $S(n) = \frac{203n}{3n+200} \approx 67$
-
-When $k = 10$ and $Q = 100$, for more iterations and less error, we'd see an Amdahl bounded max speedup of $S(n) = \frac{1011n}{11n+1000} \approx 91$
-
-It is important to note that this speedup is measured with respect to the parareal agorithm done in parallel.
-This does not compare the Parareal to doing $g_{\textnormal{fine}}$ but rather shows how given infinite processors, our speedup will approach certain asymptotic bounds.
-We do see however large constants in the denominator which means that we will get better speedup as we increase n past a certain size.
 
@@
 \[
-S(n) = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
+S_{n} = \frac{T(1)}{T(n)} = \frac{T(1)}{ T(1) \left( B + \frac{1}{n} (1-B) \right)} = \frac{T(1)}{T(1) \left( 0 + \frac{1}{n} (1-0) \right)} = \frac{T(1)}{T(1) * \left( \frac{1}{n} \right) } = n
 \]
@@
 (128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
-(128, 128) & 4 & 2 & 2.31 \\ \hline
+(128, 128) & 4 & 2 & 2.31 \\ \hline %check
 (128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
@@
 (256, 256) & 2 & 2 & 2.57 \\ \hline %real
-(256, 256) & 4 & 2 & 11.32 \\ \hline
-(256, 256) & 4 & 4 & 13.65 \\ \hline
-(256, 256) & 8 & 8 & 17.44 \\ \hline
+(256, 256) & 4 & 2 & 11.32 \\ \hline %check
+(256, 256) & 4 & 4 & 13.65 \\ \hline %check
+(256, 256) & 8 & 8 & 17.44 \\ \hline %check
 (512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline
@@
 
+\subsection{Parallelism Paradigm Comparisons}
+
+Now with 3 different parallel implementations, two in time with the parareal, one in space, we compare across the methods to see if we can make any statements about general performance of our different paradigms.
+
+TODO WESLEY
+
+TODO compare the computation times (and perhaps a slight change in accuracy of the results) between both parareal methods the best in the sections before hand
+
+TODO bsim and ashi show graphs and some times for the forward Euler parareal
+versions---the fastest ones or both if we have both implemented
+
+\subsection{Comparison to C++ Implementations}
+
+The numpy environment combined with the required structure of MPI required
+certain type/data structure conversions which will take time.not not cunot cu
+
 \subsection{Summary of Performance}
@@
 
-\subsection{Demonstrating Scaling and Efficiency}
+\subsection{Demonstrating Scaling}
 
-\subsection{Parallel in Space}
-
-bsim TODO eventually
+ashi TODO by taking some numbers and following
 
@@
 
-TODO NUMPY WRITE FATST HING
-
 On the other hand, parallelism by space already requires a more problem-specific
@@
 
-TODO WRITE ABOUT BETTER DATA COLLECTION
+TODO Wesley
 

### 2014-12-10 paper.tex
Merge remote-tracking branch 'origin/master'
Conflicts:
	paper.tex

--- 2014-12-10 paper.tex
@@
 \end{figure}
-
+`
 \begin{figure}
@@
 \end{figure}
-
-\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as Higher Order Methods}
-
-TODO team
-
-IDEALLY WE SHOW THE SAME GRAPHS AS THE ABOVE WITH ALL THE VARIATIONS TESTED
 
@@
 (128, 128) & 2 & 2 & (still queued on Odyssey)\\ \hline
-(128, 128) & 4 & 2 & 2.31 \\ \hline %check
+(128, 128) & 4 & 2 & 2.31 \\ \hline
 (128, 128) & 4 & 4 & (still queued on Odyssey) \\ \hline
@@
 (256, 256) & 2 & 2 & 2.57 \\ \hline %real
-(256, 256) & 4 & 2 & 11.32 \\ \hline %check
-(256, 256) & 4 & 4 & 13.65 \\ \hline %check
-(256, 256) & 8 & 8 & 17.44 \\ \hline %check
+(256, 256) & 4 & 2 & 11.32 \\ \hline
+(256, 256) & 4 & 4 & 13.65 \\ \hline
+(256, 256) & 8 & 8 & 17.44 \\ \hline
 (512, 512) & 2 & 2 & (still queued on Odyssey)\\ \hline

### 2014-12-10 paper.tex
new results

--- 2014-12-10 paper.tex
@@
 (256, 256) & 4 & 4 & 13.65 & \\ \hline %check
-(256, 256) & 8 & 8 & 17.44 & \\ \hline %check
+(256, 256) & 8 & 8 & (still queued on Odyssey)& \\ \hline %check17.44
 (512, 512) & 2 & 2 & 39.65 & 1.61\\ \hline %real
-(512, 512) & 4 & 2 & (still queued on Odyssey)& \\ \hline
+(512, 512) & 4 & 2 & 14.46 & 4.42\\ \hline
 (512, 512) & 4 & 4 & (still queued on Odyssey)&\\ \hline %check

### 2014-12-10 paper.tex

@@
 
-WESLEY ENDED HERE, BEFORE THIS IS WESLEY APPROVED
-
 \subsection{Expectations}
@@
 We ran the algorithm on different numbers of processors and varied $k$ and $Q$.
+We define the quality factor $Q$ in this case for how many times smaller the
+Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method
+used for $g_{\Delta t}$.
 
-Varying Iterations:
-
-Varying Quality Factor:
-We define the quality factor in this case for how many times smaller the Euler method used for $g_{\textnormal{fine}}$ is compared to the Euler method used for $g_{\Delta t}$
-
-Changing the ODE
-
-TODO ashi sort out the figure and match and replace if needbe
-
-\begin{figure}
+\begin{figure}[htb]
 \begin{center}
@@
 
-\begin{figure}
+Figure~\ref{fig:run_v_cpu} shows the running time (wall clock) for the Parareal
+algorithm compared to different number of CPUs used. The serial version is the
+Euler method with $Q$ times as many steps. As we can see, the running time
+decreases as more processors are added. However, there is a base time associated
+with setting up the parallel infrastructure, so asymptotically the running time
+is nonzero.
+
+\begin{figure}[htb]
 \begin{center}

### 2014-12-10 paper.tex
updates

@@
 
+\subsection{Summary of Performance}
+\subsubsection{Parallel in Space Paradigm}
 In serial, we have:
@@
 (256, 256) & 2 & 2 & 2.57 & 3.07\\ \hline %real
-(256, 256) & 4 & 2 & 11.32 & \\ \hline %check
-(256, 256) & 4 & 4 & 13.65 & \\ \hline %check
-(256, 256) & 8 & 8 & (still queued on Odyssey)& \\ \hline %check17.44
+(256, 256) & 4 & 2 & 11.32 & 0.70\\ \hline %check
+(256, 256) & 4 & 4 & (still queued on Odyssey) & \\ \hline %check 13.65
+(256, 256) & 8 & 8 & (still queued on Odyssey)& \\ \hline %check 17.44
 (512, 512) & 2 & 2 & 39.65 & 1.61\\ \hline %real
-(512, 512) & 4 & 2 & 14.46 & 4.42\\ \hline
-(512, 512) & 4 & 4 & (still queued on Odyssey)&\\ \hline %check
-(512, 512) & 8 & 8 & (still queued on Odyssey)&\\ \hline %check
+(512, 512) & 4 & 2 & 14.46 & 4.42\\ \hline %real
+(512, 512) & 4 & 4 & 8.02 & 7.97\\ \hline 
+(512, 512) & 8 & 8 & (still queued on Odyssey)&\\ \hline 
 \end{tabular}
@@
 
+\subsection{Demonstrating Scaling and Efficiency}
+\subsubsection{Parallel in Space Paradigm}
 Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved.
-However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed).
+However, scaling is \textbf{not} achieved for the 256$\times$ 256 grid when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed).
+We do see scaling with approximately $50\%$ efficiency for the $512\times 512$ grid as we add more processors.
+From the results, as we go from the $256\times256$ grid to the $512\times 512$ grid, we quadruple the size of the problem; quadrupling the processors from $2\times 2$ to $4\times4$ yields a somewhat linear relationship.
 Hence, this is an example of weak scaling.
-However, an advantage to this paradigm is that because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes.
+An advantage to this paradigm is as follows: because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes.
 Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.
-
-\subsection{Summary of Performance}
-
-TODO team
-
-\subsection{Demonstrating Scaling and Efficiency}
-
-\subsection{Parallel in Space}
-
-bsim TODO eventually
-
-As we can see, we do see the effects of WEAK scaling
-
-TODO analyze
 

### 2014-12-10 paper.tex
added more explanations about figures

--- 2014-12-10 paper.tex
@@
 refinement built on top of a more basic method first computed in serial. As will be discussed and analyzed in the
-methodology section, one such scheme would be the parareal algorithm. As an
-overview, the parareal algorithm first does a coarse (low-order and fast)
+methodology section, one such scheme would be the Parareal algorithm. As an
+overview, the Parareal algorithm first does a coarse (low-order and fast)
 approximation, then seeks to iteratively correct using smaller, more refined
@@
 
-The parareal algorithm, developed by Lions, Maday and Turinici in
+The Parareal algorithm, developed by Lions, Maday and Turinici in
 2001~\cite{lions2001parareal}, is a generalized algorithm which allows for
@@
 
-The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators.
+The beauty of the Parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators.
 The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.
@@
 \includegraphics[width=0.75\linewidth]{parareal_visualization.png}
-\caption{Graphical representation of the parareal scheme.}
+\caption{Graphical representation of the Parareal scheme.}
 \label{fig:pararealvis}
@@
 
-With the assumptions of the previous section, this parareal method will
+With the assumptions of the previous section, this Parareal method will
 approach, with large enough $k$ (correction iterations) to approach the error of
@@
 u_n) = \bar{g_{\textnormal{fine}}}u_n$ and $g_{\Delta t}(t_n, u_n) =
-\bar{g_{\Delta t}} u_n$. As shown by Staff et al. \cite{staff2005stability}, the parareal method becomes
+\bar{g_{\Delta t}} u_n$. As shown by Staff et al. \cite{staff2005stability}, the Parareal method becomes
 
@@
 
-The first $t$ seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$.
+The first $t$ seconds comes from the first coarse approximation, without which the Parareal algorithm degenerates to $g_{\Delta t}$.
 Then for each of the $k$ correction iterations, the $t$ term comes from $n$, element by element evaluations of $g_{\Delta t}$ followed by a parallel $\frac{1}{N}$ evaluation of $g_{\textnormal{fine}}$.
@@
 
-We first try to express $B$ as function of the quality factor $Q$ and the number of correction steps of the parareal $k$.
+We first try to express $B$ as function of the quality factor $Q$ and the number of correction steps of the Parareal $k$.
 
@@
 
-It is important to note that this speedup is measured with respect to the parareal agorithm done in parallel.
+It is important to note that this speedup is measured with respect to the Parareal algorithm done in parallel.
 This does not compare the Parareal to doing $g_{\textnormal{fine}}$ but rather shows how given infinite processors, our speedup will approach certain asymptotic bounds.
@@
 
-With embarassingly parallel problem setups, we can optimally achieve up to $n$ times speedup with $n$ processors though this number if quite ideal and will be very difficult to approach (large scaling will have different memory management and greater communication overheads and costs).
+With embarrassingly parallel problem setups, we can optimally achieve up to $n$ times speedup with $n$ processors though this number if quite ideal and will be very difficult to approach (large scaling will have different memory management and greater communication overheads and costs).
 
@@
 
-We wished to explore certain tests that would demonstrate the effects of changing various parameters be it from the solution schemes in the parareal to the exact system we were trying to solve.
+We wished to explore certain tests that would demonstrate the effects of changing various parameters be it from the solution schemes in the Parareal to the exact system we were trying to solve.
 
@@
 
-\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as a Forward Euler with
-Smaller $\Delta t$}
+\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as a Forward Euler
+with Smaller $\Delta t$}
 
@@
 
+As seen in Figures~\ref{fig:err_v_k} and~\ref{fig:err_v_q}, the error remains
+constant for the parallel versions of the algorithm even as we increase the
+number of correction steps and the quality factor. This is not exactly what we
+expect. We predicted that the error would decrease as we did more correction
+steps, or if we increased the quality of the $g_{\textnormal{fine}}$ operator.
+
 \begin{figure}[htb]
@@
 
+Figure~\ref{fig:run_v_k} shows the running time (wall clock) as a function of
+the number of correction steps. For the parallel implementations, we see an
+increase in running time as we increase the number of correction steps. This is
+to be expected, since increasing the number of correction steps increases the
+number of iterations of our algorithm. Therefore, even though we proved before
+that as the number of iterations $k \to N$, where $N$ is the number of time
+points, the order of convergence of the algorithm converges to the order of
+accuracy of $g_{\textnormal{fine}}$, this comes at the expense of increasing
+computational cost.
+
+\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as Higher Order
+Methods}
+
+We considered using a higher order $g_{\textnormal{fine}}$ in our
+implementation, but we were unable to show higher order convergence of the error
+when we tested our method serially.
+
 \subsection{Parallelism by Space Paradigm}
 
-To demonstrate parallelism by space, we solve the two-dimensional wave equation by partioning a two-dimensional space domain amongst multiple processors. Specifically, we look to numerically solve
+To demonstrate parallelism by space, we solve the two-dimensional wave equation by partitioning a two-dimensional space domain amongst multiple processors. Specifically, we look to numerically solve
 
@@
 (256, 256) & 4 & 4 & 13.65 & \\ \hline %check
-(256, 256) & 8 & 8 & (still queued on Odyssey)& \\ \hline %check17.44
+(256, 256) & 8 & 8 & 17.44 & \\ \hline %check
 (512, 512) & 2 & 2 & 39.65 & 1.61\\ \hline %real
-(512, 512) & 4 & 2 & 14.46 & 4.42\\ \hline
+(512, 512) & 4 & 2 & (still queued on Odyssey)& \\ \hline
 (512, 512) & 4 & 4 & (still queued on Odyssey)&\\ \hline %check
@@
 be ideal. Other optimizations mentioned above could also be implemented. To
-further explore the parareal algorithm would involve many test cases and to see
-how the parareal algorithm compares to other approaches---since the actual time
+further explore the Parareal algorithm would involve many test cases and to see
+how the Parareal algorithm compares to other approaches---since the actual time
 and iterations necessary depend on the system of ODEs to solve.
@@
 problem---how to divide by space so as to allow for the maximum number of even
-divisions. The lack of generality is not as beautiful as the parareal algorithm
+divisions. The lack of generality is not as beautiful as the Parareal algorithm
 but may lend itself to faster speedups.
@@
 All in all, we have implemented and begun to explore some techniques of looking
-for strong and weak scaling efficiencies for solving ODEs. The parareal
+for strong and weak scaling efficiencies for solving ODEs. The Parareal
 algorithm is beautiful in its theoretical advantages---in terms of stability,
@@
 
-Bibtex Citations for the Slides, and for \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
+\section{Appendix}
+\subsection{Explanation of Code Files}
 

### 2014-12-10 paper.tex

@@
 (256, 256) & 2 & 2 & 2.57 & 3.07\\ \hline %real
-(256, 256) & 4 & 2 & 11.32 & \\ \hline %check
-(256, 256) & 4 & 4 & 13.65 & \\ \hline %check
-(256, 256) & 8 & 8 & 17.44 & \\ \hline %check
+(256, 256) & 4 & 2 & 11.32 & 0.70\\ \hline %check
+(256, 256) & 4 & 4 & (still queued on Odyssey) & \\ \hline %check 13.65
+(256, 256) & 8 & 8 & (still queued on Odyssey)& \\ \hline %check 17.44
 (512, 512) & 2 & 2 & 39.65 & 1.61\\ \hline %real
-(512, 512) & 4 & 2 & (still queued on Odyssey)& \\ \hline
-(512, 512) & 4 & 4 & (still queued on Odyssey)&\\ \hline %check
-(512, 512) & 8 & 8 & (still queued on Odyssey)&\\ \hline %check
+(512, 512) & 4 & 2 & 14.46 & 4.42\\ \hline %real
+(512, 512) & 4 & 4 & 8.02 & 7.97\\ \hline 
+(512, 512) & 8 & 8 & (still queued on Odyssey)&\\ \hline 
 \end{tabular}

### 2014-12-10 paper.tex
added more to paper

@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$ explained in the below table.
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$ explained in the below table.
 
@@
 \begin{itemize}
-\item General Scaling with Incrased $N$
+\item General Scaling with Increased $N$
 \item Adjusting $K$ in the Parareal
@@
 \subsection{General Remarks about Results}
-LETS SEE IF I HAVE TO DISCLAIM ANY GLOBAL FAILURES IN OUR RESULTS
+
+In general, we were unable to see the promised theoretical speedups and error
+convergence for the Parareal algorithm. We believe this could be because of
+several factors. First, we may have some subtle bugs in our implementation of
+the Parareal algorithms. Additionally, we only tested our implementation on
+relatively simple problems.  he problem with this testing approach is that the
+fine serial computation of the solution $g_{\textnormal{fine}}$ is very cheap.
+We did not see the running time of the Parareal beat that for the serial
+implementation of $g_{\textnormal{fine}}$. If we tested our implementation for
+more expensive $g_{\textnormal{fine}}$ and more complicated problems, we might
+be able to see to see more speedup and better error convergence.
 
@@
 
-\subsection{Summary of Performance}
-\subsubsection{Parallel in Space Paradigm}
 In serial, we have:
@@
 
-\subsection{Demonstrating Scaling and Efficiency}
-\subsubsection{Parallel in Space Paradigm}
 Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved.
@@
 
-\section{Appendix}
-\subsection{Explanation of Code Files}
-
 \end{document}

### 2014-12-10 paper.tex
np stuff

--- 2014-12-10 paper.tex
@@
 
-TODO NUMPY WRITE FATST HING
+Some drawbacks also include the use of the \verb|mpi4py| library instead of the MPI libraries in C++.
+Because \verb|mpi4py| is a set of Python bindings for MPI, based on the standard MPI-2 C++ bindings, there is additional overhead involved with mapping \verb|mpi4py| commands to the standard C++ library.
+For example, since Python is dynamically typed, additional information about size may first need to be sent as a part of the Python bindings, creating greater overhead as compared to the C++ libraries.
+In addition, C++ compilers may be more optimizable according to specific use case, as specific \verb|-o| flags or other compiler shortcuts may be specified directly, rather than through a standard wrapper class as provided by \verb|mpi4py|.
+
+\textbf{something here about parareal? ashi}
 

### 2014-12-11 paper.tex
trying to get changes
updated paper and documentation

@@
 
-\section{Results and Discussion}
+\section{Experimental Results and Discussion}
 
@@
 
-We see that (HOPE IT WORKS WELL OR ELSE...) Most of the further work would be in
-optimization. We are trying to measure performance in Python, which is not the
-ideal benchmarking language for efficiency and speedups. A port over to C++
-using the MPI libraries in C++ rather than the mpi4py libraries in python would
-be ideal. Other optimizations mentioned above could also be implemented. To
-further explore the Parareal algorithm would involve many test cases and to see
-how the Parareal algorithm compares to other approaches---since the actual time
-and iterations necessary depend on the system of ODEs to solve.
+We do not see the theoretical speedups and error convergence in our experimental
+tests of the Parareal algorithm. In the future, we would like to debug our
+implementation of the algorithm and apply it to more complicated systems of ODEs
+and more expensive and higher fine operators $g_{\textnormal{fine}}$. We also
+would like to test different combinations of $k$ and $Q$ and see how the error
+converges after varying step size.
+
+More further work could be in optimization. We are trying to measure performance
+in Python, which is not the ideal benchmarking language for efficiency and
+speedups. A port over to C++ using the MPI libraries in C++ rather than the
+mpi4py libraries in python would be ideal. Other optimizations mentioned above
+could also be implemented. To further explore the Parareal algorithm would
+involve many test cases and to see how the Parareal algorithm compares to other
+approaches---since the actual time and iterations necessary depend on the system
+of ODEs to solve.
 
@@
 In addition, C++ compilers may be more optimizable according to specific use case, as specific \verb|-o| flags or other compiler shortcuts may be specified directly, rather than through a standard wrapper class as provided by \verb|mpi4py|.
-
-\textbf{something here about parareal? ashi}
 
@@
 
-TODO WRITE ABOUT BETTER DATA COLLECTION
-
-\section{Conclusion and Future Work}
+\section{Conclusion}
 

### 2014-12-11 paper.tex
updated abstract

@@
 \begin{abstract}
-We look to achieve strong scaling in parallelizing numerical ODE methods.
-We test both parallelism by result space as well as parallelism by time, with the Parareal algorithm, to solve various ODEs.
-We measure performance and compare in terms of accuracy, speedup and efficiency.
-We provide some theoretical analysis for the Parareal algorithm and compare its tradeoff to paralleism by space, For the Parareal agorithm, we observe
-light speedups in our small scale tests up to 64 processors on Harvard's
-Odyssey cluster. We analyze possible reasons as to why our implementation
-may not be ideal and some tradeoffs that are taken in the Parareal
-implementation. Further optimizations are proposed and rationalized as next
-steps including ports of to C++ using BLAS.
-Our code is written in python using mpi4py.
-We compare tests from various data sources from basic exponential ODEs to larger grids of ODEs to solve for in applications like diffusion and sound wave propagation.
+We look to achieve strong scaling in parallelizing numerical ODE methods. We
+test both parallelism by result space as well as parallelism by time, with the
+Parareal algorithm, to solve various ODEs. We measure performance and compare in
+terms of accuracy, speedup and efficiency. We provide some theoretical analysis
+for the Parareal algorithm's runtime, speedup, error convergence, and stability.
+For the Parareal agorithm, we observe preliminary promising but not convincing
+results in our small scale tests up to 32 processors on Harvard's Odyssey
+computing cluster. We analyze possible reasons as to why our implementation may
+not be ideal and some tradeoffs that are taken in the Parareal implementation.
+Further optimizations are proposed and rationalized as next steps including
+ports of to C++ using BLAS. Our code is written in python using mpi4py. We
+compare tests from various data sources from basic exponential ODEs to larger
+grids of ODEs to solve for in applications like diffusion and sound wave
+propagation.
 
@@
 the Parareal algorithms. Additionally, we only tested our implementation on
-relatively simple problems.  he problem with this testing approach is that the
+relatively simple problems. The problem with this testing approach is that the
 fine serial computation of the solution $g_{\textnormal{fine}}$ is very cheap.

### 2014-12-11 paper.tex
PAPER PUSH YES THEORY DONE

--- 2014-12-10 paper.tex
@@
 refinement built on top of a more basic method first computed in serial. As will be discussed and analyzed in the
-methodology section, one such scheme would be the Parareal algorithm. As an
-overview, the Parareal algorithm first does a coarse (low-order and fast)
+methodology section, one such scheme would be the parareal algorithm. As an
+overview, the parareal algorithm first does a coarse (low-order and fast)
 approximation, then seeks to iteratively correct using smaller, more refined
@@
 
-The Parareal algorithm, developed by Lions, Maday and Turinici in
+The parareal algorithm, developed by Lions, Maday and Turinici in
 2001~\cite{lions2001parareal}, is a generalized algorithm which allows for
@@
 
-The beauty of the Parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators.
+The beauty of the parareal method is that it can be adapted into many different versions with different choices of the $g_{\Delta t}$ and $g_{\textnormal{fine}}$ solution operators.
 The $g_{\textnormal{fine}}$ solution is a more computationally expensive method but assumed to have lesser error.
@@
 
-\begin{figure}[H]
+\begin{figure}[htb]
 \begin{center}
 \includegraphics[width=0.75\linewidth]{parareal_visualization.png}
-\caption{Graphical representation of the Parareal scheme.}
+\caption{Graphical representation of the parareal scheme.}
 \label{fig:pararealvis}
@@
 
-With the assumptions of the previous section, this Parareal method will
+With the assumptions of the previous section, this parareal method will
 approach, with large enough $k$ (correction iterations) to approach the error of
@@
 u_n) = \bar{g_{\textnormal{fine}}}u_n$ and $g_{\Delta t}(t_n, u_n) =
-\bar{g_{\Delta t}} u_n$. As shown by Staff et al. \cite{staff2005stability}, the Parareal method becomes
+\bar{g_{\Delta t}} u_n$. As shown by Staff et al. \cite{staff2005stability}, the parareal method becomes
 
@@
 
-The first $t$ seconds comes from the first coarse approximation, without which the Parareal algorithm degenerates to $g_{\Delta t}$.
+The first $t$ seconds comes from the first coarse approximation, without which the parareal algorithm degenerates to $g_{\Delta t}$.
 Then for each of the $k$ correction iterations, the $t$ term comes from $n$, element by element evaluations of  $g_{\Delta t}$ followed by a parallel $\frac{1}{N}$ evaluation of $g_{\textnormal{fine}}$.
@@
 
-To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$ explained in the below table.
+To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters - $\alpha$, $\beta$ and $\gamma$ explained in the below table.
+The expressions measure the communitation times spent by each processor (on average as not all processors may communicate equally depending on the size of the problem).
+This helps model the communication overhead.
 
@@
 $\gamma$ & Inverse computational performance. (sec/flop)\\
+$n$ & Number of processors\\
+$P$ & Abstracted problem size\\
 \hline
@@
 \subsubsection{Parareal}
-TODO THIS IS CONFUSING Wesley
+For the parareal method, the communications are solely group communications, a broadcast and gather for each corrective iteration, so 2k $\alpha$.
+The $\beta$ term is from the amount being sent - which is one vector for the results of the $g_{\Delta t}$.
+Finally the cost of the computation we seek to multiply $\gamma$ by an abstract function $\xi(P_n)$ which returns the cost in flops of computing on the input block, $P_n$.
+\[
+C_{\textnormal{parareal}} = \alpha*2k +  \beta \cdot kt + \gamma \cdot \xi(kQt/n)
+\]
 
 \subsubsection{Parallel By Space}
-TODO THIS IS CONFUSING TOO Wesley
+The main communication of the parallel in space method is the sendrecv (a round robin style send and receive for the ghost pounts - the corners have 2 pairs of sendrecvs) which contribute to 4 $\alpha$ plus the initial scatter and final gather each with an $\alpha$ term.
+The amount comunicated is the answer as well as the original broadcasted space which includes boundary points.
+Then, as before, there will be the gamma term multiplied into the computation of the component - given again by an abstract function $\xi(P_n)$ which returns the cost in flops of computing on the input block, $P_n$.
+\[
+C_{\textnormal{paraspace}} \leq \alpha \cdot 6 + \beta \cdot P/n + \gamma \cdot \xi(P/n)
+\]
 
@@
 
-We first try to express $B$ as function of the quality factor $Q$ and the number of correction steps of the Parareal $k$.
+We first try to express $B$ as function of the quality factor $Q$ and the number of correction steps of the parareal $k$.
 
@@
 
-It is important to note that this speedup is measured with respect to the Parareal algorithm done in parallel.
+It is important to note that this speedup is measured with respect to the parareal agorithm done in parallel.
 This does not compare the Parareal to doing $g_{\textnormal{fine}}$ but rather shows how given infinite processors, our speedup will approach certain asymptotic bounds.
@@
 
-With embarrassingly parallel problem setups, we can optimally achieve up to $n$ times speedup with $n$ processors though this number if quite ideal and will be very difficult to approach (large scaling will have different memory management and greater communication overheads and costs).
+With embarassingly parallel problem setups, we can optimally achieve up to $n$ times speedup with $n$ processors though this number if quite ideal and will be very difficult to approach (large scaling will have different memory management and greater communication overheads and costs).
 
@@
 
-We wished to explore certain tests that would demonstrate the effects of changing various parameters be it from the solution schemes in the Parareal to the exact system we were trying to solve.
+We wished to explore certain tests that would demonstrate the effects of changing various parameters be it from the solution schemes in the parareal to the exact system we were trying to solve.
 
@@
 \begin{itemize}
-\item General Scaling with Increased $N$
+\item General Scaling with Incrased $N$
 \item Adjusting $K$ in the Parareal
@@
 \subsection{Expectations}
-We hoped to show that under the properly tuned parameters (namely the $k$ number of iterations),
+We hoped to show that under the properly tuned parameters (namely the $k$ number of iterations), we would get error for the Parareal method that would approach that of running $g_{\textnormal{fine}}$.
+There would be some convergence above which increasing $k$ would not help convergence by as much.
+The expectation was that we would only being to see the best speedups for the Parareal algorithm with large $n$ due to the strong scaling.
+With limited access to processors, we expected to see speedup comparing the Parareal algorithm to itself but not necessaily to the evaluation of $g_{\textnormal{fine}}$.
+Finally, we fully expect the parallel in space algorithm to demonstrate good weak scaling, as well as for it to be faster overall since it is a more embarassingly parallel design.
+As disclaimed before, the main tradeoff for the parallel in space method is the lack of extensiblily across different ODEs applications.
 
 \subsection{General Remarks about Results}
+LETS SEE IF I HAVE TO DISCLAIM ANY GLOBAL FAILURES IN OUR RESULTS
 
-In general, we were unable to see the promised theoretical speedups and error
-convergence for the Parareal algorithm. We believe this could be because of
-several factors. First, we may have some subtle bugs in our implementation of
-the Parareal algorithms. Additionally, we only tested our implementation on
-relatively simple problems.  he problem with this testing approach is that the
-fine serial computation of the solution $g_{\textnormal{fine}}$ is very cheap.
-We did not see the running time of the Parareal beat that for the serial
-implementation of $g_{\textnormal{fine}}$. If we tested our implementation for
-more expensive $g_{\textnormal{fine}}$ and more complicated problems, we might
-be able to see to see more speedup and better error convergence.
-
-\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as a Forward Euler
-with Smaller $\Delta t$}
+\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as a Forward Euler with
+Smaller $\Delta t$}
 
@@
 
-As seen in Figures~\ref{fig:err_v_k} and~\ref{fig:err_v_q}, the error remains
-constant for the parallel versions of the algorithm even as we increase the
-number of correction steps and the quality factor. This is not exactly what we
-expect. We predicted that the error would decrease as we did more correction
-steps, or if we increased the quality of the $g_{\textnormal{fine}}$ operator.
-
-\begin{figure}[H]
+\begin{figure}[htb]
 \begin{center}
@@
 
-Figure~\ref{fig:run_v_k} shows the running time (wall clock) as a function of
-the number of correction steps. For the parallel implementations, we see an
-increase in running time as we increase the number of correction steps. This is
-to be expected, since increasing the number of correction steps increases the
-number of iterations of our algorithm. Therefore, even though we proved before
-that as the number of iterations $k \to N$, where $N$ is the number of time
-points, the order of convergence of the algorithm converges to the order of
-accuracy of $g_{\textnormal{fine}}$, this comes at the expense of increasing
-computational cost.
-
-\subsection{Comparison to Serial with $g_{\textnormal{fine}}$ as Higher Order
-Methods}
-
-We considered using a higher order $g_{\textnormal{fine}}$ in our
-implementation, but we were unable to show higher order convergence of the error
-when we tested our method serially.
-
 \subsection{Parallelism by Space Paradigm}
 
-To demonstrate parallelism by space, we solve the two-dimensional wave equation by partitioning a two-dimensional space domain amongst multiple processors. Specifically, we look to numerically solve
+To demonstrate parallelism by space, we solve the two-dimensional wave equation by partioning a two-dimensional space domain amongst multiple processors. Specifically, we look to numerically solve
 
@@
 (256, 256) & 2 & 2 & 2.57 & 3.07\\ \hline %real
-(256, 256) & 4 & 2 & 11.32 & 0.70\\ \hline %check
-(256, 256) & 4 & 4 & (still queued on Odyssey) & \\ \hline %check 13.65
-(256, 256) & 8 & 8 & (still queued on Odyssey)& \\ \hline %check 17.44
+(256, 256) & 4 & 2 & 11.32 & \\ \hline %check
+(256, 256) & 4 & 4 & 13.65 & \\ \hline %check
+(256, 256) & 8 & 8 & (still queued on Odyssey)& \\ \hline %check17.44
 (512, 512) & 2 & 2 & 39.65 & 1.61\\ \hline %real
-(512, 512) & 4 & 2 & 14.46 & 4.42\\ \hline %real
-(512, 512) & 4 & 4 & 8.02 & 7.97\\ \hline 
-(512, 512) & 8 & 8 & (still queued on Odyssey)&\\ \hline 
+(512, 512) & 4 & 2 & 14.46 & 4.42\\ \hline
+(512, 512) & 4 & 4 & (still queued on Odyssey)&\\ \hline %check
+(512, 512) & 8 & 8 & (still queued on Odyssey)&\\ \hline %check
 \end{tabular}
@@
 Note that as the workload assigned to each processing element stays constant and more processors are used to solve a larger problem, near-linear scaling is achieved.
-However, scaling is \textbf{not} achieved for the 256$\times$ 256 grid when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed).
-We do see scaling with approximately $50\%$ efficiency for the $512\times 512$ grid as we add more processors.
-From the results, as we go from the $256\times256$ grid to the $512\times 512$ grid, we quadruple the size of the problem; quadrupling the processors from $2\times 2$ to $4\times4$ yields a somewhat linear relationship.
+However, scaling is \textbf{not} achieved when we simply add more processing elements to try to solve the same-sized problem, as there is greater constant startup costs which do not necessarily improve the speed (or perhaps even decrease speed).
 Hence, this is an example of weak scaling.
-An advantage to this paradigm is as follows: because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes.
+However, an advantage to this paradigm is that because each processor only communicates with its nearest neighbors, communication overhead is relatively constant regardless of how large problem size / number of processors used becomes.
 Therefore, such a paradigm should scale well to larger problem sizes or more fine grid spacings.
+
+\subsection{Summary of Performance}
+
+TODO team
+
+\subsection{Demonstrating Scaling and Efficiency}
+
+\subsection{Parallel in Space}
+
+bsim TODO eventually
+
+As we can see, we do see the effects of WEAK scaling
+
+TODO analyze
 
@@
 be ideal. Other optimizations mentioned above could also be implemented. To
-further explore the Parareal algorithm would involve many test cases and to see
-how the Parareal algorithm compares to other approaches---since the actual time
+further explore the parareal algorithm would involve many test cases and to see
+how the parareal algorithm compares to other approaches---since the actual time
 and iterations necessary depend on the system of ODEs to solve.
 
-Some drawbacks also include the use of the \verb|mpi4py| library instead of the MPI libraries in C++.
-Because \verb|mpi4py| is a set of Python bindings for MPI, based on the standard MPI-2 C++ bindings, there is additional overhead involved with mapping \verb|mpi4py| commands to the standard C++ library.
-For example, since Python is dynamically typed, additional information about size may first need to be sent as a part of the Python bindings, creating greater overhead as compared to the C++ libraries.
-In addition, C++ compilers may be more optimizable according to specific use case, as specific \verb|-o| flags or other compiler shortcuts may be specified directly, rather than through a standard wrapper class as provided by \verb|mpi4py|.
-
-\textbf{something here about parareal? ashi}
+TODO NUMPY WRITE FATST HING
 
@@
 problem---how to divide by space so as to allow for the maximum number of even
-divisions. The lack of generality is not as beautiful as the Parareal algorithm
+divisions. The lack of generality is not as beautiful as the parareal algorithm
 but may lend itself to faster speedups.
@@
 All in all, we have implemented and begun to explore some techniques of looking
-for strong and weak scaling efficiencies for solving ODEs. The Parareal
+for strong and weak scaling efficiencies for solving ODEs. The parareal
 algorithm is beautiful in its theoretical advantages---in terms of stability,
@@
 
+Bibtex Citations for the Slides, and for \url{https://www.sharcnet.ca/help/index.php/Measuring_Parallel_Scaling_Performance}
+
 \end{document}

### 2014-12-11 paper.tex

--- 2014-12-11 paper.tex
@@
 A fully parallelizable approach allows for $n$ times speedup and no computational tradeoffs if we ignore communication.
-As we are dividing our processors into a processor grid (dividing a 2D space into a 2D arrangement of processors) we are only limited by when the processor grid size approaches the problem size - upon which further processors would not receive any tasks since that would require a division by less than a unit of computation.
+As we are dividing our processors into a processor grid (dividing a 2D space into a 2D arrangement of processors) we are only limited by when the processor grid size approaches the problem size---upon which further processors would not receive any tasks since that would require a division by less than a unit of computation.
 
@@
 To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$ explained in the below table.
+The expressions measure the communitation times spent by each processor (on average as not all processors may communicate equally depending on the size of the problem).
+This helps model the communication overhead.
 
@@
 \subsection{Expectations}
-We hoped to show that under the properly tuned parameters (namely the $k$ number of iterations),
+We hoped to show that under the properly tuned parameters (namely the $k$ number of iterations), we would get error for the Parareal method that would approach that of running $g_{\textnormal{fine}}$.
+There would be some convergence above which increasing $k$ would not help convergence by as much.
+The expectation was that we would only being to see the best speedups for the Parareal algorithm with large $n$ due to the strong scaling.
+With limited access to processors, we expected to see speedup comparing the Parareal algorithm to itself but not necessaily to the evaluation of $g_{\textnormal{fine}}$.
+Finally, we fully expect the parallel in space algorithm to demonstrate good weak scaling, as well as for it to be faster overall since it is a more embarassingly parallel design.
+As disclaimed before, the main tradeoff for the parallel in space method is the lack of extensiblily across different ODEs applications.
 

### 2014-12-11 paper.tex
fixed typos and added docs

@@
 for the Parareal algorithm's runtime, speedup, error convergence, and stability.
-For the Parareal agorithm, we observe preliminary promising but not convincing
+For the Parareal algorithm, we observe preliminary promising but not convincing
 results in our small scale tests up to 32 processors on Harvard's Odyssey
@@
 Further optimizations are proposed and rationalized as next steps including
-ports of to C++ using BLAS. Our code is written in python using mpi4py. We
+ports of to C++ using BLAS. Our code is written in Python using mpi4py. We
 compare tests from various data sources from basic exponential ODEs to larger
-grids of ODEs to solve for in applications like diffusion and sound wave
-propagation.
+grids of ODEs to solve for in applications like sound wave propagation.
 
@@
 Solving systems of differential equations can be a computationally expensive
-task. The error of most algorithms scales on the stepsize of the discretization
-of time. However, stepsize in time is also proportional to the computation
+task. The error of most algorithms scales on the step size of the discretization
+of time. However, step size in time is also proportional to the computation
 required. It would be nice to allow for strong scaling where a problem can be
@@
 method is expected to outperform the parallel in time paradigm simply because
-the task is embarassingly parallel, being able to be naturally divided
+the task is embarrassingly parallel, being able to be naturally divided
 into smaller but identical in method problems. There is no serial component in
@@
 Performance of parallel algorithms are usually measured in terms of 3 different
-metrics: speedup, strong scalling efficiency and weak scaling efficiency.
+metrics: speedup, strong scaling efficiency and weak scaling efficiency.
 Strong and weak scaling are common measures for the different ways a program can
-benefit from increased parallellism and more processors as well as the
+benefit from increased parallelism and more processors as well as the
 limitations that dictate the speedups~\cite{parallelscaling}. In each of the below subsections, $T(n)$ represents the computation of using $n$ processors and the same notation for speedup.
@@
 The final principle when discussing parallel performance is the potential existence of an Amdahl speedup limit.
-Amdahl's Law seeks to describe the maximum parallel speedup in an overall piece of code when the problem is not embarassingly parallel --- not all of the system can be divided up in parallel.
+Amdahl's Law seeks to describe the maximum parallel speedup in an overall piece of code when the problem is not embarrassingly parallel---not all of the system can be divided up in parallel.
 The parallel in time methods must always have some serial component due to the time-dependent nature of the ODE's which will fall under the limitations stated by Amdahl's Law.
@@
 To delve one step deeper in our analysis, parallel communications are usually expressed by architecture parameters---$\alpha$, $\beta$ and $\gamma$ explained in the below table.
-The expressions measure the communitation times spent by each processor (on average as not all processors may communicate equally depending on the size of the problem).
+The expressions measure the communication times spent by each processor (on average as not all processors may communicate equally depending on the size of the problem).
 This helps model the communication overhead.
@@
 \subsubsection{Parallel By Space}
-The main communication of the parallel in space method is the sendrecv (a round robin style send and receive for the ghost pounts - the corners have 2 pairs of sendrecvs) which contribute to 4 $\alpha$ plus the initial scatter and final gather each with an $\alpha$ term.
-The amount comunicated is the answer as well as the original broadcasted space which includes boundary points.
-Then, as before, there will be the gamma term multiplied into the computation of the component - given again by an abstract function $\xi(P_n)$ which returns the cost in flops of computing on the input block, $P_n$.
+The main communication of the parallel in space method is the sendrecv (a round
+robin style send and receive for the ghost points - the corners have 2 pairs of
+sendrecvs) which contribute to 4 $\alpha$ plus the initial scatter and final
+gather each with an $\alpha$ term. The amount communicated is the answer as well
+as the original broadcasted space which includes boundary points. Then, as
+before, there will be the gamma term multiplied into the computation of the
+component---given again by an abstract function $\xi(P_n)$ which returns the
+cost in flops of computing on the input block, $P_n$. 
 \[
@@
 
-Because this task is embarassingly parallel and nearly 100\% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):
+Because this task is embarrassingly parallel and nearly 100\% of the computation code can be parallelized (disregarding any uneven divisions of the space by the number of processors), we achieve the Amdahl's Law optimal speedup bound of (with $B$ of 0):
 
@@
 \subsection{Expectations}
-We hoped to show that under the properly tuned parameters (namely the $k$ number of iterations), we would get error for the Parareal method that would approach that of running $g_{\textnormal{fine}}$.
-There would be some convergence above which increasing $k$ would not help convergence by as much.
-The expectation was that we would only being to see the best speedups for the Parareal algorithm with large $n$ due to the strong scaling.
-With limited access to processors, we expected to see speedup comparing the Parareal algorithm to itself but not necessaily to the evaluation of $g_{\textnormal{fine}}$.
-Finally, we fully expect the parallel in space algorithm to demonstrate good weak scaling, as well as for it to be faster overall since it is a more embarassingly parallel design.
-As disclaimed before, the main tradeoff for the parallel in space method is the lack of extensiblily across different ODEs applications.
+We hoped to show that under the properly tuned parameters (namely the $k$ number
+of iterations), we would get error for the Parareal method that would approach
+that of running $g_{\textnormal{fine}}$. There would be some convergence above
+which increasing $k$ would not help convergence by as much. The expectation was
+that we would only being to see the best speedups for the Parareal algorithm
+with large $n$ due to the strong scaling. With limited access to processors, we
+expected to see speedup comparing the Parareal algorithm to itself but not
+necessarily to the evaluation of $g_{\textnormal{fine}}$. Finally, we fully
+expect the parallel in space algorithm to demonstrate good weak scaling, as well
+as for it to be faster overall since it is a more embarrassingly parallel design.
+As disclaimed before, the main tradeoff for the parallel in space method is the
+lack of extensibility across different ODEs applications.
 

### 2014-12-11 poster/main.tex
fixed poster

@@
 The Parareal algorithm, developed by Lions, Maday and Turinici in 2001, is
-a general algorithm that allows parallelization in time. It does so by using a
-cheaper (lower order or lesser resolution) approximation first, and then makes
+a general algorithm that allows parallelization in time. It does so by first using a
+cheaper (lower order or lesser resolution) approximation, and then makes
 corrections in parallel. The entire process is then iterative and tuned for a
